---
title: "title"
author: "David Herman"
date: [REDACTED]
output:
  html_document:
    embed-resources: true
    toc-location: right
    df_print: paged
    theme: united
    highlight: tango
    toc: true
    toc_depth: 4
    number-sections: true
    code_folding: hide
    fig_caption: true
    fig_width: 7
    fig_height: 5
    keep_md: true
    css: styles.css
editor: visual
cap-location: margin
---

```{r setup, echo = F}
knitr::opts_chunk$set(echo = F, warning = F, message = F) 
```

```{r Import & load data/packages/functions, message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls())

source("../src/config.R")
source("../src/functions.R")

# Load libraries -------------------------------------------------------
load_libraries(PACKAGES)

# Client Authentification ----------------------------------------------
# do not displayed
```

#### Patient's distribution during follow-up

```{r obtain the number of patients/IDs under follow-up by sex}
# df of interest with follow up days, patient uuid and sex variables
sex_follow_up <- rare_disease_followup_durations_days %>%
  left_join(rare_disease_demographics_new, by = "patient_uuid")

# Extract the unique patient_uuid under follow up by age group and sex
follow_up_patients_uuid <- sex_follow_up %>%
  
  group_by(patient_uuid, sex) %>%
  
  reframe(age_range = seq(floor(first_record_age_days/365.25),
                      floor(last_record_age_days/365.25))) %>%
  
  mutate(annual_age_groups = cut( 
    age_range,
    breaks = c(0:12, Inf),
    right = F, 
    include.lowest = T)) %>%

  filter(age_range < 12) %>%
  
  dplyr::select(-age_range)


# Count the number of patients under follow by group of age and sex
summary_count_uuid <- follow_up_patients_uuid %>%
  
  group_by(sex, annual_age_groups) %>%
  
  reframe(n = n_distinct(patient_uuid)) %>%
  
  bind_rows(
    follow_up_patients_uuid %>%
      
      group_by(annual_age_groups) %>%
      
      reframe(n = n_distinct(patient_uuid)) %>%
      
      mutate(sex = 'patients')
  ) %>%
  
  pivot_wider(values_from = n, names_from = sex) %>%
  
  mutate(
    percent_total = (patients / first(patients)) * 100,  # Percentage of total follow-up by age group
    percent_male = (male / patients) * 100,  # Percentage of total patients that are male
    percent_female = (female / patients) * 100  # Percentage of total patients that are female
  )

# Transpose data
summary_transposed <- summary_count_uuid %>%
  
  pivot_longer(cols = c(patients, male, female), names_to = 'Category') %>%
  
  pivot_wider(values_from = value, names_from = annual_age_groups)

# Display the summary table
summary_transposed %>%
  kable(
    caption = "Table 1. Summary of total number of patients under follow-up for each group of age based on first and last linked medical record in the RWD Health Rare Disorders Databank",
    align = "c",
    format = "html"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = F,
    position = "center"
  ) %>%
  column_spec(1, bold = TRUE, background = "#f2f2f2")

summary_transposed %>%
  write_csv("~/RWE-rare_disease/Age groups analyses/output/tables/summary_patients_table.csv")
```

```{r plot of patient under follow with reasons for loss, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, dpi=300}
# Identify deceased patients
deceased_patients <- rare_disease_primary_diagnosis %>% 
  filter(primary_diagnosis == "Deceased") %>% 
  pull(patient_uuid)

# Prepare the follow-up data
follow_up_plot_data <- rare_disease_followup_durations_days %>%
  distinct(patient_uuid, first_record_age_days, last_record_age_days) %>%
  ungroup() %>%
  mutate(first_record_age_years = first_record_age_days / 365.25,
         last_record_age_years = last_record_age_days / 365.25,
         followup_duration = last_record_age_years - first_record_age_years,
         last_record_age_years_plot = pmin(last_record_age_years, 12),
         censored = last_record_age_years > 12,
         loss_reason = case_when(
           patient_uuid %in% deceased_patients ~ "Death",
           censored ~ "Age cut-off",
           TRUE ~ "Lost to follow-up"
         )) %>%
  arrange(desc(followup_duration)) %>%  # Ordering by descending follow-up duration
  mutate(patient_label = factor(row_number(), levels = rev(row_number()))) %>%  # Assign rank explicitly
  dplyr::select(-first_record_age_days, -last_record_age_days)

# Define colors for loss to follow-up reasons
colors_ltfu <- c("Death" = "red", "Lost to follow-up" = "black", "Age cut-off" = "blue")

# Create ggplot
follow_up_plot <- ggplot(follow_up_plot_data) +
  # Mark end of follow-up
  geom_point(aes(x = last_record_age_years_plot,
                 y = patient_label,
                 color = loss_reason),
             shape = 18,
             size = 2.5) +  
  # Add dashed follow-up lines
  geom_linerange(aes(y = patient_label, 
                     xmin = first_record_age_years,
                     xmax = last_record_age_years_plot),
                 size = .4,
                 linetype = "dashed",
                 colour = "black") +
  # Add vertical lines for patients at the beggining of follow-up
  geom_segment(aes(x = first_record_age_years, xend = first_record_age_years, 
                   y = as.numeric(patient_label) - 0.3, 
                   yend = as.numeric(patient_label) + 0.3),
               size = 0.5, color = "black") +
  geom_vline(xintercept = 12, linetype = "dashed", color = "black", size = 0.4) +
  # Add text annotation above the line
  # annotate("text", x = 11.5, y = max(as.numeric(follow_up_plot_data$patient_label)) - 2, 
  #          label = "End of Study", color = "black", size = 2, fontface = "bold", vjust = 50, hjust = .7) +
  # Custom colors for legend
  scale_color_manual(name = "Follow-Up Status", values = colors_ltfu) +
  # Adjust x-axis
  scale_x_continuous(breaks = seq(0, 12, 1), limits = c(0, 12)) +
  labs(x = "Patient age (in years)", y = "Patient number") +
  # Clean theme and formatting
  theme_minimal() +
  theme(
    panel.grid = element_blank(), 
    panel.background = element_rect(fill = "white", color = NA), 
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7, face = "bold", margin = margin(r = -18)),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 8),
    legend.title = element_text(size = 9),
    legend.position = "bottom",
    legend.text = element_text(size = 9)
  )

follow_up_plot

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/follow_up_plot.png", 
       follow_up_plot, units = "mm", width = 200, height = 150 , dpi = 300)
```

#### Seizure history

```{r Seizure frequency derivation, echo=TRUE, message=FALSE, warning=FALSE}
seizure_subgrouping <-
  
  # starting from processed seizure dataset
  rare_disease_seizure_EEG_processed %>%
  
  # select variables of interest
  dplyr::select(patient_uuid,
                seizure_history_age_days,
                seizure_history_value_perday_aggr) %>%

  mutate( 
    
    # create variable to obtain the patients with at least one record of seizure per each group of age
    seizure_record = "Yes",
    
    age_seizure = seizure_history_age_days / 365.25,
    
    # Derivation of age at seizure in 1-year intervals
    annual_age_groups = cut(
      age_seizure,
      breaks = c(0:12, Inf),
      right = F,
      include.lowest = T)
  )  %>%
  
  # keep only records of age of interest and 0-2 years old
  filter(age_seizure < 12) %>%
  
  # Select relevant columns
  dplyr::select(
    patient_uuid,
    age_seizure,
    seizure_history_age_days,
    annual_age_groups,
    seizure_history_value_perday_aggr,
    seizure_record
  )


# Adding rows with patients under follow up without seizure per day records during a specific year of life
seizure_subgrouping <- follow_up_patients_uuid %>%
  
  left_join(seizure_subgrouping,
            by = c("patient_uuid", "annual_age_groups")) %>%
  
  # assuming that the patient has been under follow up during the entire year if had a record for the specific age group
  mutate(
    seizure_record = replace_na(seizure_record, "No")) %>%
  
  arrange(patient_uuid, seizure_history_age_days)
```

##### Seizure frequency

```{r violin plot of seizure frequencies, fig.align='center', message=FALSE, warning=FALSE, dpi=300, include=FALSE}
# Seizure frequency plot
seizure_violin_plot <- ggplot(seizure_subgrouping, aes(x = annual_age_groups, y = seizure_history_value_perday_aggr)) +
  
  geom_violin(trim = TRUE, scale = "count", width = 0.9) +
  
  geom_boxplot(width = 0.4, fill = "white") +
  
  labs(
    x = "Age Group (in years)",
    y = "Seizures frequency per day all patient records",
    fill = "Age group"
  ) +
  
  scale_fill_manual(values = c(
    "(0-2]" = "blue", 
    "(2-12]" = "green")) +
  
  theme_minimal() +
  
  scale_y_continuous(breaks = seq(0, 750, 100), limits = c(0, 750)) +
  
  theme(
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

seizure_violin_plot
```

```{r summary of seizure frequencies, echo=TRUE, fig.align='center', message=FALSE, warning=FALSE}
# calculate the intra patient mean to reduce variability
seizure_summary_per_patient <- seizure_subgrouping %>%
  
  group_by(patient_uuid, annual_age_groups) %>%
  
  # intra-patient mean and median seizure freq. for each group of age
  summarise(
    mean_seizures_per_day_patient =
      mean(seizure_history_value_perday_aggr, na.rm = TRUE),
    median_seizures_per_day_patient =
      median(seizure_history_value_perday_aggr, na.rm = TRUE),
    .groups = "drop"
  )
  
seizure_summary_age_group <- seizure_summary_per_patient %>%
  
  group_by(annual_age_groups) %>%
  
  # mean, median and dispersion (deviation and range-based measures) metrics for each group of age
  summarise(
    mean_seizures_per_day = round(
      mean(mean_seizures_per_day_patient, na.rm = TRUE), 1),
    sd_seizures_per_day = round(
      sd(mean_seizures_per_day_patient, na.rm = TRUE), 1),
    median_seizures_per_day = round(
      median(median_seizures_per_day_patient, na.rm = TRUE), 1),
    seizures_per_day_p25 = round(
      quantile(median_seizures_per_day_patient, 0.25, na.rm = TRUE), 1),
    seizures_per_day_p75 = round(
      quantile(median_seizures_per_day_patient, 0.75, na.rm = TRUE), 1),
    patients = n_distinct(patient_uuid),
    .groups = "drop"
  ) %>%
  
  arrange(annual_age_groups)
```

```{r violin plot seizures per patient summary table, message=FALSE, warning=FALSE, dpi=300, include=FALSE}
seizure_violin_plot_per_patient <- ggplot(seizure_summary_per_patient, aes(x = annual_age_groups, y =  median_seizures_per_day_patient)) +
  
  geom_violin(trim = TRUE, scale = "count", width = 0.9) +
  
  geom_boxplot(width = 0.4, fill = "white") +
  
  labs(
    x = "Age Group (in years)",
    y = "Seizures frequency per day per patient",
    fill = "Age group"
  ) +
  
  scale_fill_manual(values = c(
    "(0-2]" = "blue", 
    "(2-12]" = "green")) +
  
  theme_minimal() +
  
  scale_y_continuous(breaks = seq(0, 400, 100), limits = c(0, 400)) +
  
  theme(
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

seizure_violin_plot_per_patient
```

```{r seizure summary table, echo=TRUE, message=FALSE, warning=FALSE}
# transpose the table
seizure_summary_t <- seizure_summary_age_group %>%
  
  mutate(
    mean_sd = paste0(mean_seizures_per_day, " (", sd_seizures_per_day, ")"),
    median_Q1_Q3 = paste0(median_seizures_per_day, " [", seizures_per_day_p25, "-", seizures_per_day_p75, "]")
  ) %>%
  
  dplyr::select(annual_age_groups, mean_sd, median_Q1_Q3, patients) %>%
    rename( 
    `Mean (sd)` = mean_sd,
    `Median [Q1-Q3]` = median_Q1_Q3,
    `Patients (N)` = patients
  ) %>%
  
  mutate(across(everything(), as.character)) %>%
  
  pivot_longer(cols = -annual_age_groups, names_to = "Metric", values_to = "Value") %>%
  
  pivot_wider(names_from = annual_age_groups, values_from = Value)

# seizure summary table
kable(seizure_summary_t, format = "html", caption = "Table 2. Summary of seizures per day during follow-up by age groups") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  footnote(
    general = "Age groups defined by patient's age at seizure event.",
    general_title = "**",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "For metric calculation, it was obtained first the mean/median for each patient and then, the mean/median across all patients in each group of age.",
    general_title = "*",
    footnote_as_chunk = TRUE)

seizure_summary_t %>%
  write_csv("~/RWE-rare_disease/Age groups analyses/output/tables/seizure_summary_table.csv")
```

```{r Mean of seizures per day chart, echo=TRUE, fig.align='center', message=FALSE, warning=FALSE, dpi=300}
seizure_meanplot <- ggplot(seizure_summary_age_group, aes(x = annual_age_groups, y = mean_seizures_per_day, fill = case_when(
      annual_age_groups %in% c("[0,1)", "[1,2)") ~ "(0-2]", TRUE ~ "(2-12]"
    ))) +
  
  geom_col(position = "dodge", width = 0.6) + 
  
  # geom_text(aes(label = paste0("N = ", patients)), vjust = -0.5, size = 2.8) +  # Show patient count on bars
  
  labs(
    x = "Age Group (in years)",
    y = "Mean of seizures per day",
    fill = "Age group"
  ) +

  scale_fill_manual(values = c(
    "(0-2]" = "grey", 
    "(2-12]" = "black")) + 

  theme_minimal() +
  
  scale_y_continuous(breaks = seq(0, 45, 5), limits = c(0, 45)) +
  
  theme(
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

seizure_meanplot

# Save the plot as a PNG file
ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/Seizure_mean_plot.png",
       seizure_meanplot, units = "mm", width = 200, height = 150 , dpi = 300)
```

```{r median_seizure_per_day_chart, echo=TRUE, fig.align='center', message=FALSE, warning=FALSE, dpi=300}
# Seizure median plot
seizure_medianplot <- ggplot(seizure_summary_age_group, aes(x = annual_age_groups, y = median_seizures_per_day, fill = case_when(
      annual_age_groups %in% c("[0,1)", "[1,2)") ~ "(0-2]",
      TRUE ~ "(2-12]"
    ))) +
  
  geom_col(position = "dodge", width = 0.6) + 
  
  # geom_text(aes(label = paste0("N = ", patients)), vjust = -0.5, size = 2.8) +  # Show patient count on bars
  
  labs(
    x = "Age Group (in years)",
    y = "Median of seizures per day",
    fill = "Age group"
  ) +

  scale_fill_manual(values = c(
    "(0-2]" = "grey", 
    "(2-12]" = "black")) + 

  theme_minimal() +
  
  scale_y_continuous(breaks = seq(0, 13, 1), limits = c(0, 13.5)) +
  
  theme(
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

seizure_medianplot

# save as png graph
ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/Seizure_median_plot.png", 
       seizure_medianplot, units = "mm", width = 200, height = 150 , dpi = 300)
```

#### Hospital admissions

```{r admission derivation, echo=TRUE, message=FALSE, warning=FALSE}
## Hospitalizations ---------------------------------------
admissions_subgrouping <- rare_disease_hospital_admission %>%
  
  # Select relevant columns
  dplyr::select(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    admission_diagnosis
  ) %>%
  
  # Filter out birth-related and planned admissions
  filter(!str_detect(admission_diagnosis, "birth|Birth|planned|Planned")) %>%
  
  # Derivation of admission day during follow-up
  mutate(
    
    # record of hospital admission
    hosp_admission = "Yes",
    
    # Derivation of follow-up duration
    length_of_stay_days = (admission_age_days_lastDate - admission_age_days_firstDate) + 1,

    # Derivation of age at admission
    age_admission = admission_age_days_firstDate / 365.25,

    # Derivation of age at seizure in suggested sub grouping
    annual_age_groups = cut(
      age_admission,
      breaks = c(0:12, Inf),
      right = F,
      include.lowest = T),
    
    # Determine type of admission diagnosis based on seizures/epilepsy
    admission_diagnosis_rec = factor(
      ifelse(
        str_detect(
          tolower(admission_diagnosis), 
          paste(tolower(SEIZURE.EVENT), collapse = "|")
        ),
        "seizure_event",
        "non_seizure_event"
      ),
      levels = c("non_seizure_event", "seizure_event"),
      labels = c("Non-seizure-related", "Seizure-related")
    )
  ) %>%
  
  # Filter out records outside the age range
  filter(age_admission < 12) %>%
  
  # remove duplicated
  distinct(patient_uuid,
           admission_age_days_firstDate,
           admission_age_days_lastDate, 
           admission_diagnosis,
           .keep_all = TRUE) %>%
  
  # Sort by patient and the first day of hospitalization
  arrange(patient_uuid, admission_age_days_lastDate)


# adding rows with patients under risk but without admissions at the specific age group
admissions_subgrouping_complete <- follow_up_patients_uuid %>%
  
  left_join(admissions_subgrouping, by = c("patient_uuid", "annual_age_groups")) %>%
  
  mutate(
    hosp_admission = replace_na(hosp_admission, "No")
  ) %>%
  
  arrange(patient_uuid, annual_age_groups)
```

##### Total hospital admissions

```{r summary of admissions events by age group, echo=FALSE, message=FALSE, warning=FALSE}
# summary of total admissions events by age group
admissions_summary_per_patient <- admissions_subgrouping_complete %>%
  
  group_by(patient_uuid, annual_age_groups) %>%
  
  # count of hospitalizations per patient for each year of life
  summarise(
    admission_count_per_patient = sum(hosp_admission == "Yes"),
    .groups = 'drop'
  )

admissions_summary_age_group <- admissions_summary_per_patient %>%
  
  group_by(annual_age_groups) %>%
  
  # summary of hospitalizations across patients by age group
  summarise(
    # Total number of admissions events
    admission_count = round(sum(admission_count_per_patient), 1),
    # Mean admissions per patient
    mean_admission_per_patient = round(mean(admission_count_per_patient), 1),
    # Standard deviation of admissions per patient
    sd_admission_per_patient = round(sd(admission_count_per_patient), 1),
    # Median admissions per patient
    median_admission_per_patient = round(median(admission_count_per_patient), 1),
    # 25th percentile (Q1)
    p25_admission_per_patient = round(quantile(admission_count_per_patient, 0.25), 1),
    # 75th percentile (Q3)
    p75_admission_per_patient = round(quantile(admission_count_per_patient, 0.75), 1),
    # Number of patients under follow-up
    patients = n_distinct(patient_uuid),
    .groups = 'drop'
  ) %>%
  
  mutate(
    # Percentage of hospital admissions from total admissions
    admission_percentage = round((admission_count / sum(admission_count)) * 100, 1),
  ) %>%

  arrange(annual_age_groups)
```

```{r violin plot of hospital admissions, fig.align='center', message=FALSE, warning=FALSE, dpi=300, include=FALSE}
# Hospital admissions plot
admission_violin_plot <- ggplot(admissions_summary_per_patient, aes(x = annual_age_groups, y = admission_count_per_patient)) +
  
  geom_violin(trim = TRUE, scale = "count", width = 0.9) +
  
  geom_boxplot(width = 0.4, fill = "white") +
  
  labs(
    x = "Age Group (in years)",
    y = "Distribution of the count of hospital admissions per patient"
  ) +

  theme_minimal() +
  
  scale_y_continuous(breaks = seq(0, 15, 1), limits = c(0, 15)) +
  
  theme(
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

admission_violin_plot
```

```{r summary table for hospital admissions, echo=FALSE, message=FALSE, warning=FALSE}
# transpose table
admissions_summary_t <- admissions_summary_age_group %>%
  
  mutate(
    `n (%)` = paste0(admission_count, " (", admission_percentage, ")"),
    `Mean (sd)` = paste0(mean_admission_per_patient, " (", sd_admission_per_patient, ")"),
    `Median [Q1-Q3]` = paste0(median_admission_per_patient, " [", p25_admission_per_patient, "-", p75_admission_per_patient, "]")
  ) %>%
  
  dplyr::select(annual_age_groups,
                `n (%)`,
                `Mean (sd)`, 
                `Median [Q1-Q3]`,
                `Patients (N)` = patients) %>%
  
  mutate(across(everything(), as.character)) %>%
  
  pivot_longer(cols = -annual_age_groups, names_to = "Metric", values_to = "Value") %>%
  
  pivot_wider(names_from = annual_age_groups, values_from = Value)

# summary table
kable(admissions_summary_t, format = "html", caption = "Table 3. Summary of hospital admissions events during follow-up by age groups") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  footnote(
    general = "Age groups defined by patient's age at admission event.",
    general_title = "**",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "Birth and planned admissions are excluded.",
    general_title = "*",
    footnote_as_chunk = TRUE)

admissions_summary_t %>%
  write_csv("~/RWE-rare_disease/Age groups analyses/output/tables/admissions_summary_table.csv")
```

```{r admissions per patient plot, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE,  dpi = 300}
# admission plot
admission_plot <- ggplot(admissions_summary_age_group, aes(x = annual_age_groups, y =  median_admission_per_patient, fill = case_when(
      annual_age_groups %in% 
        c("[0,1)", "[1,2)") ~ "(0-2]",
      TRUE ~ "(2-12]" ))) +
  
  geom_col(position = "dodge", width = 0.6) + 
  
  #geom_text(aes(label = paste0("N = ", patients)), vjust = -2, size = 2.2) +
  
  labs(
    x = "Age Group (in years)",
    y = "Median number of admissions per patient",
    fill = "Age group") +
  
  scale_fill_manual(values = c(
    "(0-2]" = "grey", 
    "(2-12]" = "black")) + 
  
  theme_minimal() +
  
  scale_y_continuous(breaks = seq(0, 4.5, 1), limits = c(0, 4.5)) +
  
  theme(
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

admission_plot

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/Admissions_per_patient_plot.png", admission_plot, units = "mm", width = 200, height = 150 , dpi = 300)
```

```{r admissions per patient plot by admission diagnosis, echo=FALSE, message=FALSE, warning=FALSE}
# Summary of admissions by diagnosis
admissions_summary_diagnosis <- admissions_subgrouping %>%
  
  filter(!is.na(admission_diagnosis_rec)) %>%
  
  group_by(annual_age_groups, patient_uuid, admission_diagnosis_rec) %>%
  
  summarise(
    admission_count_per_patient = sum(hosp_admission == "Yes", na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  
  group_by(annual_age_groups, admission_diagnosis_rec) %>%
  
  summarise(
    # Total admissions for each diagnosis and age group
    admission_count = round(sum(admission_count_per_patient, na.rm = TRUE), 1),
    # Median of admissions per patient
    median_admission_per_patient = round(median(admission_count_per_patient, na.rm = TRUE), 1),
    # Mean of admissions per patient
    mean_admission_per_patient = round(mean(admission_count_per_patient, na.rm = TRUE), 1),
    # Standard deviation
    sd_admission_per_patient = round(sd(admission_count_per_patient, na.rm = TRUE), 1),
    # 25th percentile (Q1)
    p25_admission_per_patient = round(quantile(admission_count_per_patient, 0.25, na.rm = TRUE), 1),
    # 75th percentile (Q3)
    p75_admission_per_patient = round(quantile(admission_count_per_patient, 0.75, na.rm = TRUE), 1),
    patients = n_distinct(patient_uuid),
    .groups = 'drop'
  ) %>%
  
  arrange(annual_age_groups)
```

```{r admissions by diagnosis, fig.align='center', echo=TRUE, message=FALSE, warning=FALSE, dpi= 300}
admission_plot_diag <- ggplot() +
  
  # Bar plot for "Seizure-related" and "Non-seizure-related" admissions
  geom_col(
    data = admissions_summary_diagnosis, 
    aes(x = annual_age_groups, y = mean_admission_per_patient, fill = admission_diagnosis_rec), 
    position = "dodge", width = 0.6
  ) +
  
  # Customize labels
  labs(
    x = "Age Group (in years)",
    y = "Mean number of admissions per patient",
    fill = "Admission Diagnosis*"
  ) +
  
  # Custom colors
  scale_fill_manual(
    values = c(
      "Seizure-related" = "blue",
      "Non-seizure-related" = "red"
    )
  ) +
  
  # Minimal theme
  theme_minimal() +
  
  # Adjust y-axis scale
  scale_y_continuous(breaks = seq(0, 4, 1), limits = c(0, 4.5)) +
  
  # Customize text size and legend
  theme(
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

admission_plot_diag


ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/Admissions_diag_per_patient_plot.png", admission_plot_diag, units = "mm", width = 200, height = 150 , dpi = 300)
```

##### Length of Stay (LOS)

```{r summary of length of stay by age group, echo=TRUE, message=FALSE, warning=FALSE}
los_summary <- admissions_subgrouping_complete %>%
  
  group_by(patient_uuid, annual_age_groups) %>%
  
  # intra-patient mean and median LOS for each group of age
  summarise(
    los_mean_patient = mean(length_of_stay_days, na.rm = TRUE),
    los_median_patient =  median(length_of_stay_days, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  
  group_by(annual_age_groups) %>%
  
  # mean, median and dispersion (deviation and range-based measures) metrics for each group of age
  summarise(
    los_mean = round(mean(los_mean_patient, na.rm = TRUE), 1),
    los_sd = round(sd(los_mean_patient, na.rm = TRUE), 1),
    los_median = round(median(los_median_patient, na.rm = TRUE), 1),
    los_p25 = round(quantile(los_median_patient, 0.25, na.rm = TRUE), 1),
    los_p75 = round(quantile(los_median_patient, 0.75, na.rm = TRUE), 1),
    patients = n_distinct(patient_uuid),
    .groups = "drop"
  ) %>%
  
  arrange(annual_age_groups)
```

```{r summary table of LOS by age group, echo=TRUE, message=FALSE, warning=FALSE}
# transpose table
los_summary_t <- los_summary %>%
  
  mutate(
    `Mean (sd)` = paste0(los_mean, 
                         " (", 
                         los_sd,
                         ")"),
    `Median [Q1-Q3]` = paste0(los_median,
                              " [",
                              los_p25,
                              "-", 
                              los_p75,
                              "]")
  ) %>%
  
  dplyr::select(
    annual_age_groups,
    `Mean (sd)`,
    `Median [Q1-Q3]`, 
    `Patients (N)` = patients) %>%
  
  mutate(across(everything(), as.character)) %>%
  
  pivot_longer(cols = -annual_age_groups, names_to = "Metric", values_to = "Value") %>%
  
  pivot_wider(names_from = annual_age_groups, values_from = Value)

# summary table
kable(los_summary_t, format = "html", caption = "Table 5. Summary of Length of Stay (LOS) during follow-up by age groups.") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  footnote(
    general = "Age groups defined by patient's age at hospital admission",
    general_title = "*",
    footnote_as_chunk = TRUE)%>%
  footnote(
    general = "Birth and planned admissions are excluded.",
    general_title = "**",
    footnote_as_chunk = TRUE)

los_summary_t %>%
  write_csv("~/RWE-rare_disease/Age groups analyses/output/tables/los_summary_table.csv")
```

```{r LOS plot during follow-up, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, dpi= 300}
los_median_plot <- ggplot(los_summary, aes(x = annual_age_groups, y =  los_median, fill = case_when(
      annual_age_groups %in% 
        c("[0,1)", "[1,2)") ~ "(0-2]",
      TRUE ~ "(2-12]" ))) +
  
  geom_col(position = "dodge", width = 0.6) + 
  
 # geom_text(aes(label = paste0("N = ", patients)), vjust = -3.5, size = 2.2) +
  
  labs(
    x = "Age Group (in years)",
    y = "Length of Stay in days (Median)",
    fill = "Age group"
      ) +
  
  scale_fill_manual(values = c(
    "(0-2]" = "grey", 
    "(2-12]" = "black")) + 
  
  theme_minimal() +
  
  scale_y_continuous(breaks = seq(0, 7, 1), limits = c(0, 7)) +
  
  theme(
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

los_median_plot

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/LOS_median_plot.png", 
       los_median_plot, units = "mm", width = 200, height = 150 , dpi = 300)
```

```{r los by admission type, echo=TRUE, fig.align='center', message=FALSE, warning=FALSE, dpi= 300}
los_summary_diagnosis_type <- admissions_subgrouping_complete %>%
  
  filter(!is.na(admission_diagnosis_rec)) %>%

  group_by(patient_uuid, annual_age_groups, admission_diagnosis_rec) %>%
  
  # intra-patient mean and median LOS for each group of age
  summarise(los_mean_patient = mean(length_of_stay_days, na.rm = TRUE),
            los_median_patient = median(length_of_stay_days, na.rm = TRUE),
              .groups = "drop") %>%
  
  ungroup() %>%
  
  group_by(annual_age_groups, admission_diagnosis_rec) %>%  
  
  # mean, median and dispersion (deviation and range-based measures) metrics for each group of age
  summarise(los_mean = round(mean(los_mean_patient, na.rm = TRUE), 1),
            los_sd = round(sd(los_mean_patient, na.rm = TRUE), 1),
            los_median = round(median(los_median_patient, na.rm = TRUE), 1),
            los_p25 = round(quantile(los_median_patient, 0.25, na.rm = TRUE), 1),
            los_p75 = round(quantile(los_median_patient, 0.75, na.rm = TRUE), 1),
            .groups = "drop") %>%
  
  # join with max number of unique patients under follow up
  left_join(dplyr::select(
    summary_count_uuid, annual_age_groups, patients), by = "annual_age_groups") 
```

```{r los plot by admission type, echo=TRUE, fig.align='center', dpi= 300}
los_plot_diag <- ggplot() +
  
  # Bar plot for "Seizure-related" and "Non-seizure-related" admissions
  geom_col(
    data = los_summary_diagnosis_type, 
    aes(x = annual_age_groups, y = los_median, fill = admission_diagnosis_rec), 
    position = "dodge", width = 0.6
  ) +
  
  # Customize the labels and appearance
  labs(
    x = "Age Group (in years)",
    y = "Median of length of stay per patient by type of admission diagnosis",
    fill = "Admission diagnosis*"
  ) +
  
  scale_fill_manual(
    values = c(
      "Seizure-related" = "blue",
      "Non-seizure-related" = "red"
    )
  ) +
  
  # Minimal theme
  theme_minimal() +
  
  # Adjust y-axis scale
  scale_y_continuous(breaks = seq(0, 6, 1), limits = c(0, 6)) +
  
  # Customize text size and legend
  theme(
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

los_plot_diag

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/LOS_median_plot.by.diagnosis.png", 
       los_plot_diag, units = "mm", width = 200, height = 150 , dpi = 300)
```

##### 30-day Unplanned Hospital Readmissions (30-day UHRs)

```{r 30-day UHRs derivation, echo=TRUE, message=FALSE, warning=FALSE}
# derivation of 30 day UHRs
find_readmissions <- function(
    data,              # Input dataset containing admission and discharge records
    grouping,          # Grouping variable to identify unique patients or cohorts
    start_date,        # Column name representing the admission start date
    end_date,          # Column name representing the discharge date
    window_days = 30){ # Time window in days to define readmission

  comparison <- FALSE  # Flag to track whether the data has stabilized after iterations

  while(!comparison){  # Iterate until no new groups are formed
    if(!any(grepl('adm_group', names(data)))){  
      data <- data %>%
        mutate(adm_groups = 0)  # Initialize admission group column if it does not exist
    }

    prev_data <- data  # Store a copy of the previous iteration's data for comparison

    data <- data %>%
      arrange(!!sym(start_date)) %>%  # Sort the dataset by the admission start date
      
      group_by_at(c('adm_groups', grouping)) %>%  # Group by current admission groups and provided grouping variable
      
      mutate(within_window = !!sym(start_date) - first(!!sym(end_date)) <= window_days, 
             # Compute whether the admission falls within the 30-day readmission window
             adm_groups = cur_group_id() + cumsum(within_window != lag(within_window, default = first(within_window)))) %>%
             # Assign new admission groups based on readmission criteria
      
      mutate(readmission = case_when(
        row_number() == 1 ~ 'index',  # Mark the first admission in a group as the index admission
        TRUE ~ glue::glue('30_days_uhr_{row_number()-1}')  # Label readmissions sequentially
        )) %>%
      
      ungroup()  # Remove grouping to allow further processing

    comparison <- identical(prev_data, data)  # Check if any changes were made; if not, stop iterating
  }

  return(data %>%
           dplyr::select(-adm_groups, -within_window))  # Remove intermediate columns before returning the final dataset
}

# Run prior function to find 30-day UHRs
readmissions_subgrouping <- find_readmissions(
  data = admissions_subgrouping,
  grouping = 'patient_uuid',
  start_date = 'admission_age_days_firstDate',
  end_date = 'admission_age_days_lastDate'
  )


# adding variable to validate only the first 30-day readmissions if many occurs in the same time 30-day windows after last index discharge day
readmissions_subgrouping <- readmissions_subgrouping %>%
  
  arrange(patient_uuid, admission_age_days_firstDate, admission_age_days_lastDate) %>%
  
  group_by(patient_uuid) %>%
  
  mutate(
    # Dynamically update last_index_discharge by keeping the most recent "index" discharge
    last_index_discharge = if_else(readmission == "index", admission_age_days_lastDate, NA_real_),
    
    # Fill forward the last_index_discharge for the patient
    last_index_discharge = zoo::na.locf(last_index_discharge, na.rm = FALSE),
    
    # Parse the numeric part of the readmission string for comparison
    readmission_number = as.integer(str_extract(readmission, "\\d+")),
    
    # Calculate the time difference between last admission with the last index discharge
    time_diff = admission_age_days_firstDate - last_index_discharge,
    
    # Determine if the current row is a valid readmission
    valid_readmission = case_when(
      # Mark index as index admission
      readmission == "index" ~ 'Index',
      
      # Mark the first valid as always 30-day readmission
      readmission == "30_days_uhr_1" ~ 'Yes',
      
      # check the distance between 2nd and further 30 UHRs in the sequence with the last index.
      readmission_number == lag(readmission_number, default = NA) + 1 & time_diff > 30 ~ 'Yes',
      
      # All other cases are not valid readmissions
      TRUE ~ 'No'
    )
  ) %>%
  
  ungroup() %>%
  
  dplyr::select(-readmission_number, -time_diff, -last_index_discharge)
```

```{r obtaining the number of patient under risk of 30-day UHRs (with at least one admission event at follow-up), message=FALSE, warning=FALSE, include=FALSE}
# Merge follow-up and hospital admission data to calculate patient under risk
hosp_admission_follow_up_risk <- sex_follow_up %>%
  left_join(rare_disease_hospital_admission, by = "patient_uuid") %>%
   # Keep only those with at least one admission
  filter(!is.na(admission_age_days_firstDate))

# Extract patients under risk of 30-day admission
readmission_follow_patients_uuid <- hosp_admission_follow_up_risk %>%
  
  group_by(patient_uuid, sex) %>%
  
  summarise(first_admission_age_days = min(admission_age_days_firstDate, na.rm = TRUE), 
            last_record_age_days = max(last_record_age_days, na.rm = TRUE)) %>%
  
  # Define age groups from first admission onwards
  reframe(age_range = seq(floor(first_admission_age_days/365.25),
                          floor(last_record_age_days/365.25))) %>%
  
  mutate(annual_age_groups = cut(
    age_range,
    breaks = c(0:12, Inf),
    right = F,
    include.lowest = T)) %>%
  
  filter(age_range < 12) %>%
  
  dplyr::select(-age_range)

# adding rows with patients under risk but without readmissions at the specific age group
readmissions_subgrouping <- readmission_follow_patients_uuid %>%
  
  left_join(readmissions_subgrouping, by = c("patient_uuid", "annual_age_groups")) %>%
  
  mutate(
    valid_readmission = replace_na(valid_readmission, "No")
  ) %>%
  
  arrange(patient_uuid, annual_age_groups)
```

```{r summary of 30-day readmissions events by age group, echo=FALSE, message=FALSE, warning=FALSE}
readmissions_summary <- readmissions_subgrouping %>%

  group_by(patient_uuid, annual_age_groups) %>%  
  
  # count of readmissions per patient for each year of life
  summarise(
    readmission_count_per_patient = sum(valid_readmission == "Yes", na.rm = TRUE),
    .groups = 'drop') %>%
  
  group_by(annual_age_groups) %>%
  
  # summary of readmissions across patients by age group
  summarise(
    # Total number of readmissions
    readmission_count = round(sum(readmission_count_per_patient), 1),  
    # Mean of readmissions per patient
    mean_readmission_per_patient = round(mean(readmission_count_per_patient), 1),  
    # Standard deviation of readmissions per patient
    sd_readmission_per_patient = round(sd(readmission_count_per_patient), 1), 
    # Median of readmissions per patient
    median_readmission_per_patient = round(median(readmission_count_per_patient), 1), 
    # 25th percentile (Q1)
    p25_readmission_per_patient = round(quantile(readmission_count_per_patient, 0.25), 1), 
    # 75th percentile (Q3)
    p75_readmission_per_patient = round(quantile(readmission_count_per_patient, 0.75), 1), 
    patients = n_distinct(patient_uuid),
    .groups = 'drop') %>%
  
  mutate(
    # Percentage of readmissions
    readmission_percentage = round((readmission_count / sum(readmission_count)) * 100, 1),
  ) %>%

  arrange(annual_age_groups)
```

```{r summary table of 30-day UHRs, echo=FALSE, message=FALSE, warning=FALSE}
# transpose data
readmissions_summary_t <- readmissions_summary %>%
  
  mutate(
    `n (%)` = paste0(readmission_count, " (", readmission_percentage, ")"),
    `Mean (sd)` = paste0(mean_readmission_per_patient, " (", sd_readmission_per_patient, ")"),
    `Median [Q1-Q3]` = paste0(median_readmission_per_patient, " [", p25_readmission_per_patient, "-", p75_readmission_per_patient, "]")
  ) %>%
  
  dplyr::select(
    annual_age_groups,
    `n (%)`,
    `Mean (sd)` ,
    `Median [Q1-Q3]`,
    `Patients (N)` = patients) %>%
  
  mutate(across(everything(), as.character)) %>%
  
  pivot_longer(cols = -annual_age_groups, names_to = "Metric", values_to = "Value") %>%
  
  pivot_wider(names_from = annual_age_groups, values_from = Value)

# summary table
kable(readmissions_summary_t, format = "html", caption = "Table 4. Summary of 30-day unplanned hospital readmissions (30-day UHRs) during follow-up by age groups") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  footnote(
    general = "Patients at risk of 30-day UHRs are those with at least one prior admission since start of follow-up.",
    general_title = "***",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "Age groups defined by patient's age at admission event.",
    general_title = "**",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "A 30-day unplanned hospital readmission is defined as a subsequent unplanned admission within 30 days following an index discharge date. If multiple 30-day UHRs occur within the same 30-day window after last index discharge, only the first event of 30-day UHRs is counted.",
    general_title = "*",
    footnote_as_chunk = TRUE)

readmissions_summary_t %>%
  write_csv("~/RWE-rare_disease/Age groups analyses/output/tables/readmissions_summary_table.csv")
```

```{r 30-day UHRs per patient plot, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, dpi = 300}
readmissions_plot <- ggplot(readmissions_summary, aes(x = annual_age_groups, y =  mean_readmission_per_patient, fill = case_when(annual_age_groups %in% 
        c("[0,1)", "[1,2)") ~ "(0-2]",
      TRUE ~ "(2-12]" ))) +
  
  geom_col(position = "dodge", width = 0.6) + 
  
  # geom_text(aes(label = paste0("N = ", patients)), vjust = -3.5, size = 2.2) +
  
  labs(
    x = "Age Group (in years)",
    y = "Mean number of 30-day UHRs per patient",
    fill = "Age group") +
  
  scale_fill_manual(values = c( "(2-12]" = "black", 
                                  "(0-2]" = "grey")) + 
  
  theme_minimal() +
  
  scale_y_continuous(breaks = seq(0, 2, 1), limits = c(0, 2)) +
  theme(
    panel.background = element_rect(fill = "white", color = NA), 
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11, face = "bold", margin = margin(r = -18)),
    axis.title.x = element_text(size = 11),
    axis.title.y = element_blank(),
    legend.title = element_text(size = 11),
    legend.position = "bottom",
    legend.text = element_text(size = 11),
    
    # Adjust margins and panel spacing
    plot.margin = margin(5, 5, 5, 5),  # Reduce whitespace around the plot
    panel.spacing = unit(0, "lines"),  # Reduce spacing between plot elements
    
    # Fix aspect ratio to reduce vertical stretch
    aspect.ratio = 0.3  # Lower values make it more compact
  )

readmissions_plot

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/30-day_UHRs_per_patient_plot.png", 
       readmissions_plot, units = "mm", width = 200, height = 150 , dpi = 300)
```

```{r illustrative plot of 30-day UHRs, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, dpi = 300}
#### Explanatory figure ------------------------------------------------
readmission_example_plot <- readmissions_subgrouping %>%
  
  dplyr::select(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    length_of_stay_days,
    readmission,
    valid_readmission
  ) %>%
  
  distinct() %>%
  
  ungroup() %>%
  
  # patient with 3 subsequent admissions in less than 30 day after index admission
  filter(patient_uuid == "b6808b7f-9119-475a-bd19-e3a7dc723fed") %>%
  
  left_join(rare_disease_followup_durations_days, by = "patient_uuid") %>%
  
  mutate(
    patient_label = "Patient example",
    followup_duration = last_record_age_days - first_record_age_days,
  ) %>%
  mutate(
    # Dynamically update last_index_discharge by keeping the most recent "index" discharge
    last_index_discharge = if_else(readmission == "index", admission_age_days_lastDate, NA_real_),
    
    # Fill forward the last_index_discharge for the patient
    last_index_discharge = zoo::na.locf(last_index_discharge, na.rm = FALSE),
    
    # Calculate the time difference between last admission with the last index discharge
    time_diff = admission_age_days_firstDate - last_index_discharge
    ) %>%
  
  mutate(
    readmission_label = case_when(
      valid_readmission == "Index" ~ "Index discharge day",
      valid_readmission == "Yes" ~ "30-day UHRs",
      valid_readmission == "No" ~ "not counted readmission"
    )
  ) %>%
  ggplot(.) +
  scale_alpha_manual(name = "", values = 1) +
  geom_linerange(
    aes(y = patient_label, xmin = 1, xmax = 80),
    linetype = "dashed",
    size = .2,
    
  ) +
  
  geom_segment(aes(x = 39, xend = 69, y = 1.05, yend = 1.05), 
               color = "black", size = 0.2) +
  
  # Add left arrow at x = 39
  geom_segment(aes(x = 39, xend = 36, y = 1.05, yend = 1.05), 
               color = "black", size = 0.2, 
               arrow = arrow(type = "closed", length = unit(0.2, "cm"))) +
  
  # Add right arrow at x = 69
  geom_segment(aes(x = 66, xend = 69, y = 1.05, yend = 1.05), 
               color = "black", size = 0.2, 
               arrow = arrow(type = "closed", length = unit(0.2, "cm"))) +
  
  # Add label above the line
  annotate("text", x = 52, y = 1.1, label = "30-day window after index discharge (day 39)",
           size = 3) +
  
  geom_point(
    aes(
      x = admission_age_days_firstDate,
      y = patient_label,
      colour = factor(readmission_label)
    ),
    shape = 18,
    size = 5,
    show.legend = TRUE
  ) +
  geom_text(
    aes(
      x = admission_age_days_firstDate,
      y = patient_label,
      label = case_when(
        readmission_label == "30-day UHRs" ~ as.character(paste0("day ", abs(admission_age_days_firstDate))),
        readmission_label == "not counted readmission" ~ as.character(paste0("day ", abs(admission_age_days_firstDate))
      ))
    ),
    size = 3,
    hjust = 0.5,
    vjust = -1,
    nudge_y = - 0.08) +
  scale_colour_manual(
    values = c(
      "30-day UHRs" = "blue",
      "not counted readmission" = "grey",
      "Index discharge day" = "red"
    ),
    labels = c(  
      "Index discharge day" = "Index admission",
      "30-day UHRs" = "30-day UHRs",
      "not counted readmission" = "Not counted readmission"
    ),
    name = "Legend"
  ) +
  scale_x_continuous(
    breaks = seq(0, 80, 5),
    limits = c(0, 80)
  ) +
  labs(x = "Follow-up time (days)") +
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA), 
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7, face = "bold", margin = margin(r = -18)),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_blank(),
    legend.title = element_text(size = 9),
    legend.position = "bottom",
    legend.text = element_text(size = 9)
  )


readmission_example_plot

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/readmissions_example.png", 
       readmission_example_plot, units = "mm", width = 200, height = 150 , dpi = 300)
```

#### Respiratory support

```{r ther. procedures derivation updated, echo=TRUE, message=FALSE, warning=FALSE}
rare_disease_therapeutic_subgr_all <- rare_disease_therapeutic_procedures %>%
  
  # Filter rows with terms related to respiratory support
  filter(str_detect(procedure, regex("ventil|oxygen|tracheostomy|tracheotomy", ignore_case = TRUE))) %>%
  
  # Select relevant columns with dates of procedures
  ## first and last date columns were not used in ther.procedures due to procedures with first/last date are not related to respiratory support (i.e. ketogenic diet, Nissen fundoplication, Insertion of gastrostomy tube and Phototherapy)
  dplyr::select(patient_uuid, procedure, procedure_age_days_date) %>%
  
  # remove duplicates
  distinct() %>%
  
  # Convert procedure to factor for consistency
  mutate(procedure = factor(procedure)) %>%
  
  # Group by patient and admission periods
  group_by(patient_uuid) %>%
  
  # Derivation of therapeutic procedure follow-up day and get some missing dates from first date
  mutate(
    
    procedure_age = procedure_age_days_date / 365.25,
    
    # Derivation of age at seizure in 1-year intervals
    annual_age_groups = cut(
      procedure_age,
      breaks = c(0:12, Inf),
      right = F,
      include.lowest = T)
  ) %>%
  
  # Filter out ages greater than 12 years
  filter(procedure_age < 12)
```

```{r clinical diagnosis resp. derivation, echo=TRUE, message=FALSE, warning=FALSE}
# obtaining events with respiratory support term
rare_disease_clinical_diagnosis_subgr_all <- rare_disease_clinical_diagnosis %>%
    
  # Convert procedure to factor for consistency and rename as procedure for binding rows
  mutate(procedure = factor(clinical_diagnosis)) %>%
  
  # Filter rows with terms related to respiratory support and exclude hypoventilation
  filter(
    str_detect(procedure, regex("ventil|oxygen|tracheostomy|tracheotomy", ignore_case = TRUE)) &
    !str_detect(procedure, regex("hypoventilation", ignore_case = TRUE))
  ) %>%

  # Select relevant columns
  dplyr::select(patient_uuid, procedure, clinical_diagnosis_age_days_firstDate, clinical_diagnosis_age_days_lastDate) %>%
  
  # remove duplicates
  distinct(.keep_all = TRUE)


# derivation of group of age for each diagnosis
rare_disease_clinical_diagnosis_subgr_all <- rare_disease_clinical_diagnosis_subgr_all %>%
  
  group_by(patient_uuid, procedure) %>% 
  
  # Arrange data for detecting gaps
  arrange(
    clinical_diagnosis_age_days_firstDate,
    desc(clinical_diagnosis_age_days_lastDate)
  ) %>%
  
  # Identify episodes of respiratory support
  mutate(
    ## find the largest end date from current row to row 1 (for each pat/medic combination)
    max_end = purrr::accumulate(clinical_diagnosis_age_days_lastDate, max),
    
    ## computes the difference between current start and the end of the previous row
    date_diff = clinical_diagnosis_age_days_firstDate - lag(max_end, default = T),
    
    ## if date_diff is >=2 (episod starts more than 1 day after the end of the previous), episod is incremented 1
    episod = purrr::accumulate(date_diff >= 2, sum, na.rm = TRUE)
  ) %>%
  
  group_by(patient_uuid, procedure, episod) %>%

  arrange(patient_uuid, procedure, clinical_diagnosis_age_days_firstDate) %>%

  summarize(
    procedure_age_days_firstDate = min(clinical_diagnosis_age_days_firstDate),
    procedure_age_days_lastDate = max(clinical_diagnosis_age_days_lastDate),
    procedure_duration = (procedure_age_days_lastDate - procedure_age_days_firstDate) + 1,
    .groups = "drop"
  ) %>%

  # Expand medication periods to capture all ages during follow-up
  rowwise() %>%
  
  # create list with age sequence per patient increasing by 1 unit to capture an aprox moment of the year window in which the patient was under the medication
  mutate(age_sequence = list(
    seq(
      from = procedure_age_days_firstDate / 365.25,
      to = procedure_age_days_lastDate / 365.25,
      by = 1
    ) %>%
      unique() %>%  # Ensure unique ages in case rounding causes overlaps
      c(procedure_age_days_lastDate / 365.25) # Explicitly include the exact last medication intake
  )) %>%
  
  unnest(cols = c(age_sequence)) %>% # returns list to rows
  
  mutate(
    procedure_age = age_sequence, # exact age for first and last record

    # Derivation of age at seizure in 1-year intervals
    annual_age_groups = cut(
      procedure_age,
      breaks = c(0:12, Inf),
      right = F,
      include.lowest = T)
    
  ) %>%
  
  filter(procedure_age < 12) %>%
  
  arrange(patient_uuid, procedure, episod, procedure_age_days_lastDate) %>%
  
  group_by(patient_uuid, procedure, episod, annual_age_groups) %>%
  
  slice_tail(n = 1) %>%  # Keep the last row (last medication intake) within each sub grouping if duplicates exist
  
  ungroup() %>%
  
  dplyr::select(patient_uuid, annual_age_groups, procedure_age, procedure_age_days_firstDate, procedure_age_days_lastDate, procedure)
```

```{r joining databases. derivation, echo=TRUE, message=FALSE, warning=FALSE}
# Joining clinical diagnosis and therapeutic procedures datasets
respiratory_sup_subgrouping <-  
  bind_rows(
    rare_disease_clinical_diagnosis_subgr_all %>% mutate(source = "clinical_diagnosis"),
    rare_disease_therapeutic_subgr_all %>% mutate(source = "therapeutic")
  )  %>%
  
  # Filter records from therapeutic procedures that do not have the same date for clinical diagnosis
  filter(!(source == "therapeutic" & procedure_age_days_firstDate %in% rare_disease_clinical_diagnosis_subgr_all$procedure_age_days_firstDate)) %>%
  
  mutate(
    resp_support = "Yes"
  )


# add rows with patients under risk but without records
respiratory_sup_subgrouping <- follow_up_patients_uuid %>%
  
  left_join(respiratory_sup_subgrouping, by = c("patient_uuid", "annual_age_groups")) %>%
  
  mutate(
    resp_support = replace_na(resp_support, "No")
  ) %>%
  
  arrange(patient_uuid, annual_age_groups)
```

```{r summary of respiratory support events by age group, echo=TRUE, message=FALSE, warning=FALSE}
# respiratory support events by age group
respiratory_support_summary <- respiratory_sup_subgrouping %>%
  
  group_by(patient_uuid, annual_age_groups) %>%
  
  summarise(
    respiratory_sup_count_per_patient = sum(resp_support == "Yes"),
    .groups = 'drop'
  ) %>%
  
  group_by(annual_age_groups) %>%
  
  summarise(
    # Total number of respiratory support procedures
    respiratory_sup_count = round(sum(respiratory_sup_count_per_patient), 1),
    # Mean of respiratory support procedures per patient
    mean_resp_sup_per_patient = round(mean(respiratory_sup_count_per_patient), 1),
    # Standard deviation of respiratory support procedures per patient
    sd_resp_sup_per_patient = round(sd(respiratory_sup_count_per_patient), 1),
    # Median of respiratory support procedures per patient
    median_resp_sup_per_patient = round(median(respiratory_sup_count_per_patient), 1),
    # 25th percentile (Q1)
    p25_resp_sup_per_patient = round(quantile(respiratory_sup_count_per_patient, 0.25), 1),
    # 75th percentile (Q3)
    p75_resp_sup_per_patient = round(quantile(respiratory_sup_count_per_patient, 0.75), 1),
    # Number of patients under follow up
    patients = n_distinct(patient_uuid),
    .groups = 'drop'
  ) %>%

  mutate(
    # Percentage of total rescue medications
    resp_supp_percentage = round((respiratory_sup_count / sum(respiratory_sup_count)) * 100, 1),
  ) %>%

  arrange(annual_age_groups)
```

```{r summary table of respiratory support events by age group, echo=TRUE, message=FALSE, warning=FALSE}
# transpose table
respiratory_support_summary_t <- respiratory_support_summary %>%
  
  mutate(
    `n (%)` = paste0(respiratory_sup_count,
                     " (",
                     resp_supp_percentage,
                     ")"),
    `Mean (sd)` = paste0(mean_resp_sup_per_patient,
                         " (",
                         sd_resp_sup_per_patient,
                         ")"),
    `Median [Q1-Q3]` = paste0(median_resp_sup_per_patient,
                              " [",
                              p25_resp_sup_per_patient,
                              "-",
                              p75_resp_sup_per_patient,
                              "]")
    ) %>%
  
  dplyr::select(
    annual_age_groups,
    `n (%)`,
    `Mean (sd)`, 
    `Median [Q1-Q3]`, 
    `Patients (N)` = patients
    ) %>%
  
  mutate(across(everything(), as.character)
         ) %>%
  
  pivot_longer(cols = -annual_age_groups, 
               names_to = "Metric", 
               values_to = "Value") %>%
  
  pivot_wider(names_from = annual_age_groups, 
              values_from = Value) 

# summary kable
kable(respiratory_support_summary_t, format = "html", caption = "Table 6. Summary of respiratory support* events during follow-up by age groups") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2")  %>%
  footnote(
    general = "Age groups defined by age at start and end of clinical diagnosis event or age at therapeutic procedure event.",
    general_title = "**",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "Respiratory support records for a specific procedure/clinical diagnosis with continuous dates (< 2 days of gap) were grouped into episodes.",
    general_title = "***",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "Respiratory support procedures are based on the occurrence of 'ventil|oxygen|tracheostomy|tracheotomy' procedures, according to the 'procedure' and 'clinical diagnosis' variables from clinical diagnosis and therapeutic procedures files.",
    general_title = "*",
    footnote_as_chunk = TRUE)

respiratory_support_summary_t %>%
  write_csv("~/RWE-rare_disease/Age groups analyses/output/tables/resp.support_summary_table.csv")
```

```{r respiratory support per patient plot, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, dpi = 300}
resp_supp_plot <- ggplot(respiratory_support_summary, aes(x = annual_age_groups, y =  mean_resp_sup_per_patient, fill = 
                                                            case_when(annual_age_groups %in% c("[0,1)", "[1,2)") ~ "(0-2]",
                                                                      TRUE ~ "(2-12]" ))) +
  
  geom_col(position = "dodge", width = 0.6) +  
  
  #geom_text(aes(label = paste0("N = ", patients)), vjust = -3.5, size = 2.2) +
  
  labs(
    x = "Age Group (in years)",
    y = "Mean number of respiratory support procedures per patient",
    fill = "Age group") +
  
  scale_fill_manual(values = c(
    "(0-2]" = "grey", 
    "(2-12]" = "black")) + 
  
  theme_minimal() +
  
  scale_y_continuous(breaks = seq(0, 2, 1), limits = c(0, 2)) +
  
  theme(
  axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

resp_supp_plot

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/resp_supp_per_patient_plot.png", 
       resp_supp_plot, units = "mm", width = 200, height = 150 , dpi = 300)
```

#### Medication

##### Rescue medication

```{r Medication derivation, echo=TRUE, message=FALSE, warning=FALSE}
rescue_med_subgrouping <- rare_disease_medication %>%
  
  # Select relevant columns
  dplyr::select(
    patient_uuid,
    medication_age_days_firstDate,
    medication_age_days_lastDate,
    medication
    ) %>%

  # Derivation of medication day during follow-up
  mutate(
    
    # Derivation of age at medication start
    age_medication = medication_age_days_firstDate / 365.25, 

    # Derivation of age groups based on age at medication prescription
    annual_age_groups = cut(
      age_medication,
      breaks = c(0:12, Inf),
      right = F,
      include.lowest = T),
    
    # Derivation of rescue medication based on age at first and last day of medication
  rescue_medication = factor(
    ifelse(
      # list obtained from epilepsy foundation
      medication %in% c("Diazepam",
                        "Lorazepam",
                        "Midazolam", 
                        "Clonazepam", 
                        "Clorazepate"
                        ) &
        medication_age_days_firstDate == medication_age_days_lastDate,
      "Yes",
      "No"
    ),
    levels = c("Yes", "No")
  )
  ) %>% 

  # Filter out records outside the age range 
  filter(age_medication < 12) %>%

  # distinct records to remove duplicates
  distinct(
    patient_uuid,
    medication_age_days_firstDate,
    medication_age_days_lastDate,
    medication,
    .keep_all = TRUE) %>%
  
  # keep selected columns in order
  dplyr::select(
    patient_uuid,
    medication,
    rescue_medication,
    age_medication,
    annual_age_groups,
    medication_age_days_firstDate,
    medication_age_days_lastDate
  ) %>%
  
  # arrange by patient uuid
  arrange(patient_uuid, medication_age_days_firstDate)


# add rows with patients under risk but without records
rescue_med_subgrouping <- follow_up_patients_uuid %>%
  
  left_join(rescue_med_subgrouping, by = c("patient_uuid", "annual_age_groups")) %>%
  
  mutate(
    rescue_medication = replace_na(rescue_medication, "No")
  ) %>%
  
  arrange(patient_uuid, annual_age_groups)
```

```{r summary of rescue medication by age group, echo=FALSE, message=FALSE, warning=FALSE}
# rescue medication summary
rescue_med_summary <- rescue_med_subgrouping %>%
  
  group_by(patient_uuid, annual_age_groups) %>%
  
  # summary by patient at age group
  summarise(
    resc_med_count_per_patient = sum(rescue_medication == "Yes"),
    .groups = 'drop'
  ) %>%
  
  group_by(annual_age_groups) %>%
  
  # summary across age group
  summarise(
    # Total number of rescue medications
    resc_med_count = round(sum(resc_med_count_per_patient), 1),
    # Mean of rescue medications per patient
    mean_resc_med_per_patient = round(mean(resc_med_count_per_patient), 1),
    # Standard deviation of rescue medications per patient
    sd_resc_med_per_patient = round(sd(resc_med_count_per_patient), 1),
    # Median of rescue medications per patient
    median_resc_med_per_patient = round(median(resc_med_count_per_patient), 1),
    # 25th percentile (Q1)
    p25_resc_med_per_patient = round(quantile(resc_med_count_per_patient, 0.25), 1),
    # 75th percentile (Q3)
    p75_resc_med_per_patient = round(quantile(resc_med_count_per_patient, 0.75), 1),
    # patient under follow up
    patients = n_distinct(patient_uuid),
    .groups = 'drop'
  ) %>%
  
  mutate(
    # Percentage of total rescue medications
    resc_med_percentage = round((resc_med_count / sum(resc_med_count)) * 100, 1),
  ) %>%

  arrange(annual_age_groups)
```

```{r summary table - rescue medication by age group, echo=FALSE, message=FALSE, warning=FALSE}
# transpose table
rescue_med_summary.t <- rescue_med_summary %>%
  
  mutate( 
    `n (%)` = paste0(resc_med_count, 
                     " (",
                     resc_med_percentage,
                     ")"),
    `Mean (sd)` = paste0(mean_resc_med_per_patient,
                         " (",
                         sd_resc_med_per_patient,
                         ")"),
    `Median [Q1-Q3]` = paste0(median_resc_med_per_patient, 
                              " [", 
                              p25_resc_med_per_patient, 
                              "-",
                              p75_resc_med_per_patient,
                              "]")
  ) %>%
  dplyr::select(
    annual_age_groups,
    `n (%)`, 
    `Mean (sd)`, 
    `Median [Q1-Q3]`,
    `Patients (N)` = patients
    ) %>%
  
  mutate(across(everything(), as.character)
         ) %>%
  
  pivot_longer(cols = -annual_age_groups, names_to = "Metric", values_to = "Value") %>%
  
  pivot_wider(names_from = annual_age_groups, values_from = Value)

# kable summary table
kable(rescue_med_summary.t, format = "html", caption = "Table 7. Summary of total rescue medication prescriptions during follow-up by age groups") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2")  %>%
  footnote(
    general = "Records with missing values in dose and unit are counted.",
    general_title = "***",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "Age groups defined by age at medication (earliest report of medication/dose in record).",
    general_title = "**",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "Rescue medication is defined as one-off use administration of benzodiazepines to treat acute repetitive seizures or clusters. This includes diazepam, lorazepam, midazolam, clorazepate and clonazepam.",
    general_title = "*",
    footnote_as_chunk = TRUE)

rescue_med_summary.t %>%
  write_csv("~/RWE-rare_disease/Age groups analyses/output/tables/rescue_med_summary_table.csv")
```

```{r rescue medication per patient plot, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE,  dpi = 300}
resc_med_plot <- ggplot(rescue_med_summary, aes(x = annual_age_groups, y =  mean_resc_med_per_patient, fill = case_when(
      annual_age_groups %in% c("[0,1)", "[1,2)") ~ "(0-2]", TRUE ~ "(2-12]"
    ))) +
  
  geom_col(position = "dodge", width = 0.6) + 
  
  #geom_text(aes(label = paste0("N = ", patients)), vjust = -3.5, size = 2.2) +
  
  labs(
    x = "Age Group (in years)",
    y = "Mean of rescue medication per patient",
    fill = "Age group")  +
  
  scale_fill_manual(values = c( "(2-12]" = "black", 
                                  "(0-2]" = "grey")) + 
  
  theme_minimal() +
  
  scale_y_continuous(breaks = seq(0, 7, 1), limits = c(0, 7)) +
  
  theme(
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

resc_med_plot

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/resc_med_per_patient_plot.png", 
       resc_med_plot, units = "mm", width = 200, height = 150 , dpi = 300)
```

##### Anti-seizure medication

```{r derivation of antiseizure medication episodes, echo=TRUE, message=FALSE, warning=FALSE}
medication_antiseizure_subgrouping <- rare_disease_medication %>% 
      
  # 35 total medications (A Summary of Antiseizure Medications Available in the United States: 4th Edition) 2024
  filter(medication %in% c("Brivaracetam", 
                           "Carbamazepine", 
                           "Cenobamate", 
                           "Clobazam", 
                           "Clonazepam", 
                           "Diazepam", 
                           "Divalproex", 
                           "Eslicarbazepine acetate",
                           "Ethosuximide", 
                           "Epidiolex", # cannabidiol (CBD)
                           "Everolimus", 
                           "Felbamate", 
                           "Fenfluramine", 
                           "Fosphenytoin", 
                           "Gabapentin", 
                           "Ganaloxone", 
                           "Lacosamide", 
                           "Lamotrigine", 
                           "Levetiracetam", 
                           "Lorazepam", 
                           "Midazolam", 
                           "Oxcarbazepine",
                           "Perampanel", 
                           "Phenobarbital", 
                           "Phenytoin",
                           "Pregabalin", 
                           "Primidone", 
                           "Rufinamide", 
                           "Stiripentol", 
                           "Tiagabine", 
                           "Topiramate", 
                           "Valproic acid", 
                           "Vigabatrin", 
                           "Zonisamide")) %>% 
  
  dplyr::select(
    patient_uuid, 
    medication, 
    medication_age_days_firstDate, 
    medication_age_days_lastDate) %>% 
  
  ## remove duplicate rows (not taking dose into account for this analysis)
  distinct() %>% 
  
  group_by(patient_uuid, medication) %>%
  
  arrange(
    medication_age_days_firstDate,
    desc(medication_age_days_lastDate)
  ) %>%
  
  # Identify episodes (Identify the longest prescription windows and remove the nested ones)
  mutate(
    ## find the largest end date from current row to row 1 (for each pat/medic combination)
    max_end = purrr::accumulate(medication_age_days_lastDate, max),
    
    ## computes the difference between current start and the end of the previous row
    date_diff = medication_age_days_firstDate - lag(max_end, default = T),
    
    ## if date_diff is >=2 (episod starts more than 1 day after the end of the previous), episod is incremented 1
    episod = purrr::accumulate(date_diff >= 2, sum, na.rm = TRUE)
  ) %>%
  
  # Group by patient, medication, and episode to identify distinct prescription windows
  group_by(patient_uuid, medication, episod) %>%
  
  arrange(patient_uuid, medication, medication_age_days_firstDate) %>%
  
  summarise(
    medication_age_days_firstDate = min(medication_age_days_firstDate),
    medication_age_days_lastDate = max(medication_age_days_lastDate),
    .groups = "drop"
  ) %>%
  
  # remove prescriptions initiated after 12 years in days
  filter(medication_age_days_firstDate <= MAX_DAYS_12_YEARS) %>%
  
  # Truncate medication_age_days_lastDate to the last day is higher than 12 years (in days) to avoid overestimation of prescription duration in the lattest group of age (to check)
  mutate(
    medication_age_days_lastDate = if_else(
      medication_age_days_lastDate > MAX_DAYS_12_YEARS,
      MAX_DAYS_12_YEARS,
      medication_age_days_lastDate
    ),
    prescr_duration = medication_age_days_lastDate - medication_age_days_firstDate + 1
  )
```

###### a. Total ASMs prescriptions

```{r total ASM by age group, echo=TRUE, message=FALSE, warning=FALSE}
total_asm_subgr <- medication_antiseizure_subgrouping %>%

  # Expand medication periods to capture all ages during follow-up
  rowwise() %>%

  mutate(
    age_sequence = list(
      seq(
      from = medication_age_days_firstDate / 365.25,
      to = medication_age_days_lastDate / 365.25,
      by = 1 # by 1 year increase
    ) %>%
      
      # Ensure unique ages in case rounding causes overlaps
      unique() %>%  
      
      # finish the sequence with the last date of medication intake
      c(medication_age_days_lastDate / 365.25)
    
  )) %>%

  unnest(cols = c(age_sequence)) %>%

  mutate(
    
    # age at medication prescription
    age_medication = age_sequence,

    # Derivation of in 1-year intervals
    annual_age_groups = cut(
      age_medication,
      breaks = c(0:12, Inf),
      right = F,
      include.lowest = T)
  ) %>%
  
  arrange(patient_uuid, medication, episod, medication_age_days_lastDate) %>%

  # Keep the last row (last medication intake) within each sub grouping if duplicates exist
  group_by(patient_uuid, medication, episod, annual_age_groups) %>%
  
  slice_tail(n = 1) %>%  
  
  ungroup() %>%

  # Filter out records outside the age range
  filter(age_medication < 12) %>%

  mutate(
    antiseizure_med = "Yes"
  ) %>%

  # Keep selected columns for further processing
  dplyr::select(
    patient_uuid,
    medication,
    age_medication,
    annual_age_groups,
    medication_age_days_firstDate,
    medication_age_days_lastDate,
    prescr_duration,
    antiseizure_med
  ) %>%

  # Arrange by patient uuid
  arrange(patient_uuid, medication, medication_age_days_firstDate)


# add rows with patients under risk but without records
total_asm_subgr <- follow_up_patients_uuid %>%
  
  left_join(total_asm_subgr, by = c("patient_uuid", "annual_age_groups")) %>%
  
  mutate(
    antiseizure_med = replace_na(antiseizure_med, "No")
  ) %>%
  
  # Arrange by patient uuid
  arrange(patient_uuid, medication, medication_age_days_firstDate)
```

```{r summary of total ASM prescriptions by age group, echo=FALSE, message=FALSE, warning=FALSE}
# asm medication summary
antiseizure_med_summary <- total_asm_subgr %>%
  
  group_by(patient_uuid, annual_age_groups) %>%
  
  # summary by patient at age group
  summarise(
    antiseizure_med_count_per_patient =
      sum(antiseizure_med == "Yes"),
    .groups = 'drop'
  ) %>%
  
  group_by(annual_age_groups) %>%
  
  # summary across age groups
  summarise(
    # Total number of rescue medications per age group
    antiseizure_med_count = round(sum(antiseizure_med_count_per_patient), 1),
    # Mean of rescue medications per patient
    mean_antiseizure_med_per_patient = round(mean(antiseizure_med_count_per_patient), 1),
    # Standard deviation of rescue medications per patient
    sd_antiseizure_med_per_patient = round(sd(antiseizure_med_count_per_patient), 1),
    # Median of rescue medications per patient
    median_antiseizure_med_per_patient = round(median(antiseizure_med_count_per_patient), 1),
    # 25th percentile (Q1)
    p25_antiseizure_med_per_patient = round(quantile(antiseizure_med_count_per_patient, 0.25), 1),
    # 75th percentile (Q3)
    p75_antiseizure_med_per_patient = round(quantile(antiseizure_med_count_per_patient, 0.75), 1),
    # patiens under follow up
    patients = n_distinct(patient_uuid),
    .groups = 'drop'
  ) %>%
  
  mutate(
    # Percentage of total rescue medications
    antiseizure_med_percentage = round((antiseizure_med_count / sum(antiseizure_med_count)) * 100, 1),
  ) %>%

  arrange(annual_age_groups)
```

```{r summary table - total ASM prescriptions by age group, echo=FALSE, message=FALSE, warning=FALSE}
# transpose table
antiseizure_med_summary_t <- antiseizure_med_summary %>%
  
  mutate( 
    `n (%)`= paste0(antiseizure_med_count,
                    " (", 
                    antiseizure_med_percentage,
                    ")"),
    `Mean (sd)` = paste0(mean_antiseizure_med_per_patient,
                         " (", 
                         sd_antiseizure_med_per_patient,
                         ")"),
    `Median [Q1-Q3]` = paste0(median_antiseizure_med_per_patient, 
                              " [",
                              p25_antiseizure_med_per_patient, 
                              "-", 
                              p75_antiseizure_med_per_patient, 
                              "]")
  ) %>%
  dplyr::select(
    annual_age_groups,
    `n (%)`, 
    `Mean (sd)` ,
    `Median [Q1-Q3]`,
    `Patients (N)` = patients
    ) %>%
  
  mutate(across(everything(), as.character)
         ) %>%
  
  pivot_longer(cols = -annual_age_groups, names_to = "Metric", values_to = "Value") %>%
  
  pivot_wider(names_from = annual_age_groups, values_from = Value)

# kable summary table
kable(antiseizure_med_summary_t, format = "html", caption = "Table 8. Summary of total anti-seizure medication prescriptions during follow-up by age groups") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2")  %>%
  footnote(
    general = "Age groups reflect the patient's age range during the medication period, calculated from the first and last recorded dates of intake.",
    general_title = "***",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "Antiseizure medication is defined as pharmacological treatments used to prevent or reduce the frequency and severity of seizures. This includes medications from various classes such as Brivaracetam, Lamotrigine, Levetiracetam, and others commonly prescribed for seizure management taken from https://www.epilepsy.com/tools-resources/seizure-medication-list",
    general_title = "**",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "Prescriptions with no information about dose/unit are counted.",
    general_title = "*",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "Medication use records with a specific medication name with continuous dates (< 2 days of gap) were grouped into the same episode.",
    general_title = "****",
    footnote_as_chunk = TRUE)

antiseizure_med_summary_t %>%
  write_csv("~/RWE-rare_disease/Age groups analyses/output/tables/antiseizure_med_summary_table.csv")
```

```{r antiseizure medication per patient plot, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE,  dpi = 300}
antiseizure_med_plot <- ggplot(antiseizure_med_summary, aes(x = annual_age_groups, y =  mean_antiseizure_med_per_patient, fill = case_when(
      annual_age_groups %in% c("[0,1)", "[1,2)") ~ "(0-2]", TRUE ~ "(2-12]"
    ))) +
  
  geom_col(position = "dodge", width = 0.6) + 
  
  #geom_text(aes(label = paste0("N = ", patients)), vjust = -3.5, size = 2.2) +
  
  labs(
    x = "Age Group (in years)",
    y = "Mean of ASMs prescriptions per patient",
    fill = "Age group")  +
  
  scale_fill_manual(values = c( "(2-12]" = "black", 
                                  "(0-2]" = "grey")) + 
  
  theme_minimal() +
  
  scale_y_continuous(breaks = seq(0, 23, 1), limits = c(0, 23)) +
  
  theme(
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

antiseizure_med_plot

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/antiseizure_med_per_patient_plot.png", 
       antiseizure_med_plot, units = "mm", width = 200, height = 150 , dpi = 300)
```

###### b. Distinct Anti-seizure medication prescription

```{r summary of number of different ASM prescriptions by age group, echo=FALSE, message=FALSE, warning=FALSE}
# Summary of different medications by age group
different_med_summary <- total_asm_subgr %>%
  
  group_by(patient_uuid, annual_age_groups) %>%
  
  # number of different medications by patient at the age group
  summarise(different_med_count_per_patient = n_distinct(medication),
            .groups = 'drop') %>%
  
  group_by(annual_age_groups) %>%
  
  # Summary of different medications across age groups
  summarise(
    # Total number of different medications per age group
    mean_different_med_per_patient = round(mean(different_med_count_per_patient), 1),
    # Standard deviation of different medications per patient
    sd_different_med_per_patient = round(sd(different_med_count_per_patient), 1),
    # Median of different medications per patient
    median_different_med_per_patient = round(median(different_med_count_per_patient), 1),
    # 25th percentile (Q1)
    p25_different_med_per_patient = round(quantile(different_med_count_per_patient, 0.25), 1),
    # 75th percentile (Q3)
    p75_different_med_per_patient = round(quantile(different_med_count_per_patient, 0.75), 1),
    # Number of patients under follow up
    patients = n_distinct(patient_uuid),
    .groups = 'drop'
  ) %>%

  arrange(annual_age_groups)
```

```{r summary table - number of different ASM prescriptions by age group, echo=FALSE, message=FALSE, warning=FALSE}
# Transpose table for display
different_med_summary_t <- different_med_summary %>%
  mutate( 
    `Mean (sd)` = paste0(mean_different_med_per_patient,
                         " (",
                         sd_different_med_per_patient,
                         ")"),
    `Median [Q1-Q3]` = paste0(median_different_med_per_patient, 
                               " [", 
                               p25_different_med_per_patient, 
                               "-", 
                               p75_different_med_per_patient, 
                               "]")
  ) %>%
  dplyr::select(annual_age_groups,
                `Mean (sd)` ,
                `Median [Q1-Q3]`,
                `Patients (N)` = patients) %>%
  
  mutate(across(everything(), as.character)) %>%
  
  pivot_longer(cols = -annual_age_groups, names_to = "Metric", values_to = "Value") %>%
  
  pivot_wider(names_from = annual_age_groups, values_from = Value)

# Create kable summary table
kable(different_med_summary_t, format = "html", caption = "Table 9. Summary of unique anti-seizure medications during follow-up by age groups") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2")  %>%
  footnote(
    general = "Age groups reflect the patient's age range during the medication period, calculated from the first and last recorded dates of intake.",
    general_title = "**",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "A 'different medication' is defined as a unique medication name (e.g., Brivaracetam, Lamotrigine, etc.) taken at any point during the follow-up period.",
    general_title = "*",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "Medication use records with a specific medication name with continuous dates (< 2 days of gap) were grouped into the same episode.",
    general_title = "***",
    footnote_as_chunk = TRUE)

different_med_summary_t %>%
  write_csv("~/RWE-rare_disease/Age groups analyses/output/tables/different_med_summary_table.csv")
```

```{r number different antiseizure medication per patient plot, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE,  dpi = 300}
asm_diff_plot <- ggplot(different_med_summary, aes(x = annual_age_groups, y =  mean_different_med_per_patient, fill = case_when(
      annual_age_groups %in% c("[0,1)", "[1,2)") ~ "(0-2]", TRUE ~ "(2-12]"
    ))) +
  
  geom_col(position = "dodge", width = 0.6) + 
  
 # geom_text(aes(label = paste0("N = ", patients)), vjust = -3.5, size = 2.2) +
  
  labs(
    x = "Age Group (in years)",
    y = "Mean of unique ASMs per patient",
    fill = "Age group")  +
  
  scale_fill_manual(values = c( "(2-12]" = "black", 
                                  "(0-2]" = "grey")) + 
  
  theme_minimal() +
  
  scale_y_continuous(breaks = seq(0, 8, 1), limits = c(0, 8)) +
  
  theme(
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

asm_diff_plot

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/asm.diff.plot.png", 
       asm_diff_plot, units = "mm", width = 200, height = 150 , dpi = 300)
```

###### c. Active ASMs number

```{r count of active ASM prescriptions by follow up day, echo=TRUE, message=FALSE, warning=FALSE}
## Creating a new table to count active ASM prescriptions
## To identify episodes with the same number of active ASMs, we need to count how many ASMs the patient receives for each day of follow-up
act_asm_subgr <- medication_antiseizure_subgrouping %>%

  rowwise() %>%

  mutate(days =
           list(
             seq(
               medication_age_days_firstDate, medication_age_days_lastDate,
               by = 1))) %>%

  unnest_longer(col = days) %>% # unnest days

  group_by(patient_uuid, days) %>%

  tally() %>% # n = number of concomitant ASMs for each day of follow-up

  group_by(patient_uuid) %>%

  mutate(episode_2 = data.table::rleid(n)) %>% ## Tally episodes when the patient has the same total number of ASM prescriptions

  group_by(patient_uuid, episode_2) %>%

  mutate(
    start_episode_age_days = min(days),
    end_episode_age_days = max(days),
    duration = end_episode_age_days - start_episode_age_days + 1) %>% # compute the duration of each episodes with same number of ASM prescriptions

  dplyr::select(-days) %>%

  distinct()
```

```{r active ASM by age group, echo=TRUE, message=FALSE, warning=FALSE}
act_asm_subgr <- act_asm_subgr %>%

  # Add age subgrouping
  mutate(
    age_sequence = list(
      seq(
        from = start_episode_age_days / 365.25,
        to = end_episode_age_days / 365.25,
        by = 1
      ) %>%
        unique() %>%
        c(end_episode_age_days / 365.25)
    )
  ) %>%
  
  unnest(cols = c(age_sequence)) %>%
  
  mutate(
    age_medication = age_sequence,
    
    # Derivation of in 1-year intervals
    annual_age_groups = cut(
      age_medication,
      breaks = c(0:12, Inf),
      right = F,
      include.lowest = T)
  ) %>%

  arrange(patient_uuid, episode_2, end_episode_age_days) %>%
  
  # Remove duplicates for the last age subgrouping records
  group_by(patient_uuid, episode_2, annual_age_groups) %>%
  
  slice_tail(n = 1) %>%
  
  ungroup() %>%
  
  # Filter out records outside the defined age range
  filter(!age_medication > 12) %>%

  # Select and organize relevant columns
  dplyr::select(
    patient_uuid,
    annual_age_groups,
    start_episode_age_days,
    end_episode_age_days,
    n,
    duration
    ) %>%
  
  # Sort the final dataset
  arrange(patient_uuid, start_episode_age_days)


# add rows with patients under risk but without records
act_asm_subgr <- follow_up_patients_uuid %>%
  
  left_join(act_asm_subgr, by = c("patient_uuid", "annual_age_groups")) %>%
  
  # n will be kept as NA for patients without records due as computed 0 active during follow-up were not included in the derivation
  mutate(
    n = replace_na(n, NA),
  ) %>%
  
  # Arrange by patient uuid
  arrange(patient_uuid, start_episode_age_days)
```

```{r summary of active ASMs by age group, echo=TRUE, message=FALSE, warning=FALSE, cache=FALSE}
# Calculate mean number of active medications per patient for each age group
act_asm_summary <- act_asm_subgr %>%
  
  group_by(patient_uuid, annual_age_groups) %>%
  
  # mean number of active medications per patient
  summarise(
    mean_n_per_patient = mean(n, na.rm = TRUE), 
    median_n_per_patient = median(n, na.rm = TRUE), 
    .groups = "drop"
  ) %>%
  
  group_by(annual_age_groups) %>%
  
  # mean number of active medications across age groups
  summarise(
    # mean of active.medications per patient by age group
    act_mean = round(mean(mean_n_per_patient, na.rm = TRUE), 1), 
    # standard deviation for active medications per patient 
    act_sd = round(sd(mean_n_per_patient, na.rm = TRUE), 1),
    # median number of active.med per patient by age group
    act_median = round(median(median_n_per_patient, na.rm = TRUE), 1),
    # percentile 25%
    act_p25 = round(quantile(median_n_per_patient, 0.25, na.rm = TRUE), 1),
    # percentile 75%
    act_p75 = round(quantile(median_n_per_patient, 0.75, na.rm = TRUE), 1),
    # patient under follow-up
    patients = n_distinct(patient_uuid),
    .groups = "drop"
  ) %>%

  arrange(annual_age_groups)
```

```{r summary table - active ASMs by age group, echo=TRUE, message=FALSE, warning=FALSE, cache=FALSE}
# transpose table
act_asm_med_summary.t <- act_asm_summary %>%
  
  mutate( 
    `Mean (sd)` = paste0(act_mean, 
                         " (",
                         act_sd,
                         ")"),
    `Median [Q1-Q3]` = paste0(act_median,
                              " [",
                              act_p25,
                              "-",
                              act_p75,
                              "]")
  ) %>%
  
  dplyr::select(
    annual_age_groups,
    `Mean (sd)`,
    `Median [Q1-Q3]`,
    `Patients (N)` = patients) %>%
  
  mutate(across(everything(), as.character)) %>%
  
  pivot_longer(cols = -annual_age_groups, names_to = "Metric", values_to = "Value") %>%
  
  pivot_wider(names_from = annual_age_groups, values_from = Value)

# kable summary table
kable(act_asm_med_summary.t, format = "html", caption = "Table 8. Summary of concomitant medications during follow-up by age groups") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  footnote(
    general = "Age groups reflect the patient's age range during the medication period, calculated from the first and last recorded dates of intake.",
    general_title = "**",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "Antiseizure medication is defined as pharmacological treatments used to prevent or reduce the frequency and severity of seizures. This includes medications from various classes such as Brivaracetam, Lamotrigine, Levetiracetam, and others commonly prescribed for seizure management taken from https://www.epilepsy.com/tools-resources/seizure-medication-list",
    general_title = "*",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "A concomitant medication is defined as the simultaneous use of two or more drugs to achieve the purpose of treatment  during the follow-up period.",
    general_title = "*",
    footnote_as_chunk = TRUE) %>%
  footnote(
    general = "Medication use records with a specific medication name with continuous dates (< 2 days of gap) were grouped into the same episode.",
    general_title = "***",
    footnote_as_chunk = TRUE)

act_asm_med_summary.t %>%
  write_csv("~/RWE-rare_disease/Age groups analyses/output/tables/act_asm_med_summary_table.csv")
```

```{r active asm per patient plot, echo=TRUE, fig.align='center', message=FALSE, warning=FALSE, dpi=300}
act_med_plot <- ggplot(act_asm_summary, aes(x = annual_age_groups, y =  act_mean, fill = case_when(
      annual_age_groups %in% c("[0,1)", "[1,2)") ~ "(0-2]", TRUE ~ "(2-12]"
    ))) +
  
  geom_col(position = "dodge", width = 0.6) + 
  
 # geom_text(aes(label = paste0("N = ", patients)), vjust = -3.5, size = 2.2) +
  
  labs(
    x = "Age Group (in years)",
    y = "Mean of concomitant ASMs per patient",
    fill = "Age group")  +
  
  scale_fill_manual(values = c( "(2-12]" = "black", 
                                  "(0-2]" = "grey")) + 
  
  theme_minimal() +
  
  scale_y_continuous(breaks = seq(0, 4, 1), limits = c(0, 5)) +
  
  theme(
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.position = "bottom",   
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )

act_med_plot

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/act_med_per_patient_plot.png", 
       act_med_plot, units = "mm", width = 200, height = 150 , dpi = 300)
```

#### Growth parameters

```{r Growth parameters derivation, echo=TRUE, message=FALSE, warning=FALSE}
growth_parameters_subgrouping <- rare_disease_growth_parameters %>%
  
  # Select relevant columns
  dplyr::select(
    patient_uuid,
    growth_parameter,
    growth_parameter_value,
    growth_parameter_unit,
    growth_parameter_age_days
  ) %>%
  
  # keeping only records without missing value/unit/age
  filter(!is.na(growth_parameter_value)  | !is.na(growth_parameter_unit) | !is.na(growth_parameter_age_days)) %>% 
  
  # remove duplicates
  distinct() %>%
  
  # For records with the same patient, growth parameter, and day, keep the record with the maximum value
  group_by(patient_uuid, growth_parameter, growth_parameter_age_days) %>%
  
  slice_max(growth_parameter_value, with_ties = FALSE) %>%
  
  ungroup() %>%
  
  # Derivation of measurement day during follow-up
  mutate(
    
    growth_parameter_measurement = "Yes",
    
    # Derivation of age at measurement
    age_measurement = growth_parameter_age_days / 365.25, # age in years
    
    # age groups by 1-year intervals
    annual_age_groups = cut(
      age_measurement,  
      breaks = c(0:12, Inf),  
      right = FALSE,  
      include.lowest = TRUE  
    )
  ) %>%
  
  # correct units in g to kg
  mutate(
      growth_parameter_value = ifelse(
        growth_parameter_unit == "g",
        growth_parameter_value/1000, # correct the unit to kg
        growth_parameter_value)
    ) %>%
  
  # Filter out records outside the age range 
  filter(age_measurement < 12) %>%
  
  # joining gender of the patient
  left_join(rare_disease_demographics_new, by = "patient_uuid") %>%
  
  mutate(sex = factor(sex)) %>%
  
  # keep selected columns for further processing
  dplyr::select(
    patient_uuid,
    growth_parameter,
    growth_parameter_value,
    growth_parameter_unit,
    growth_parameter_age_days,
    age_measurement,
    annual_age_groups,
    sex,
    growth_parameter_measurement
  ) %>%
  
  # arrange by patient uuid
  arrange(patient_uuid, growth_parameter_age_days)
```

##### Height overall

```{r summary of height parameter by age group, echo=FALSE, message=FALSE, warning=FALSE}
### CDC height for age 
h_cdc_data <- read.csv("~/RWE-rare_disease/Age groups analyses/raw.data/statage.csv") 

h_cdc <- h_cdc_data %>%
  
  mutate(
    # Creating subgroups for both sexes
    annual_age_groups = cut(
      Agemos / 12,  
      breaks = c(0:12, Inf),  
      right = FALSE,  
      include.lowest = TRUE  
    )
  ) %>%
  
  filter(annual_age_groups != "[12,Inf]") %>%
  
  group_by(annual_age_groups) %>%
  
  summarise(
    # Simple Median (Un-weighted)
    p50 = round(median(P50, na.rm = TRUE), 1),
    p05 = round(median(P5, na.rm = TRUE), 1),
    p25 = round(median(P25, na.rm = TRUE), 1),
    p75 = round(median(P75, na.rm = TRUE), 1),
    p95 = round(median(P95, na.rm = TRUE), 1),
    .groups = "drop"
  )


# summary of height parameter from RWD data
height_gp_summary <- follow_up_patients_uuid %>%
  
  # adding height parameter for patients under follow up without body height records for each age group
  left_join(filter(growth_parameters_subgrouping, growth_parameter == "Body height"), 
            by = c("patient_uuid", "annual_age_groups")) %>%
  
  mutate(
    growth_parameter_measurement = replace_na(growth_parameter_measurement, "No")
    ) %>%

  group_by(annual_age_groups) %>%  
  
  # summarise by age group to compare with the CDC data
  summarise(height_mean = round(mean(growth_parameter_value, na.rm = TRUE), 1),
            height_sd = round(sd(growth_parameter_value, na.rm = TRUE), 1),
            height_median = round(median(growth_parameter_value, na.rm = TRUE), 1),
            height_25 = round(quantile(growth_parameter_value, 0.25, na.rm = TRUE), 1),
            height_75 = round(quantile(growth_parameter_value, 0.75, na.rm = TRUE), 1),
            .groups = 'drop') %>%
  
  left_join(summary_count_uuid, by = "annual_age_groups")

  
# join CDC weight data
joined_height_data <- height_gp_summary %>%
  
  left_join(h_cdc, by = "annual_age_groups") %>%

  arrange(annual_age_groups)
```

```{r summary table - height parameter by age group, echo=FALSE, message=FALSE, warning=FALSE}
# transpose table and add CDC height median
height_gp_summary_t <- joined_height_data %>%
  
  mutate(
    `Mean (sd)` = paste0(height_mean, " (", height_sd, ")"),
    `Median [Q1-Q3]` = paste0(height_median, " [", height_25, "-", height_75, "]")
  ) %>%
  
  dplyr::select(
    annual_age_groups,
    `Mean (sd)`, 
    `Median [Q1-Q3]`,
    `Participants (N)` = patients,
    `Male (N)` = male,
    `Female (N)` = female) %>%
  
  mutate(across(everything(), as.character)) %>%
  
  pivot_longer(cols = -annual_age_groups, names_to = "Metric", values_to = "Value") %>%
  
  pivot_wider(names_from = annual_age_groups, values_from = Value)

kable(height_gp_summary_t, format = "html", caption = "Table 8. Summary of height parameter (cm) during follow-up by age groups") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  footnote(
    general = "Age groups defined by patient's age at measurement.",
    general_title = "*",
    footnote_as_chunk = TRUE)

height_gp_summary_t %>%
  write_csv("~/RWE-rare_disease/Age groups analyses/output/tables/height_gp_summary_table.csv")
```

```{r heigth plot, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, dpi= 300}
# Create the plot using percentiles (P5, P25, P50, P75, P95)
height_plot <- ggplot(joined_height_data, aes(x = annual_age_groups)) +
  
  # RWD data (using height median values for the example)
  geom_line(aes(y = height_median, group = 1, color = "RWD"), size = 0.5, linetype = "dashed") +      
  geom_point(aes(y = height_median, color = "RWD"), size = 2) +      

  # CDC data (using median of CDC percentiles)
  geom_line(aes(y = p50, group = 1, color = "CDC"), size = 0.5, linetype = "dashed") +  
  geom_point(aes(y = p50, color = "CDC"), size = 2) +     

  # Adding ribbons for the 25th and 75th percentiles (using another color)
  geom_ribbon(aes(ymin = p25, ymax = p75, fill = "P25–P75 (25th–75th range)", group = 2), alpha = 0.3) +

  # Adding ribbons for the 5th and 95th percentiles (using a different color 
  geom_ribbon(aes(ymin = p05, ymax = p95, fill = "P5–P95 (5th–95th range)", group = 2), alpha = 0.3) +
  
  # Customizing plot labels and appearance
  labs(
    x = "Age Group",
    y = "Median of height (cm)",
    color = "Source",
    fill = "Percentile"
  ) +
  scale_y_continuous(breaks = seq(50, 160, 10), limits = c(50, 160)) +
  theme_minimal() +
  theme(
    [BQ_REF_REDACTED] = element_blank(),        
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8),
    legend.position = "bottom"
  ) +
  scale_color_manual(values = c("RWD" = "blue", "CDC" = "red")) +
  scale_fill_manual(values = c("P25–P75 (25th–75th range)" = "lightcoral", "P5–P95 (5th–95th range)" = "orange"))

height_plot

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/height_plot.png", 
       height_plot, units = "mm", width = 200, height = 150 , dpi = 300)
```

###### Height by sex

```{r summary of height parameter by sex, echo=TRUE, message=FALSE, warning=FALSE}
h_cdc_sex <- h_cdc_data %>%
  
  mutate(
    # Creating subgroups for both sexes
    annual_age_groups = cut(
      Agemos / 12,  
      breaks = c(0:12, Inf),  
      right = FALSE,  
      include.lowest = TRUE  
    )
  ) %>%

  filter(annual_age_groups != "[12,Inf]") %>%
  
  group_by(annual_age_groups, Sex) %>%
  
  # summarization of medians for each percentile
  summarise(
    p_50 = round(median(P50, na.rm = TRUE), 1), # or median of P50
    p_05 = round(median(P5, na.rm = TRUE), 1),
    p_25 = round(median(P25, na.rm = TRUE), 1),
    p_75 = round(median(P75, na.rm = TRUE), 1),
    p_95 = round(median(P95, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>% 
  
  rename(sex = Sex) %>%
  
  mutate(
    sex = factor(
      case_when(
        sex == 1 ~ "male" ,  
        sex == 2 ~ "female" 
      ),
      levels = c("male", "female")     
    )
  )

# summary of height parameter from RWD data
height_gp_summary_sex <- follow_up_patients_uuid %>%
  
  # adding height parameter for patients under follow up without body height records for each age group 
  left_join(filter(growth_parameters_subgrouping, growth_parameter == "Body height"), 
            by = c("patient_uuid", "annual_age_groups", "sex")) %>%
  
  mutate(
    growth_parameter_measurement = replace_na(growth_parameter_measurement, "No")
  ) %>%

  group_by(annual_age_groups, sex) %>%
  
  # summarise by age group and sex and not by patient first due to low number of patients with height records by sex (to check here)
  summarise(height_mean = round(mean(growth_parameter_value, na.rm = TRUE), 1),
            height_sd = round(sd(growth_parameter_value, na.rm = TRUE), 1),
            height_median = round(median(growth_parameter_value, na.rm = TRUE), 1),
            height_25 = round(quantile(growth_parameter_value, 0.25, na.rm = TRUE), 1),
            height_75 = round(quantile(growth_parameter_value, 0.75, na.rm = TRUE), 1),
            patients = n_distinct(patient_uuid),
            .groups = 'drop') 


# join CDC weight data
joined_height_sex <- left_join(
  
  height_gp_summary_sex, h_cdc_sex, by = c("annual_age_groups", "sex")) %>%

  arrange(annual_age_groups)
```

```{r heigth plot by sex, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, dpi= 300}
# Create the plot using percentiles (P5, P25, P50, P75, P95)
height_plot_sex <- ggplot(joined_height_sex, aes(x = annual_age_groups)) +
  
  # RWD data (using height median values for the example)
  geom_line(aes(y = height_median, group = 1, color = "RWD"), size = 0.5, linetype = "dashed") +      
  geom_point(aes(y = height_median, color = "RWD"), size = 2) +      

  # CDC data (using median of CDC percentiles)
  geom_line(aes(y = p_50, group = 1, color = "CDC"), size = 0.5, linetype = "dashed") +  
  geom_point(aes(y = p_50, color = "CDC"), size = 2) +     

  # Adding ribbons for the 25th and 75th percentiles (using another color)
  geom_ribbon(aes(ymin = p_25, ymax = p_75, fill = "P25–P75 (25th–75th range)", group = 2), alpha = 0.3) +

  # Adding ribbons for the 5th and 95th percentiles (using a different color 
  geom_ribbon(aes(ymin = p_05, ymax = p_95, fill = "P5–P95 (5th–95th range)", group = 2), alpha = 0.3) +
  
  # Customizing plot labels and appearance
  labs(
    x = "Age Group",
    y = "Median of height (cm)",
    color = "Source",
    fill = "Percentile"
  ) +
  scale_y_continuous(breaks = seq(50, 160, 10), limits = c(50, 160)) +
  theme_minimal() +
  theme(
    [BQ_REF_REDACTED] = element_blank(),        
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8),
    legend.position = "bottom"
  ) +
  scale_color_manual(values = c("RWD" = "blue", "CDC" = "red")) +
  scale_fill_manual(values = c("P25–P75 (25th–75th range)" = "lightcoral", "P5–P95 (5th–95th range)" = "orange")) +
  
  facet_wrap(~ sex)

height_plot_sex

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/height_plot_by_sex.png", 
       height_plot_sex, units = "mm", width = 200, height = 150 , dpi = 300)
```

##### Weight

```{r summary of weight parameter by age group, echo=TRUE, message=FALSE, warning=FALSE}
### CDC weight for age
w_cdc_data <- read.csv("~/RWE-rare_disease/Age groups analyses/raw.data/wtage.csv") 

w_cdc <- w_cdc_data %>%
  
  mutate(
    # Creating subgroups for both sexes
    annual_age_groups = cut(
      Agemos / 12,  
      breaks = c(0:12, Inf),  
      right = FALSE,  
      include.lowest = TRUE  
    )
  ) %>%

  filter(annual_age_groups != "[12,Inf]") %>%
  
  group_by(annual_age_groups) %>%

  summarise(
    p_50 = round(median(P50, na.rm = TRUE), 1),
    p_05 = round(median(P5, na.rm = TRUE), 1),
    p_25 = round(median(P25, na.rm = TRUE), 1),
    p_75 = round(median(P75, na.rm = TRUE), 1),
    p_95 = round(median(P95, na.rm = TRUE), 1),
    .groups = "drop"
  )

# summary of weight parameter from RWD data
weight_gp_summary <- follow_up_patients_uuid %>%
  
  # adding height parameter for patients under follow up without body height records for each age group
  left_join(filter(growth_parameters_subgrouping, growth_parameter == "Body weight"), 
            by = c("patient_uuid", "annual_age_groups")) %>%
  
  mutate(
    growth_parameter_measurement = replace_na(growth_parameter_measurement, "No")
    ) %>%  

  group_by(annual_age_groups) %>%
  
  summarise(weight_mean = round(mean(growth_parameter_value, na.rm = TRUE), 1),
            weight_sd = round(sd(growth_parameter_value, na.rm = TRUE), 1),
            weight_median = round(median(growth_parameter_value, na.rm = TRUE), 1),
            weight_25 = round(quantile(growth_parameter_value, 0.25, na.rm = TRUE), 1),
            weight_75 = round(quantile(growth_parameter_value, 0.75, na.rm = TRUE), 1),
            .groups = 'drop') %>%
  
  left_join(summary_count_uuid, by = "annual_age_groups")


# join CDC weight data
joined_weight_data <- weight_gp_summary %>%
  
  left_join(w_cdc, by = "annual_age_groups") %>%

  arrange(annual_age_groups)
```

```{r summary table - weight parameter by age group, echo=TRUE, message=FALSE, warning=FALSE}
# summary both in transpose table
weight_gp_summary_t <- joined_weight_data %>%
  
  mutate(
    `Mean (sd)` = paste0(weight_mean, " (", weight_sd, ")"),
    `Median [Q1-Q3]` = paste0(weight_median, " [", weight_25, "-", weight_75, "]")
  ) %>%
  
  dplyr::select(
    annual_age_groups, 
    `Mean (sd)`,
    `Median [Q1-Q3]`,
    `Patients (N)` = patients,
    `Male (N)` = male,
    `Female (N)` = female) %>%
  
  mutate(across(everything(), as.character)) %>%
  
  pivot_longer(cols = -annual_age_groups, names_to = "Metric", values_to = "Value") %>%
  
  pivot_wider(names_from = annual_age_groups, values_from = Value)

kable(weight_gp_summary_t, format = "html", caption = "Table 8. Summary of weight parameter (kg) during follow-up by age groups") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  footnote(
    general = "Age groups defined by patient's age at measurement.",
    general_title = "*",
    footnote_as_chunk = TRUE)

weight_gp_summary_t %>%
  write_csv("~/RWE-rare_disease/Age groups analyses/output/tables/weight_gp_summary_table.csv")
```

```{r weigth plot, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE,  dpi = 300}
# Create the plot using percentiles (P5, P25, P50, P75, P95)
weight_plot <- ggplot(joined_weight_data, aes(x = annual_age_groups)) +
  
  # RWD data (using height median values for the example)
  geom_line(aes(y = weight_median, group = 1, color = "RWD"), size = 0.5, linetype = "dashed") +      
  geom_point(aes(y = weight_median, color = "RWD"), size = 2) +      

  # CDC data (using median of CDC percentiles)
  geom_line(aes(y = p_50, group = 1, color = "CDC"), size = 0.5, linetype = "dashed") +  
  geom_point(aes(y = p_50, color = "CDC"), size = 2) +     

  # Adding ribbons for the 25th and 75th percentiles (using another color)
  geom_ribbon(aes(ymin = p_25, ymax = p_75, fill = "P25–P75 (25th–75th range)", group = 2), alpha = 0.3) +

  # Adding ribbons for the 5th and 95th percentiles (using a different color 
  geom_ribbon(aes(ymin = p_05, ymax = p_95, fill = "P5–P95 (5th–95th range)", group = 2), alpha = 0.3) +
  
  # Customizing plot labels and appearance
  labs(
    x = "Age Group",
    y = "Median of weight (kg)",
    color = "Source",
    fill = "Percentile"
  ) +
  scale_y_continuous(breaks = seq(0, 60, 10), limits = c(0, 60)) +
  theme_minimal() +
  theme(
    [BQ_REF_REDACTED] = element_blank(),        
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8),
    legend.position = "bottom"
  ) +
  scale_color_manual(values = c("RWD" = "blue", "CDC" = "red")) +
  scale_fill_manual(values = c("P25–P75 (25th–75th range)" = "lightcoral","P5–P95 (5th–95th range)" = "orange"))

weight_plot

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/weight_plot.png", 
       weight_plot, units = "mm", width = 200, height = 150 , dpi = 300)
```

###### Weight by sex

```{r summary of weight parameter by sex, echo=TRUE, message=FALSE, warning=FALSE}
w_cdc_sex <- w_cdc_data %>%
  
  mutate(
    # Creating subgroups for both sexes
    annual_age_groups = cut(
      Agemos / 12,  
      breaks = c(0:12, Inf),  
      right = FALSE,  
      include.lowest = TRUE  
    )
  ) %>%

  filter(annual_age_groups != "[12,Inf]") %>%
  
  group_by(annual_age_groups, Sex) %>%
  
  summarise(
    p_50 = round(median(P50, na.rm = TRUE), 1),
    p_05 = round(median(P5, na.rm = TRUE), 1),
    p_25 = round(median(P25, na.rm = TRUE), 1),
    p_75 = round(median(P75, na.rm = TRUE), 1),
    p_95 = round(median(P95, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>% 
  
  rename(sex = Sex) %>%
  
  mutate(
    sex = factor(
      case_when(
        sex == 1 ~ "male" ,  
        sex == 2 ~ "female" 
      ),
      levels = c("male", "female")     
    )
  )


# summary of height parameter from RWD data
weight_gp_summary_sex <- follow_up_patients_uuid %>%
  
  # adding height parameter for patients under follow up without body height records for each age group 
  left_join(filter(growth_parameters_subgrouping, growth_parameter == "Body weight"), 
            by = c("patient_uuid", "annual_age_groups", "sex")) %>%
  
  mutate(
    growth_parameter_measurement = replace_na(growth_parameter_measurement, "No")
  ) %>%
  
  group_by(annual_age_groups, sex) %>%  
  
  summarise(weight_mean = round(mean(growth_parameter_value, na.rm = TRUE), 1),
            weight_sd = round(sd(growth_parameter_value, na.rm = TRUE), 1),
            weight_median = round(median(growth_parameter_value, na.rm = TRUE), 1),
            weight_25 = round(quantile(growth_parameter_value, 0.25, na.rm = TRUE), 1),
            weight_75 = round(quantile(growth_parameter_value, 0.75, na.rm = TRUE), 1),
            weight_min = round(min(growth_parameter_value, na.rm = TRUE), 1),
            weight_max = round(max(growth_parameter_value, na.rm = TRUE), 1),
            .groups = 'drop')

  
# join CDC weight data
joined_weight_sex <- left_join(
  
  weight_gp_summary_sex, w_cdc_sex, by = c("annual_age_groups", "sex")) %>%

  arrange(annual_age_groups)
```

```{r weigth plot by sex, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, dpi= 300}
# Create the plot using percentiles (P5, P25, P50, P75, P95)
weight_plot_sex <- ggplot(joined_weight_sex, aes(x = annual_age_groups)) +
  
  # RWD data (using height median values for the example)
  geom_line(aes(y = weight_median, group = 1, color = "RWD"), size = 0.5, linetype = "dashed") +      
  geom_point(aes(y = weight_median, color = "RWD"), size = 2) +      

  # CDC data (using median of CDC percentiles)
  geom_line(aes(y = p_50, group = 1, color = "CDC"), size = 0.5, linetype = "dashed") +  
  geom_point(aes(y = p_50, color = "CDC"), size = 2) +     

  # Adding ribbons for the 25th and 75th percentiles (using another color)
  geom_ribbon(aes(ymin = p_25, ymax = p_75, fill = "P25–P75 (25th–75th range)", group = 2), alpha = 0.3) +

  # Adding ribbons for the 5th and 95th percentiles (using a different color 
  geom_ribbon(aes(ymin = p_05, ymax = p_95, fill = "P5–P95 (5th–95th range)", group = 2), alpha = 0.3) +
  
  # Customizing plot labels and appearance
  labs(
    x = "Age Group",
    y = "Median of weight (kg)",
    color = "Source",
    fill = "Percentile"
  ) +
  scale_y_continuous(breaks = seq(0, 60, 10), limits = c(0, 60)) +
  theme_minimal() +
  theme(
    [BQ_REF_REDACTED] = element_blank(),        
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8),
    legend.position = "bottom"
  ) +
  scale_color_manual(values = c("RWD" = "blue", "CDC" = "red")) +
  scale_fill_manual(values = c("P25–P75 (25th–75th range)" = "lightcoral", "P5–P95 (5th–95th range)" = "orange")) +
  
  facet_wrap(~ sex) 

weight_plot_sex

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/weight_plot_by_sex.png", 
       weight_plot_sex, units = "mm", width = 200, height = 150 , dpi = 300)
```

##### Head circumference

```{r summary of head circumference parameter by age group, echo=TRUE, message=FALSE, warning=FALSE}
### WHO head circumference until 5 years old
hc_who_data <- read.csv("~/RWE-rare_disease/Age groups analyses/raw.data/tab_hcfa_girls_boys_p_0_5.csv") 

hc_who <- hc_who_data %>%
  mutate(
    # Creating subgroups for both sexes
    annual_age_groups = cut(
      Month / 12,  
      breaks = c(0:5, Inf),  
      right = FALSE,  
      include.lowest = TRUE  
    )
  ) %>%

  filter(annual_age_groups != "[5,Inf]") %>%
  
  group_by(annual_age_groups) %>%
  
  summarise(
    p_50 = round(median(P50, na.rm = TRUE), 1), # Median of WHO percentiles
    p_05 = round(median(P5, na.rm = TRUE), 1),
    p_25 = round(median(P25, na.rm = TRUE), 1),
    p_75 = round(median(P75, na.rm = TRUE), 1),
    p_95 = round(median(P95, na.rm = TRUE), 1),
    .groups = "drop"
  )

# summary of head circumference parameter from RWD data
head_circum_gp_summary <- follow_up_patients_uuid %>%
  
  # adding height parameter for patients under follow up without body height records for each age group
  left_join(filter(growth_parameters_subgrouping, growth_parameter == "Head circumference"), 
            by = c("patient_uuid", "annual_age_groups")) %>%
  
  mutate(
    growth_parameter_measurement = replace_na(growth_parameter_measurement, "No")
    ) %>% 
  
  group_by(annual_age_groups) %>%
  
  summarise(head_circum_mean = round(mean(growth_parameter_value, na.rm = TRUE), 1),
            head_circum_sd = round(sd(growth_parameter_value, na.rm = TRUE), 1),
            head_circum_median = round(median(growth_parameter_value, na.rm = TRUE), 1),
            head_circum_25 = round(quantile(growth_parameter_value, 0.25, na.rm = TRUE), 1),
            head_circum_75 = round(quantile(growth_parameter_value, 0.75, na.rm = TRUE), 1),
            .groups = 'drop') %>%
  
  left_join(summary_count_uuid, by = "annual_age_groups")

# join WHO head circumference data
joined_hc_data <- head_circum_gp_summary %>%
  
  left_join(hc_who, by = "annual_age_groups") %>%

  arrange(annual_age_groups)
```

```{r summary table - head circumference parameter by age group, echo=TRUE, message=FALSE, warning=FALSE}
# transpose table
head_circum_gp_summary_t <- head_circum_gp_summary %>%
  
  mutate(
    `Mean (sd)` = paste0(head_circum_mean, " (", head_circum_sd, ")"),
    `Median, [Q1-Q3]` = paste0(head_circum_median, " [", head_circum_25, "-", head_circum_75, "]")
  ) %>%
  
  dplyr::select(
    annual_age_groups,
    `Mean (sd)`,
    `Median, [Q1-Q3]`,
    `Patients (N)` = patients,
    `Male (N)` = male,
    `Female (N)` = female) %>%
  
  mutate(across(everything(), as.character)) %>%
  
  pivot_longer(cols = -annual_age_groups, names_to = "Metric", values_to = "Value") %>%
  
  pivot_wider(names_from = annual_age_groups, values_from = Value) 

kable(head_circum_gp_summary_t, format = "html", caption = "Table 9. Summary of head circumference (cm) parameter during follow-up by age groups") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  footnote(
    general = "Age groups defined by patient's age at measurement.",
    general_title = "*",
    footnote_as_chunk = TRUE)

head_circum_gp_summary_t %>%
  write_csv("~/RWE-rare_disease/Age groups analyses/output/tables/head_circum_gp_summary_table.csv")
```

```{r head circumference plot, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE,  dpi = 300}
# Create the plot using percentiles (P5, P25, P50, P75, P95)
head_circum_plot <- ggplot(joined_hc_data, aes(x = annual_age_groups)) +
  
  # RWD data (using height median values for the example)
  geom_line(aes(y = head_circum_median, group = 1, color = "RWD"), size = 0.5, linetype = "dashed") +      
  geom_point(aes(y = head_circum_median, color = "RWD"), size = 2) +      

  # CDC data (using median of CDC percentiles)
  geom_line(aes(y = p_50, group = 1, color = "WHO"), size = 0.5, linetype = "dashed") +  
  geom_point(aes(y = p_50, color = "WHO"), size = 2) +     

  # Adding ribbons for the 25th and 75th percentiles (using another color)
  geom_ribbon(aes(ymin = p_25, ymax = p_75, fill = "P25–P75 (25th–75th range)", group = 2), alpha = 0.3) +

  # Adding ribbons for the 5th and 95th percentiles (using a different color 
  geom_ribbon(aes(ymin = p_05, ymax = p_95, fill = "P5–P95 (5th–95th range)", group = 2), alpha = 0.3) +
  
  # Customizing plot labels and appearance
  labs(
    x = "Age Group",
    y = "Median of head circumference (cm)",
    color = "Source",
    fill = "Percentile"
  ) +
  scale_y_continuous(breaks = seq(30, 60, 5), limits = c(30, 60)) +
  theme_minimal() +
  theme(
    [BQ_REF_REDACTED] = element_blank(),        
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8),
    legend.position = "bottom"
  ) +
  scale_color_manual(values = c("RWD" = "blue", "WHO" = "red")) +
  scale_fill_manual(values = c("P25–P75 (25th–75th range)" = "lightcoral", "P5–P95 (5th–95th range)" = "orange"))

head_circum_plot

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/head.[BQ_REF_REDACTED]", 
       head_circum_plot, units = "mm", width = 200, height = 150 , dpi = 300)
```

###### Head circumference by sex

```{r summary of head circumference parameter by sex, echo=TRUE, message=FALSE, warning=FALSE}
### WHO head circumference until 5 years old
hc_who_sex <- hc_who_data %>%
  mutate(
    # Creating subgroups for both sexes
    annual_age_groups = cut(
      Month / 12,  
      breaks = c(0:5, Inf),  
      right = FALSE,  
      include.lowest = TRUE  
    )
  ) %>%

  filter(annual_age_groups != "[5,Inf]") %>%
  
  group_by(annual_age_groups, L) %>%
  
  summarise(
    p_50 = round(median(P50, na.rm = TRUE), 1), # Median of WHO percentiles
    p_05 = round(median(P5, na.rm = TRUE), 1),
    p_25 = round(median(P25, na.rm = TRUE), 1),
    p_75 = round(median(P75, na.rm = TRUE), 1),
    p_95 = round(median(P95, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  
  rename(sex = L) %>%
  
  mutate(
    sex = factor(
      case_when(
        sex == 1 ~ "male" ,  
        sex == 2 ~ "female" 
      ),
      levels = c("male", "female")     
    )
  )

# summary of head circumference parameter from RWD data
head_circum_gp_summary.sex <- follow_up_patients_uuid %>%
  
  # adding height parameter for patients under follow up without body height records for each age group 
  left_join(filter(growth_parameters_subgrouping, growth_parameter == "Head circumference"), 
            by = c("patient_uuid", "annual_age_groups", "sex")) %>%
  
  mutate(
    growth_parameter_measurement = replace_na(growth_parameter_measurement, "No")
  ) %>%

  group_by(annual_age_groups, sex) %>%  
  
  summarise(head_circum_mean = round(mean(growth_parameter_value, na.rm = TRUE), 1),
            head_circum_sd = round(sd(growth_parameter_value, na.rm = TRUE), 1),
            head_circum_median = round(median(growth_parameter_value, na.rm = TRUE), 1),
            head_circum_25 = round(quantile(growth_parameter_value, 0.25, na.rm = TRUE), 1),
            head_circum_75 = round(quantile(growth_parameter_value, 0.75, na.rm = TRUE), 1),
            .groups = 'drop') 


# join WHO head circumference data
joined_hc_data_sex <- left_join(
  head_circum_gp_summary.sex, hc_who_sex, by = c("annual_age_groups", "sex")) %>%
  
  arrange(annual_age_groups)
```

```{r head circumference plot by sex, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE,  dpi = 300}
# Create the plot using percentiles (P5, P25, P50, P75, P95)
head_circum_plot_sex <- ggplot(joined_hc_data_sex, aes(x = annual_age_groups)) +
  
  # RWD data (using height median values for the example)
  geom_line(aes(y = head_circum_median, group = 1, color = "RWD"), size = 0.5, linetype = "dashed") +      
  geom_point(aes(y = head_circum_median, color = "RWD"), size = 2) +      

  # CDC data (using median of CDC percentiles)
  geom_line(aes(y = p_50, group = 1, color = "WHO"), size = 0.5, linetype = "dashed") +  
  geom_point(aes(y = p_50, color = "WHO"), size = 2) +     

  # Adding ribbons for the 25th and 75th percentiles (using another color)
  geom_ribbon(aes(ymin = p_25, ymax = p_75, fill = "P25–P75 (25th–75th range)", group = 2), alpha = 0.3) +

  # Adding ribbons for the 5th and 95th percentiles (using a different color 
  geom_ribbon(aes(ymin = p_05, ymax = p_95, fill = "P5–P95 (5th–95th range)", group = 2), alpha = 0.3) +
  
  # Customizing plot labels and appearance
  labs(
    x = "Age Group",
    y = "Median of head circumference (cm)",
    color = "Source",
    fill = "Percentile"
  ) +
  scale_y_continuous(breaks = seq(30, 60, 5), limits = c(30, 60)) +
  theme_minimal() +
  theme(
    [BQ_REF_REDACTED] = element_blank(),       
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8),
    legend.position = "bottom"
  ) +
  scale_color_manual(values = c("RWD" = "blue", "WHO" = "red")) +
  scale_fill_manual(values = c("P25–P75 (25th–75th range)" = "lightcoral", "P5–P95 (5th–95th range)" = "orange")) + 
  
  facet_wrap(~ sex)

head_circum_plot_sex

ggsave("~/RWE-rare_disease/Age groups analyses/output/plots/head.[BQ_REF_REDACTED].sex.png", 
       head_circum_plot_sex, units = "mm", width = 200, height = 150 , dpi = 300)
```
