---
title: "Data management and modeling pipeline (portfolio example)"
subtitle: "Section 1: data management"
author: "Redacted"
css: styles.css
---

```{r setup, echo = F}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
options(knitr.kable.NA = "")
```

#Load config, functions files and import data from GCP

```{r}
#| label: load-functions and config files
#| echo: false
#| warning: false
#| results: false

#' Import config file: includes list of packages used accross reports
#' and files to upload from cloud
source("../src/config.R") 
# Import all developed functions: import data, load packages, readmission_find...
source("../src/functions.R")
```

```{r}
#| label: load-packages
#| echo: false
#| warning: false
#| results: false
#| message: false
#| error: false
#| cache: false
#| include: false

load_libraries(PACKAGES)
```

```{r}
#| label: import-data
#| echo: false
#| warning: false
#| results: false
#| message: false
#| include: false

# Client Authentification
# do not displayed

# Import epilepsy-related genes for genetic mutation variable derivation
# Source https://bahlolab.github.io/Genes4Epilepsy/
gene4epilepsy_list_202503 <- gcs_get_object(
  "gene4epilepsy/gene4epilepsy_list_202503.rds",
  bucket = "YOUR_GCS_BUCKET",
  parseFunction = gcs_parse_rds
)
```

# Variable derivation based on the statistical analyses plan

## Seizure-related admissions and length of hospital stay (LOHS) derivation

```{r}
#| label: Derivation of hospitalization-related variables from admission dataset
#| results: false

# Definitions:
#' Length of hospital stay (LOHS):
#' The LOHS for each hospital admission record from the “Hospital Admission”
#' dataset will be coded as a quantitative variable describing the difference
#'  in days between the admission date (“admission_age_days_firstDate”) and
#'   discharge date (“admission_age_days_lastDate”) plus one
#'   (to account for the inclusion of the admission day in the duration).
#'    Additionally, the LOHS will be derived as categorical variable based on
#'    tertiles (i.e., bottom, middle, and top tertiles), as proposed
#'    by Han TS et al., (2022).

#' SR admission diagnosis:
#' This variable will be treated as a binary variable (i.e., coded as “yes”/”no”)
#' indicating whether a hospital admission record from the “Hospital Admission”
#'  dataset has a seizure-related diagnosis. A seizure-related diagnosis will be
#'   identified based on the variable ‘admission_diagnosis‘ in the "Hospital
#'   Admission" dataset. Admission diagnoses records will be classified as
#'   seizure-related type if they contain any of the following terms: “seizure,
#'   ” “epilepsy,” or “status epilepticus”

#' Note 1: for derivation of all binary variables:
#' Although binary variables in the SAP are defined as "yes"/"no" to improve
#' readability in documentation, they were implemented in the analysis
#' code using 0 and 1 for practical and computational efficiency.

#' Note 2: this df will be used as main input for the derivation of the rest
#' of covariates and 30-day UHRs in the following steps

#' Seizure event terms detected with the words:
#'  “seizure,” “epilepsy,” & “status epilepticus”
SEIZURE_EVENT <- c(
  "Increased seizure frequency", # 'seizure' term
  "Seizure disorder", # 'seizure' term
  "Epilepsy monitoring unit admission", # 'epilepsy' term
  "Epilepsy", # 'epilepsy' term
  "Status epilepticus" # 'status epilepticus' term
)

# Admission diagnoses with the word "birth"
#' Prematurity of infant is not included as it is a disorder on SNOMED
#' It was checked and revised with PEs.
BIRTH_EVENT <- c(
  "Birth",
  "Term birth of newborn",
  "Preterm birth"
)

# Considered planned admission diagnoses with the word "planned"
PLANNED_EVENT <- c(
  "Planned procedure",
  "Planned admission",
  "Planned admission for research study"
)

hospital_admission_rec <- rare_hospital_admission %>%
  # Select variables of interest
  dplyr::select(
    # Patient UUID
    patient_uuid,
    # Admission age date
    admission_age_days_firstDate,
    # Discharge age date
    admission_age_days_lastDate,
    # Admission diagnosis
    admission_diagnosis
  ) %>%
  # derive variables of interest
  mutate(
    # Flag birth admissions
    birth_admission = factor((admission_diagnosis %in% BIRTH_EVENT) * 1),

    # Flag birth admissions
    planned_admission = factor((admission_diagnosis %in% PLANNED_EVENT) * 1),

    # Flag outpatient admissions (admission and discharge in the same day)
    outpatient_admission = factor((
      admission_age_days_firstDate == admission_age_days_lastDate) * 1),

    # Derivation of continuous length of hospital stays in days
    length_of_stay_days =
      (admission_age_days_lastDate - admission_age_days_firstDate) + 1,

    #' Derivation of length of hospital stays in 3 categories by tertiles
    length_of_stay_days_cat = cut(
      length_of_stay_days,
      # cut in tertiles
      quantile(
        length_of_stay_days,
        c(0:3 / 3)
      ),
      include.lowest = T,
      labels = c("Bottom", "Middle", "Top")
    ),

    # Derivation of age at hospital admission in years
    age_admission_yr = as.numeric(admission_age_days_firstDate / 365.25),

    # Determine type of admission diagnosis based on seizures/epilepsy
    seizure_related_admission = factor((admission_diagnosis %in% SEIZURE_EVENT
      ) * 1,
      # having not seizure-related as reference
      levels = c(0, 1)
    )
  ) %>%
  #' Remove duplicated based on patient age at admission day, patient age at
  #'  discharge day, admission diagnosis and patient unique identifier
  distinct(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    admission_diagnosis,
    .keep_all = TRUE
  ) %>%
  # Select and order variables of interest
  dplyr::select(
    patient_uuid,
    age_admission_yr,
    admission_diagnosis,
    seizure_related_admission,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    length_of_stay_days,
    length_of_stay_days_cat,
    birth_admission,
    planned_admission,
    outpatient_admission
  ) %>%
  # Sort by patient and the first day of hospitalization
  arrange(patient_uuid, admission_age_days_firstDate)
```

## Sex variable derivation

```{r}
#| label: patient sex
#| results: false

#' Definition:
#' Patient sex at birth will be defined as a categorical variable
#' according to the “sex” variable from the “Demographics” dataset.

demographics_rec <- rare_disease_demographics %>%
  # check deceased patients with admission history data
  dplyr::select(
    patient_uuid,
    sex
  ) %>%
  mutate(
    sex = factor(sex, levels = c("female", "male"))
  )
```

## Follow-up variable derivation

```{r}
#| label: patient follow-up days
#| results: false
#' Definition:
#' Patient age (in days) at first record will be identified for each patient,
#'  by extracting the minimum patient age from all the patient’s records across
#'   all datasets (i.e., listed in Section 3.3). It will be defined as a 
#'   quantitative variable and will be converted into months and years using 
#'   the conversion conventions described in Section 4.1.

#' Patient age (in days) at last record will be identified for each patient, by
#'  extracting the maximum patient age from all the patient’s records across all
#'   datasets (i.e., listed in Section 3.3). It will be defined as a quantitative
#'    variable and will be converted into months and years using the conversion
#'     conventions described in Section 4.1.

followup_rec <- rare_disease_followup_durations_days %>%
  # Select variables of interest
  dplyr::select(
    patient_uuid,
    first_record_age_days,
    last_record_age_days
  )
```

## Flag deceased patients

```{r}
#| label: flag deceased patients
#| results: false
#' Definition based on patient age at death:
#' Patient age (in days) at death will be defined from the “Primary Diagnosis”
#' dataset. It will be defined as a quantitative variable, and its value will
#'  be based on the “primary_diagnosis_age_days” variable at the death record.
#'  This variable will only be computed for the patients with a documented death.

# Identify id of deceased patients and date of death for further use
deceased_patients <- rare_disease_primary_diagnosis %>%
  filter(primary_diagnosis == "Deceased") %>%
  dplyr::select(
    patient_uuid,
    primary_diagnosis_age_days
  )
```

## 30-day UHRs derivation (derived from prior output df)

```{r}
#| label: 30-day unplanned hospital readmissions (UHR) derivation
#| results: false

#' Definition:
#' Raw hospital admissions records will be used to derive the 30-day UHR outcome.
#'  A hospital admission record will be considered as a healthcare episode in 
#'  which a patient is admitted to a hospital facility with more than 24 hs
#'  It will be defined by an admission start day and
#'    the discharge day variables, which are dated to a specific patient age 
#'    in days in the “Hospital admission” database:  
#'    
#'    - Admission start day: patient age at admission day 
#'    (variable ”admission_age_days firstDate”)  
#'  
#'    - Discharge day: patient age at discharge day 
#'    (variable “admission_age_days_lastDate”)   
#'    
#'  “Planned”, “birth” and outpatient admissions admission diagnosis reasons
#'     will be excluded from the analysis.
#'    
#'  Each hospital admission record occurring after an index discharge day 
#'    (discharge day of a hospital admission that marks the start of a 30-day 
#'    observation window) will be assessed to determine if it corresponds to a 
#'    30-day UHR event.If multiple 30-day UHRs occur within the same 30-day 
#'    window after last index discharge date, all hospital admission records
#'     will be counted. After this 30-day period, any subsequent hospital 
#'     admission is considered a new index admission (with a respective index 
#'     discharge day), initiating a new 30-day observation window.  

# 1) Flag index and UHRs-------------------------------------------------------
#' Find index and 30-day readmissions function in scr/functions.R
hospital_readmission_rec <- hospital_admission_rec %>%
  # Filtering out birth, planned and outpatient admissions from derivation
  filter(birth_admission == 0 &
           planned_admission == 0 &
           outpatient_admission == 0
         ) %>%
  dplyr::select(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  #' The "find_readmissions" function identifies and flags all admissions that
  #' occur within 30 days after an index admission
  find_readmissions(
    grouping = "patient_uuid",
    start_date = "admission_age_days_firstDate",
    end_date = "admission_age_days_lastDate"
  ) %>%
  arrange(patient_uuid, admission_age_days_firstDate)


# 2) Count of 30-day UHRs per index admission --------------------------
#' Generate df with final count of 30-day UHR (count readmissions)
#' per index admission with time under exposure (exptime) to use as offset
hospital_readmission_counts_rec <- hospital_readmission_rec %>%
  # Bring in deceased patient information
  left_join(deceased_patients, by = "patient_uuid") %>%
  # Bring last record age days to calculate last exposure time at risk
  left_join(
    rare_disease_followup_durations_days %>% dplyr::select(
      patient_uuid,
      last_record_age_days
    ), by = "patient_uuid"
  ) %>%
  #' Exclude admission records in which the patient died during during the
  #' admission period, but keeping previous admissions.
  #' Note: The 2 deceased patients did not died at the hospital
  filter(!(patient_uuid %in% deceased_patients &
    primary_diagnosis_age_days >= admission_age_days_firstDate &
    primary_diagnosis_age_days <= admission_age_days_lastDate)) %>%
  # order by patient and admission date
  arrange(patient_uuid, admission_age_days_firstDate) %>%
  # identify 30-day readmissions for each index admission
  group_by(patient_uuid) %>%
  mutate(
    
    # flag index admissions based on readmission variable
    index_admission = if_else(readmission == "index", 1, 0),

    # create a sequential number for each index admission within patient
    index_number = cumsum(index_admission),

    # flag last index admission for each patient for exptime derivation
    last_index = if_else(index_number == max(index_number), 1, 0),

    # flag 30-day readmissions
    is_uhr = str_detect(readmission, "30_days_uhr")
  ) %>%
  # Group by patient and index number
  group_by(patient_uuid, index_number) %>%
  mutate(
    # count readmissions for each index
    count_readmissions = sum(is_uhr, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  #  Keep only index admissions
  filter(index_admission == 1) %>%
  #' Define exposure time (exptime) which follows each index admission:
  #' Use 30 days as default unless censored earlier for last index admission
  #' Note: 6 index had a offset different to 30 days (2 of these with 0 days)
  mutate(
    exptime = as.integer(
      case_when(
        last_index == 1 &
          #' if diff between last record and admission last day is less than 30, 
          #' put the difference as exptime
          (last_record_age_days - admission_age_days_lastDate < 30) ~
          last_record_age_days - admission_age_days_lastDate,
        # otherwise, default exptime is 30 days by definition
        TRUE ~ 30
      )
    )
  ) %>%
  #' Exclude exp time = 0 (i.e., last record == admission last day),
  #' as there is no time to follow up and count uhr after last index admission
  #' Note: 2 index admissions with this condition
  filter(exptime > 0) %>%
  # order, rename and select variables of interest
  dplyr::select(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    count_readmissions,
    exptime
  )
```

## Technological assistance variable derivation

```{r}
#| label: tech assistance during hospitalizations
#| results: false

#' Definition:
#' The presence of technological assistance prior to hospital admissions
#'  will be computed from the “Therapeutic procedures” and “Hospital admission”
#'   datasets. It will be derived as a binary variable for each hospital
#'    admission record, defined as “yes” if the patient received at least one 
#'    procedure identified as technological assistance on or before the start 
#'    of the hospital admission (variable "admission_age_days firstDate"), 
#'    and “no” otherwise.  
#'    
#' Technological assistance procedures will be identified based on the 
#' procedure variable, using the predefined terms: “gastrostomy”, 
#' “tracheostomy”, “ventricular shunt”, “indwelling”, “nasogastric”,
#'  “orogastric”, “jeujostomy” and “gastrojejunostomy”, as defined
#'   by Harini (2023). The earliest recorded date of any technological 
#'   assistance procedure for each patient will be identified using the 
#'   variables ‘procedure_age_days_firstDate’ or ‘procedure_age_days_date’, 
#'   if available 

# List of terms used to detect technological assistance procedures
#' Note: orogastric, jejunostomy and gastrojejunostomy terms were added to the 
#' prior definition to get related events after reviewing events with 
#' pharmaco-epidemiologist
#' Note 2: indwelling catheter doens't have an equivalent in RWD patients.

TECH_ASSISTANCE_TERMS <- c(
  "gastrostomy", # includes "Insertion of gastrostomy tube", among others.
  "tracheostomy", # includes "Tracheostomy"
  "ventriculoperitoneal shunt", # includes "Revision of ventriculoperitoneal shunt"
  "nasogastric", # includes "Insertion of nasogastric tube"
  "orogastric", # includes "Insertion of orogastric tube",
  "jejunostomy", # includes "Insertion of jejunostomy tube", among others
  "gastrojejunostomy" # includes "fluoroscopy guided gastrojejunostomy tube procedure"
)

# create df to get the first tech procedure for each patient
first_tech_assistance_proc <- rare_disease_therapeutic_procedures %>%
  # filter procedures which matches with defined terms
  filter(
    str_detect(
      procedure,
      regex(paste(TECH_ASSISTANCE_TERMS, collapse = "|"), ignore_case = TRUE)
    )
  ) %>%
  # extract one date for each record in case procedure_age_days_date is NA
  mutate(
    procedure_date = coalesce(
      procedure_age_days_date, 
      procedure_age_days_firstDate)
  ) %>%
  arrange(patient_uuid, procedure_date) %>%
  group_by(patient_uuid) %>%
  #' Select the row(s) with the earliest date for each patient.
  slice_min(procedure_date, with_ties = FALSE) %>%
  summarise(
    # earliest data with a technological assistance
    # no time window after the date of the procedure (immediate effect)
    first_event_date = procedure_date,
    # name of the earliest technological assistance to check device
    name_first_proc = procedure
  ) %>%
  ungroup()

# match hospitalization records with first_tech_assistance dates
first_tech_assistance_proc_rec <- hospital_admission_rec %>%
  left_join(
    first_tech_assistance_proc,
    by = "patient_uuid"
  ) %>%
  #' create tech assistance binary variable Presence "1" or absence "0"
  #' indicating if before or at hospitalization event, the patient had at least
  #' 1 or more tech assistance procedures on or before the admission day.
  mutate(
    tech_assistance_ever = factor( 
      (!is.na(first_event_date) &
         admission_age_days_firstDate >= first_event_date) * 1,
      levels = c(0, 1)
  )) %>%
  dplyr::select(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    tech_assistance_ever
  )
```

## Vagus Nerve Stimulation variable derivation

```{r}
#| label: vagal nerve stimulation insertion 
#| results: false

#' Definition:
#' The presence of VNS prior to hospital admissions will be computed from 
#' the “Therapeutic procedures” and “Hospital admission” datasets. It will be
#'  derived as a binary variable for each hospital admission record, defined 
#'  as “yes” if the patient received at least one procedure identified as VNS
#'   insertion on or before the start of the hospital admission (variable
#'    "admission_age_days firstDate"), and “no” otherwise.  
#'  VNS insertion procedures will be identified from records in the ‘procedure’
#'   variable with the term “vagal nerve stimulator”. The earliest recorded 
#'   date of VNS insertion for each patient will be identified using the 
#'   variables ‘procedure_age_days_firstDate’ or ‘procedure_age_days_date’,
#'    if available. Additionally, the effective period of a VNS insertion will
#'     be assumed to begin 14 days after the insertion date for each patient, 
#'     in line with the device activation period.

VNS_INSERTION <- "vagal nerve stimulator" # includes insertion and replacement

# create df to get the first tech procedure for each patient
first_vns_insertion_proc <- rare_disease_therapeutic_procedures %>%
  # filter only procedures with VNS based on the agreed term
  filter(
    str_detect(
      procedure,
      regex(paste(VNS_INSERTION, collapse = "|"), ignore_case = TRUE)
    )
  ) %>%
  # Create unified date column in case procedure_age_days_date is NA
  mutate(
    procedure_date = coalesce(
      procedure_age_days_date, 
      procedure_age_days_firstDate) 
  ) %>%
  arrange(patient_uuid, procedure_date) %>%
  group_by(patient_uuid) %>%
  # Select the row(s) with the earliest date for each patient
  slice_min(procedure_date, with_ties = FALSE) %>%
  summarise(
    # first event data with assummed 14 days of effect after insertion
    first_event_date = procedure_date + 14,
    # name of the procedure
    name_first_proc = procedure,
    .groups = "drop"
  )

# match hospitalization records with vns dates
first_vns_insertion_proc_rec <- hospital_admission_rec %>%
  left_join(
    first_vns_insertion_proc,
    by = "patient_uuid"
  ) %>%
  #' create vns_insertion binary variable Presence "1" or absence "0"
  #' indicating if for each hospitalization event, the patient had at least 
  #' 1 vns_insertion procedures on or before the admission day. 
  mutate(
    vns_insertion_ever = factor(
      (!is.na(first_event_date) &
         admission_age_days_firstDate >= first_event_date) * 1
      )) %>%
  dplyr::select(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    vns_insertion_ever
  )
```

## Ketogenic diet variable derivation

```{r}
#| label: Ketogenic diet during hospitalizations
#| results: false
#| warning: false
#| message: false

#' Definition:
#' The use of ketogenic diet treatment during hospital admissions will be 
#' computed from the “Therapeutic procedures” and “Hospital admission” datasets.
#'  It will be derived as a binary variable for each hospital admission record,
#'   defined as “yes” if the period of a ketogenic diet procedure (defined by
#'    the variables ‘procedure_age_days_firstDate’ and 
#'    ‘procedure_age_days_lastDate’) coincides with at least one day of the 
#'    admission period, and “no” otherwise. Only procedures 
#'    recorded as “Ketogenic diet” in the ‘procedure’ variable will be 
#'    considered. Additionally, the effective period of a ketogenic diet
#'     procedure will be assumed to start 14 days after the start date of the 
#'     ketogenic diet for each patient (Kosssof et al 2007). 

KETOGENIC_DIET <- "Ketogenic diet" 

ketogenic_diet_episodes <- rare_disease_therapeutic_procedures %>%
# Filter for ketogenic diet procedures only
  filter(procedure %in% KETOGENIC_DIET) %>%
  #' Computing ketogenic diet episodes with 5 days of grace period 
  #' between interventions
  group_by(patient_uuid) %>% 
  mutate(episode_lag = procedure_age_days_firstDate - 
           lag(procedure_age_days_lastDate, default = T), 
         episode_rank = purrr::accumulate(
           abs(episode_lag) > 5, sum, na.rm = T)) %>% 
  # group by patient and episode rank
  group_by(patient_uuid, episode_rank) %>% 
  # #' compute the effective start day (day in which the intervention start to have
  # #' an effect) and the last day based on episodes
  # #' summarize to merge if records with a lag of < 5 days (only one patient)
  summarise(
    keto_effective_first_day= min(procedure_age_days_firstDate) + 14, # 14th day is considered stable 
    keto_effective_last_day = max(procedure_age_days_lastDate),
    .groups = "drop"
  ) %>%
  #' remove episodes with no effective days as the length of the intervention
  #'  was less equal or less than 14 days.
  filter(keto_effective_first_day <= keto_effective_last_day)

#'Create df to get the first ketogenic diet procedure for each patient and
#' hospitalizations in which the patient was under effective keto diet

# a. Derivation of ketogenic diet during admission period for 30-day and lohs analyses
adm_keto_overlap <- hospital_admission_rec %>%
  left_join(
    ketogenic_diet_episodes,
# Check for overlap between *effective* procedure window and admission
  #' any part of the period starting from day 14 to the end of the
  #' ketogenic treatment overlaps with the admission
    join_by(
      patient_uuid,
      overlaps(
        admission_age_days_firstDate,
        admission_age_days_lastDate,
        keto_effective_first_day,
        keto_effective_last_day
      )
    )
  ) %>%
    # Summarise as binary presence indicator per admission
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
    #' Summarise a binary variable indicating was under ketogenic diet
  #'  for each hospital admission record (at least one day of match)
  summarise(
    #' if there is a match between the effective period of the
    #' ketogenic diet and the admission period, then it is an 1
    ketogenic_diet_admission = factor(any(!is.na(keto_effective_first_day)) * 1),
    .groups = "drop"
  )

# b. Derivation of ketogenic diet at admission day for sr analysis
adm_keto_on_admday <- hospital_admission_rec %>%
  left_join(
    ketogenic_diet_episodes,
    # Check for overlap between *effective* procedure window and admission
  #' any part of the period starting from day 14 to the end of the
  #' ketogenic treatment overlaps with the admission
    join_by(
      patient_uuid,
      between(admission_age_days_firstDate,
              keto_effective_first_day,
              keto_effective_last_day)
    )
  ) %>%
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  #' Summarise a binary variable indicating was under ketogenic diet
  #'  for each hospital admission record (if there is a match at the day of admission)
  summarise(
    #' if the admission day is within the effective period, yes = 1, otherwise 0
    ketogenic_diet_at_admission_day = factor(any(!is.na(keto_effective_first_day)) * 1),
    .groups = "drop"
  )

ketogenic_during_admission_rec <- adm_keto_overlap %>%
  left_join(
    adm_keto_on_admday,
    by = c(
      "patient_uuid",
      "admission_age_days_firstDate",
      "admission_age_days_lastDate"
    )
  )
```

## Respiratory support during admissions derivation

```{r}
#| label: Respiratory support during admission - derivation
#| results: false

#' Definition:
#' Respiratory support procedures performed during hospital admissions will 
#' be computed from the “Therapeutic Procedures”, “Clinical Diagnosis” and 
#' “Hospital admission” datasets. It will include invasive and non-invasive 
#' respiratory support procedures to maintain or restore respiratory function.
#'  It will be derived for hospital admission record as a binary variable 
#'  defined as “yes” in the presence of at least one respiratory support event 
#'  and “no” in its absence during a corresponding admission. 
#'  
#'  The following rules will be applied: 
#'  
#'  *From clinical diagnosis records:  
#'  A clinical diagnosis event will be assessed using the period defined by
#'   the variables ‘clinical_diagnosis_age_days_firstDate’ and 
#'   ‘clinical_diagnosis_age_days_lastDate’. The event will be considered 
#'   to have occurred during the admission period if the clinical diagnosis
#'    period coincides with at least one day within the hospital admission. 
#'  
#'  *From therapeutic procedures records:  
#'  If the variable procedure_age_days_date is available, the procedure 
#'  will be considered to have occurred if the procedure date falls within the
#'   hospital admission period. 
#'  If the variable procedure_age_days_date is missing, the procedure will be 
#'  considered to have occurred if its duration, defined by 
#'  procedure_age_days_firstDate (start of the procedure) and 
#'  procedure_age_days_lastDate (end of the procedure), 
#'  coincides with at least one day of the hospital admission period. 
#'  
#'  Therapeutic procedures recorded on the same date as a clinical diagnosis 
#'  will be excluded. The remaining therapeutic procedures and clinical diagnosis
#'   records will then be evaluated to identify respiratory support events. 
#'   This will be based on the ‘procedure’ and ‘clinical_diagnosis’ variables
#'    by searching for the terms: “ventil” (excluding “acute hypoventilation”),
#'     “oxygen”, “tracheostomy” or “tracheotomy”

#### 1. identify therapeutic procedures during admissions
therapeutic_rec <- hospital_admission_rec %>%
  # join therapeutic procedures with hospital admissions
  left_join(
    rare_disease_therapeutic_procedures,
    relationship = "many-to-many",
    by = "patient_uuid"
  ) %>%
  # compute overlap flags (either a date or an interval)
  mutate(
    procedure_overlap = if_else(
      !is.na(procedure_age_days_firstDate) &
      !is.na(procedure_age_days_lastDate) &
      !is.na(admission_age_days_firstDate) &
      !is.na(admission_age_days_lastDate) &
      (admission_age_days_firstDate <= procedure_age_days_lastDate) &
      (admission_age_days_lastDate >= procedure_age_days_firstDate),
      TRUE, FALSE
    ),
    # keep procedure name only if overlap
    procedure = if_else(procedure_overlap, procedure, NA_character_)
  ) %>%
  # group by admission
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  # summarise overlapping procedures (or NA if none)
  summarise(
    ther_procedure_name = str_c(
      unique(na.omit(procedure)), collapse = "; "
    ),
    .groups = "drop"
  ) %>%
  # rename to procedure for binding rows with clinical diagnoses
  rename(procedure = ther_procedure_name) %>%
  arrange(
    patient_uuid,
    admission_age_days_firstDate)


#### 2. idenfity clinical diagnoses during admissions
#' Note: it was tested first the derivation of clinc. diag by episodes
#' but there is no change in the final output as the affected records
#' are not resp support. Thus it was kept the simplest derivation
clinical_diagnosis_rec <- hospital_admission_rec %>%
  # Joining clinical diagnosis with hospital admission by patient UUID
  left_join(
    rare_disease_clinical_diagnosis,
    relationship = "many-to-many",
    by = "patient_uuid"
  ) %>%
  mutate(
    #' Boolean to determine if the clinical diagnosis (procedure) has an
    #' explicit period that matches the admission period in at least one day
    procedure_overlap = if_else(
      !is.na(clinical_diagnosis_age_days_firstDate) &
        !is.na(clinical_diagnosis_age_days_lastDate) &
        !is.na(admission_age_days_firstDate) &
        !is.na(admission_age_days_lastDate) &
        (admission_age_days_firstDate <= clinical_diagnosis_age_days_lastDate) &
        (admission_age_days_lastDate >= clinical_diagnosis_age_days_firstDate),
      TRUE, FALSE
    ),
      clinical_diagnosis = if_else(
         procedure_overlap, clinical_diagnosis, NA_character_
    )
  ) %>%
  #' Filter to retain procedures at hospital admissions:
  #' either a single-day procedure falls within admission period
  #' or the procedure period overlaps with the admission period
  #filter(procedure_overlap) %>%
  # Group by patient and admission periods
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  # Concatenate distinct values into a single cell as temporary variable
  summarise(
    clinical_diagnosis = str_c(
      unique(na.omit(clinical_diagnosis)), collapse = "; "
    ),
    .groups = "drop"
  ) %>%
  # rename to procedure for binding rows with therapeutic procedures
  rename(procedure = clinical_diagnosis) %>%
  arrange(patient_uuid, admission_age_days_firstDate)

#### 3. identify resp.support during admissions
# Define respiratory support terms defined in SAP
RESP_SUPPORT_TERMS <- c("ventil", "oxygen", "tracheostomy", "tracheotomy")

resp_ther_proc_rec <- rbind(
    # bind therapeutic procedures and clinical diagnoses
    clinical_diagnosis_rec,
    therapeutic_rec
  ) %>%
  # add a flag if procedure matches respiratory support terms
  mutate(
    respiratory_support_flag= case_when(
      str_detect(
        procedure,
        regex(str_c(RESP_SUPPORT_TERMS, collapse = "|"), ignore_case = TRUE)
      ) &
      # exclude hypo ventilation term from the list after agreement with PE
      !str_detect(
        procedure,
        regex("hypoventilation", ignore_case = TRUE)
      ) ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>%
  # group by admission
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  # summaries as binary factor: "1" if any respiratory support, else "0"
  summarise(
   respiratory_support_admission = factor(
      any(respiratory_support_flag, na.rm = TRUE) * 1,
       levels = c(0, 1)
    ),
  .groups = "drop"
  ) %>%
  arrange(patient_uuid, admission_age_days_firstDate)
```

## Neonatal onset derivation

```{r}
#| label: Age at disease onset - first seizure record
#| results: false
#| warning: false
#| message: false

#' Definition:
#' Derived as categorical variable to classify the number of patients whose
#'  disease onset was recorded across in the following time ranges: 
#'  0 to ≤ 28 days of life (i.e., neonatal onset) 
#'  and >  28 days of life (i.e., non-neonatal onset)

neonatal_onset_rec <- rare_disease_seizure_history %>% 
  group_by(patient_uuid) %>% 
  # # get the first seizure record for each patient
  slice_min(
    seizure_history_age_days
  ) %>%
  # remove duplicates based on patient UUID
  distinct(patient_uuid, .keep_all = TRUE) %>%
  ungroup() %>% 
  mutate(
    # continuous variable in days as temporary variable
    age_disease_onset_days = seizure_history_age_days,
    #' 28 days (including 28) as described by CDC and agreed with Pharm.epi
    #' binary variable, being 1 if neonatal, otherwise 0.
    neonatal_onset = factor((age_disease_onset_days <= 28) * 1)
  ) %>%
  dplyr::select(
    patient_uuid,
    age_disease_onset_days,
    neonatal_onset
  )
```

## Number of prescribed ASM at hospital admission derivation

```{r}
#| label: Antiseizure medications during hospitalizations
#| results: false
#| warning: false
#| message: false

#' Definition: 
#' The number of ASM received at hospital admission will be computed from the
#'  “Medication” and “Hospital admission" datasets. It will be derived as a
#'   quantitative variable and defined for each hospital admission record, as 
#'   the number of distinct ASMs received (variable “medication”) at the day of
#'   hospital admission (variable "admission_age_days firstDate"). 
#'   
#'  ASM records with a continuous start and end date for the same drug will
#'   be grouped into episodes. An episode ended if a gap of two or more day
#'    s between records was detected. ASMs will be identified from the list 
#'    obtained from: A Summary of Antiseizure Medications Available in
#'     the United States - 4th Edition (2024)

# 1. create list of antiseizure medications
medication_asm_df <- rare_disease_medication %>%
  filter(medication %in% c(
    "Adrenocorticotropin (ACTH 1-18),I-125 (TYR)",
    "Brivaracetam",
    "Carbamazepine",
    "Cenobamate",
    "Clobazam",
    "Clonazepam",
    "Diazepam",
    "Divalproex",
    "Eslicarbazepine acetate",
    "Ethosuximide",
    "Epidiolex", # cannabidiol (CBD)
    "Everolimus",
    "Felbamate",
    "Fenfluramine",
    "Fosphenytoin",
    "Gabapentin",
    "Ganaloxone",
    "Lacosamide",
    "Lamotrigine",
    "Levetiracetam",
    "Lorazepam",
    "Midazolam",
    "Oxcarbazepine",
    "Perampanel",
    "Phenobarbital",
    "Phenytoin",
    "Pregabalin",
    "Primidone",
    "Rufinamide",
    "Stiripentol",
    "Tiagabine",
    "Topiramate",
    "Valproate", # Valproic acid
    "Vigabatrin",
    "Zonisamide"
  )) %>%
  # select variables of interest
  dplyr::select(
    patient_uuid,
    medication,
    medication_age_days_firstDate,
    medication_age_days_lastDate
  ) %>%
  # remove duplicate rows (not taking dose into account)
  distinct() %>%
  group_by(
    patient_uuid, 
    medication
    ) %>%
  arrange(
    medication_age_days_firstDate,
    desc(medication_age_days_lastDate)
  )

#' 2 Identify episodes to identify the longest administration windows and 
#' remove the nested ones
medication_asm_df_episodes <- medication_asm_df %>%
  mutate(
    ##' find the largest end date from current row to row 1
    ##' for each pat/medic combination
    max_end = purrr::accumulate(medication_age_days_lastDate, max),
    
    ## difference between current start and the end of the previous row
    date_diff = medication_age_days_firstDate - lag(max_end, default = T),
    
    ##' if date_diff is >=2 (i.e., episod starts more than 1 day after
    ##'  the end of the previous), episod is incremented by 1
    episod = purrr::accumulate(date_diff >= 2, sum, na.rm = TRUE)
  ) %>%
  # Group by to identify distinct administration windows
  group_by(
    patient_uuid, 
    medication,
    episod
    ) %>%
  arrange(
    patient_uuid, 
    medication, 
    medication_age_days_firstDate
    ) %>%
  # summarise with the first and last date of administration for each drug
  summarise(
    medication_age_days_firstDate = min(medication_age_days_firstDate),
    medication_age_days_lastDate = max(medication_age_days_lastDate),
    .groups = "drop"
  )

# 3. Join the processed ASM dataframe with admission records
asm_at_admission_rec <- hospital_admission_rec %>%
  left_join(
    medication_asm_df_episodes,
    by = "patient_uuid",
    relationship = "many-to-many"
  ) %>%
  # Create flag for ASM overlap only at the day of the admission start 
  mutate(
    medication_overlap = 
      !is.na(medication_age_days_firstDate) &
      !is.na(medication_age_days_lastDate) &
      admission_age_days_firstDate >= medication_age_days_firstDate &
      admission_age_days_firstDate <= medication_age_days_lastDate
  ) %>%
  # Only keep medications that were present at day of admission, if not, NA
  mutate(
    medication = if_else(medication_overlap, medication, NA_character_)
  ) %>%
  group_by(
    patient_uuid,
    admission_age_days_firstDate, 
    admission_age_days_lastDate
    ) %>%
  summarise(
    # main variable to count distinct medications at day of admission
    n_distinct_asm_admission = n_distinct(medication, na.rm = TRUE),
    # temporary variable to check medications
    medication_name = str_c(unique(na.omit(medication)), collapse = "; "),
    # categorical variable created for potential add-hoc analyses
    current_asm_at_admission = factor(
      case_when(
        n_distinct_asm_admission == 0 ~ "0",
        n_distinct_asm_admission == 1 ~ "1",
        n_distinct_asm_admission == 2 ~ "2",
        n_distinct_asm_admission >= 3 ~ ">=3",
      ),
      levels = c("0", "1", "2", ">=3"),
    ),
    .groups = "drop"
  )

#' Calculate number of distinct ASM before first admission for SR analysis 
#' at first hospital admission
# count number of distinct ASM received before first admission
asm_before_first_admission <- hospital_admission_rec %>%
  arrange(patient_uuid, admission_age_days_firstDate) %>%
  # ensure to remove birth, planned and outpatient admissions
  filter(
    birth_admission == 0,
    planned_admission == 0,
    outpatient_admission == 0
  ) %>%
  group_by(patient_uuid) %>%
  # keep only first admission per patient
  slice_head(n = 1) %>% 
  ungroup() %>%
  # join with medication episodes
  left_join(
    medication_asm_df_episodes,
    by = "patient_uuid",
    relationship = "many-to-many"
  ) %>%
  # join first record age days to define pre-admission period
  left_join(
    rare_disease_followup_durations_days %>% dplyr::select(
      patient_uuid,
      first_record_age_days
    ),
    by = "patient_uuid",
    relationship = "many-to-many"
  ) %>%
  distinct() %>%
  mutate(
    pre_admission_end = admission_age_days_firstDate - 1,
    overlap_before_first_admission =
      !is.na(first_record_age_days) &
      !is.na(pre_admission_end) &
      !is.na(medication_age_days_firstDate) &
      !is.na(medication_age_days_lastDate) &
      medication_age_days_lastDate >= first_record_age_days &
      medication_age_days_firstDate <= pre_admission_end,
    medication_before_first_admission = if_else(overlap_before_first_admission, medication, NA_character_)
  ) %>%
    group_by(
    patient_uuid,
    admission_age_days_firstDate, 
    admission_age_days_lastDate
    ) %>%
  summarise(
    n_distinct_asm_start_fu = n_distinct(medication_before_first_admission, na.rm = TRUE),
    meds_before_first_admission = str_c(sort(unique(na.omit(medication_before_first_admission))), collapse = "; "),
    .groups = "drop"
  )
```

## Quinidine treatment derivation

```{r}
#| label: Quinidine use during hospitalizations
#| results: false
#| warning: false
#| message: false

#' Definition:
#' Quinidine treatment effective during hospital admission 
#' The use of quinidine during hospital admissions will be computed from the
#'  “Medication” and Hospital admission” datasets. It will be derived as a 
#'  binary variable for each hospital admission record, defined as “yes” if
#'   the period quinidine medication (defined by the (variables
#'    ‘medication_age_days_firstDate’ and ‘medication_age_days_lastDate’) 
#'    coincides with at least one day of the hospital admission period 
#'    ,and “no” otherwise.  
#'
#' Quinidine records with a continuous start and end date were grouped into
#'  episodes. An episode ended if a gap of two or more days between records 
#'  was detected. ‘medication_age_days_firstDate’‘medication_age_days_firstDate’

QUINIDINE <- c("Quinidine", "quinidine")

#1. detecting and merging episodes of quinidine use
quinidine_use_episodes <- rare_disease_medication %>%
  # filter quinidine administration only
  filter(medication %in% QUINIDINE) %>%
  dplyr::select(
    patient_uuid,
    medication,
    medication_age_days_firstDate,
    medication_age_days_lastDate
  ) %>%
  ## remove duplicate rows (not taking dose into account for this analysis)
  distinct() %>%
  group_by(patient_uuid, medication) %>%
  arrange(
    medication_age_days_firstDate,
    desc(medication_age_days_lastDate)
  ) %>%
  #' Identify episodes to identify the longest administration windows
  #' and remove the nested ones
  mutate(
    ##' find the largest end date from current row to row 1
    ##' for each pat/medic combination
    max_end = purrr::accumulate(medication_age_days_lastDate, max),
    
    ## difference between current start and the end of the previous row
    date_diff = medication_age_days_firstDate - lag(max_end, default = T),
    
    ##' keep all records, just merged continuous (without gap) one 
    ##' and remove nested 
    episod = purrr::accumulate(date_diff > 0, sum, na.rm = TRUE)
  ) %>%
  # Identify distinct prescription windows (no effective windows)
  arrange(patient_uuid, episod, medication_age_days_firstDate) %>%
  # Group by patient and episode
  group_by(patient_uuid, episod) %>%
  summarise(
    medication_age_days_firstDate = min(medication_age_days_firstDate),
    medication_age_days_lastDate = max(medication_age_days_lastDate),
    .groups = "drop"
  ) %>%
  ungroup()

#2.  identify admissions with matching use of quinidine in at least one day
quinidine_during_admission_rec_any <- hospital_admission_rec %>%
  # Join hospital admissions and quinidine episodes on patient_uuid and overlap of periods
  left_join(
    quinidine_use_episodes,
    join_by(
      patient_uuid,
      overlaps(
        admission_age_days_firstDate,
        admission_age_days_lastDate,
        medication_age_days_firstDate,
        medication_age_days_lastDate
      )
    )
  ) %>%
  # Group by patient and admission periods
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  #' Summarise a binary variable indicating quinidine use for each hospital 
  #' admission record (at least one day of match)
  summarise(
    # if there is a match between the quinidine use and the admission
    # period, then it is an 1, otherwise 0
    quinidine_admission = factor(any(!is.na(medication_age_days_firstDate)) * 1),
    .groups = "drop"
  )

#2.a identify admissions with quinidine already active on the admission day
quinidine_during_admission_rec_on_day <- hospital_admission_rec %>%
  # Join hospital admissions and quinidine episodes where admission start is within the quinidine use period
  left_join(
    quinidine_use_episodes,
    join_by(
      patient_uuid,
      between(
        admission_age_days_firstDate,
        medication_age_days_firstDate,
        medication_age_days_lastDate
      )
    )
  ) %>%
  # Group by patient and admission periods
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  #' Summarise a binary variable indicating quinidine use at the admission day
  summarise(
    quinidine_at_admission_day = factor(any(!is.na(medication_age_days_firstDate)) * 1),
    .groups = "drop"
  )

#2.b merge both variables into the final df (keep original name)
quinidine_during_admission_rec <- quinidine_during_admission_rec_any %>%
  left_join(
    quinidine_during_admission_rec_on_day,
    by = c(
      "patient_uuid",
      "admission_age_days_firstDate",
      "admission_age_days_lastDate"
    )
  )
```

## Adverse effects during admission derivation

```{r}
#| label: Adverse effects during admissions
#| results: false
#| warning: false
#| message: false

#' Definition:
#' The presence of adverse effects for each hospital admission record will be
#'  computed from the “Adverse effects” and “Hospital admission” datasets. It
#'   will be derived for each hospital admission record as a binary variable 
#'   defined as “yes” in the presence of at least one adverse effect of any type
#'    (variable ‘adverse_effect’) and “no” otherwise.  
#'    An adverse effect event will be considered to have occurred if its 
#'    duration, defined by adverse_effect_age_days_firstDate (i.e., start 
#'    of the adverse effect) and adverse_effect_age_days_lastDate (i.e., end of
#'     the adverse effect), coincides with at least one day of the hospital 
#'     admission period 
#'     Note: the variable is called adverse effects and not adverse events**
adverse_effect_during_admission_rec <- hospital_admission_rec %>%
  # Join adverse event episodes
  left_join(
    rare_disease_adverse_effects,
    by = "patient_uuid",
    relationship = "many-to-many"
  ) %>%
  distinct(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    adverse_effect,
    medication_procedure,
    adverse_effect_age_days_firstDate,
    adverse_effect_age_days_lastDate
  ) %>%
  # Boolean overlap: adverse event overlaps with admission
  mutate(
    adverse_effect_overlap = if_else(
      !is.na(adverse_effect_age_days_firstDate) &
        !is.na(adverse_effect_age_days_lastDate) &
        !is.na(admission_age_days_firstDate) &
        !is.na(admission_age_days_lastDate) &
        (adverse_effect_age_days_firstDate <= admission_age_days_lastDate) &
        (adverse_effect_age_days_lastDate >= admission_age_days_firstDate),
      TRUE, FALSE
    )
  ) %>%
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  # Add binary variable (0/1) for adverse event presence during admission
  # Note: for those admissions without AE, it was assumed no event (i.e., 0)
  summarise(
    # below some temporary variables to check AE types & medications
    number_of_adverse_effects = sum(adverse_effect_overlap, na.rm = TRUE),
    adverse_effect_types = if_else(any(adverse_effect_overlap, na.rm = TRUE),
      paste(sort(unique(adverse_effect[adverse_effect_overlap])), 
            collapse = "; "),
      NA_character_
    ),
    medication_procedure = if_else(any(adverse_effect_overlap, na.rm = TRUE),
      paste(sort(unique(medication_procedure[adverse_effect_overlap])), 
            collapse = "; "),
      NA_character_
    ),
    # create the binary variable if at least one AE for each admission
    adverse_effect_admission = factor(
      # if there is a match between the adverse event and the admission
      # period, then it is an 1, otherwise 0
      any(adverse_effect_overlap, na.rm = TRUE) * 1),
    .groups = "drop"
  )
```

## Mutated genes derivation

```{r}
#| label: rare_disease mutated genes
#| results: false
#| warning: false
#| message: false

# list of mutated genes taken from gene4epilepsy
LIST_MUTATED_GENES <- gene4epilepsy_list_202503$Gene

# clean duplicates
mutations <- rare_disease_genetic_findings %>% 
  dplyr::select(
    patient_uuid,
    gene,
    variant_DNA,
    variant_protein,
    interpretation,
    inheritance
    ) %>% 
  group_by(
    patient_uuid,
    gene, 
    variant_DNA,
    variant_protein
    ) %>% 
  mutate(
    interpretation_corr = paste0(interpretation, collapse = ", "), 
    inheritance_corr = paste0(inheritance, collapse = ", ")
  ) %>% 
  # there is a patient with two mutations in rare_disease gen (51 unique)
  distinct(
    patient_uuid,
    gene,
    variant_DNA,
    variant_protein,
    .keep_all = TRUE)

# Step 2 — count rare_disease + epilepsy-related genes
mutations_epilepsy_rec <- mutations %>%
  ungroup() %>%
  group_by(patient_uuid) %>%
  # count mutations in rare_disease and epilepsy-related genes
  summarise(
    rare_disease = any(gene == "rare_disease"),
    n_mutations_epilepsy = n_distinct(
      gene[gene != "rare_disease" & gene %in% LIST_MUTATED_GENES]),
    .groups = "drop"
  ) %>% 
  # only patients with rare_disease
  filter(rare_disease) %>%
  # define binary variable
  mutate(
    # when there rare_disease and another/other genes assign an 1, otherwise 0
    other_epilepsy_mutations = factor((n_mutations_epilepsy > 0) * 1)
  ) %>%
  dplyr::select(patient_uuid, other_epilepsy_mutations)
```

## Highly recurrent variants derivation

```{r}
#| label: rare_disease highly recurrent variants
#| results: false
#| warning: false
#| message: false

# List of recurrent variants which are present in >= 10% of RWD patients
# The variants were confirmed with Borlot et al (2020)
variant_DNA_summary <- mutations %>%
  group_by(variant_DNA) %>%
  summarise(
    # variants per patient
    n_patients = n(),
    .groups = "drop"
  ) %>%
  mutate(
    # proportion from total patients 
    proportion = round(
      n_patients / sum(n_distinct(mutations$patient_uuid)) * 100, 2)
  ) %>%
  arrange(desc(n_patients))

# List of recurrent variants in >= 10% of patients
RECURRENT_VARIANTS <- variant_DNA_summary %>%
  filter(proportion >= 10) %>%
  pull(variant_DNA)

#' Classify patients by presence of at least one recurrent variant using
#'  curated mutation_rec df
recurrent_variant_rec <- mutations %>%
  group_by(patient_uuid) %>%
  summarise(
    # assing an 1 if at least one recurrent variant, otherwise 0
    recurrent_variant = factor(any(variant_DNA %in% RECURRENT_VARIANTS) * 1),
    .groups = "drop"
  ) %>%
  dplyr::select(patient_uuid, recurrent_variant)
```

# Main analyses dataframes

## Dataframe for secondary endpoints

```{r}
#| label: Joining patient level dataframes
#| results: false

# List of dataframes to join by patient_uuid
patient_level_dfs <- list(
  hospital_admission_rec, # start, end day, seizure diagnosis type & lohs vars.
  followup_rec, # first and last day follow up variables
  demographics_rec, # sex variable
  neonatal_onset_rec, # neonatal onset variable
  mutations_epilepsy_rec, # other epilepsy mutations variable
  recurrent_variant_rec # recurrent variant status variable
)

# merge patient-level dataframes by patient_uuid
hospital_joined_df <- purrr::reduce(
  patient_level_dfs,
  left_join, 
  by = "patient_uuid")
```

```{r}
#| label: Joining admission-varying dataframes
#| results: false

admission_join_var <- c(
  "patient_uuid",
  "admission_age_days_firstDate",
  "admission_age_days_lastDate"
)

# List of time-varying covariates to join
time_varying_dfs <- list(
  hospital_readmission_rec, # readmission type variable
  hospital_readmission_counts_rec, # count uhrs and exptime variables
  resp_ther_proc_rec, # respiratory support variable
  first_tech_assistance_proc_rec, # tech assistance variable
  first_vns_insertion_proc_rec, # vns insertion variable
  asm_at_admission_rec, # n_distinct asm variable
  asm_before_first_admission, # n_distinct asm at start of follow-up variable for sr admission at first admission
  adverse_effect_during_admission_rec, # adverse events admission variable
  ketogenic_during_admission_rec, # ketogenic diet admission variable
  quinidine_during_admission_rec # quinidine use admission variable
)

# joining hosptialization-varying covariates
hospital_joined_df <- purrr::reduce(
  time_varying_dfs,
  left_join,
  .init = hospital_joined_df,
  by = admission_join_var
)
```

```{r}
#| label: Final processing of the hospital admission joined main analyses df
#| results: false

# Final processing of the joined dataframe for main analyses
# Note: dataframe that will be used for all secondary endpoint main analyses
hospitalization_main_df <- hospital_joined_df %>%
  mutate(
    patient_uuid = as.factor(patient_uuid)
  ) %>%
  # select and order variables of interest in joined df
  dplyr::select(
    # patient identifier
    patient_uuid,
    # admissions related features
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    birth_admission,
    planned_admission,
    outpatient_admission,
    # demographics
    sex,
    age_admission_yr,
    # follow up and disease onset
    first_record_age_days,
    last_record_age_days,
    neonatal_onset,
    # genetic mutations
    other_epilepsy_mutations,
    recurrent_variant,
    # clinical history
    admission_diagnosis,
    seizure_related_admission,
    length_of_stay_days,
    length_of_stay_days_cat,
    respiratory_support_admission,
    tech_assistance_ever,
    readmission,
    count_readmissions,
    exptime,
    # treatment history
    vns_insertion_ever,
    n_distinct_asm_admission,
    n_distinct_asm_start_fu, # last variable added for SR at first admission
    adverse_effect_admission,
    ketogenic_diet_admission,
    ketogenic_diet_at_admission_day, # added for sr analysis
    quinidine_admission,
    quinidine_at_admission_day # added for sr analysis
  ) %>%
  # arrange by patient uuid and admission day
  arrange(patient_uuid, admission_age_days_firstDate)
```

```{r}
#| label: save-main-df-file-GCS
#| results: false
#| warning: false
#| message: false

# create temp dir to save file
dir.create("~/RWE-rare_disease/Hospitalizations/output/data")

# save df as rds file
saveRDS(hospitalization_main_df, "data/hospitalization_main_df.rds")

# Upload final df to GCS
gcs_upload("~/RWE-rare_disease/Hospitalizations/output/data/hospitalization_main_df.rds",
  bucket = "YOUR_GCS_BUCKET",
  name = "MSM_Hospitalizations/hospitalization_main_df.rds"
)
```

## Dataframe for exploratory endpoint

```{r}
#| label: patient-trajectory-derivation-df
#| echo: false
#| results: false
#| warning: false

# Patient trajectory across hospitalization states #
#' The patient trajectory describes the patient’s health status over time during
#'  follow-up, showing whether they are out of hospital, admitted with a
#'  seizure-related diagnosis, or admitted for another reason at any given
#'  time of follow-up until death or lost to follow-up. It will be defined
#'  as a time-dependent categorical variable that classify each patient into
#'   one of three different states:
#'
# 1. Out-of-hospital state.
# 2. Seizure-related admission state == SR state
# 3. Non-seizure-related admission state == non SR state


# 1). Define the hospitalization states from initial df (SR and non SR only)
#' Note 1: to start it will be used the curated hospital_admission_rec derived
#'  for all secondary endpoints.
#' Note: 2 in this analysis, all admissions, including birth, planned and
#'  outpatient admissions will be used to capture the full trajectory of patients
#'  Note 3: the time-scale is based on patient's age at admission
#'  start/discharge in days, and then renamed to tstart and tstop.
admissions_initial_states <- hospital_admission_rec %>%
  # bring first and last record to
  left_join(rare_disease_followup_durations_days, by = "patient_uuid") %>%
  mutate(
    state = if_else(seizure_related_admission == 1,
      "SR admission",
      "Non-SR admission"
    ),
    # use patient age in days as main timescale
    age_state_start = admission_age_days_firstDate,
    age_state_end = admission_age_days_lastDate,
    # duration of each state in days without including the end day
    duration_state = age_state_end - age_state_start
  ) %>%
  dplyr::select(
    patient_uuid,
    state,
    age_state_start,
    age_state_end,
    duration_state,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    followup_duration,
    first_record_age_days,
    last_record_age_days
  )


# 2). Merge consecutive admissions of same type into single states
# 2.1 flag next day admissions for same type of states into episodes
#' gap_threshold = 1 → next hospitalization record starts on the next day (+1)
#'  after the previous discharge (for the same patient and with the same type
#'   of sr diagnosis)
#'  Note: there are  9 transitions from SR to SR and 7 non-SR to non-SR in raw data.
admissions_initial_states_flagged <- flag_admission_episodes(
  admissions_initial_states,
  gap_threshold = 1 # next day
)

# 2.2 merge flagged episodes into single states to remove transitions between same states
admissions_initial_states_merged <- admissions_initial_states_flagged %>%
  group_by(
    patient_uuid, # by patient
    episode_id, # by episode identified earlier
    state # and by seizure-related or non-seizure-related state
  ) %>%
  summarise(
    age_state_start = min(age_state_start),
    age_state_end = max(age_state_end),
    duration_state = age_state_end - age_state_start,
    followup_duration = first(followup_duration),
    first_record_age_days = first(first_record_age_days),
    last_record_age_days = first(last_record_age_days),
    .groups = "drop"
  ) %>%
  dplyr::select(-episode_id)

# 3) Derive out of hospital (OOH) state
# 3.1. Derive out of hospital (OOH) gaps between admissions
admissions_states_out_hospital <- admissions_initial_states_merged %>%
  arrange(patient_uuid, age_state_start) %>%
  group_by(patient_uuid) %>%
  mutate(
    prev_end = lag(age_state_end),
    next_start = lead(age_state_start),
    # OOH starts at the day after previous discharge or first record age days
    ooh_start = case_when(
      is.na(prev_end) & age_state_start > first_record_age_days ~ first_record_age_days,
      prev_end < age_state_start - 1 ~ prev_end
    ),
    # OOH ends the same day of the next admission start day
    ooh_end = case_when(
      is.na(prev_end) & age_state_start > first_record_age_days ~ age_state_start,
      prev_end < age_state_start - 1 ~ age_state_start,
      is.na(next_start) ~ NA_real_
    )
  ) %>%
  ungroup() %>%
  # remove invalid OOH intervals
  filter(!is.na(ooh_start), ooh_end >= ooh_start) %>%
  transmute(
    patient_uuid,
    state = "Out-of-hospital",
    age_state_start = ooh_start, # age in days
    age_state_end = ooh_end, # age in days
    duration_state = ooh_end - ooh_start,
    followup_duration,
    first_record_age_days,
    last_record_age_days
  )

# 3.2. Derive OOH state for patients under follow-up but NO admissions
admissions_states_out_hospital_none <- rare_disease_followup_durations_days %>%
  dplyr::anti_join(
    admissions_states_out_hospital %>% dplyr::distinct(patient_uuid),
    by = "patient_uuid"
  ) %>%
  transmute(
    patient_uuid,
    state = "Out-of-hospital",
    # OOH starts at first record age days
    age_state_start = first_record_age_days,
    # OOH ends at last record age days
    age_state_end = last_record_age_days,
    duration_state = age_state_end - age_state_start,
    followup_duration,
    first_record_age_days,
    last_record_age_days
  )

# 4) Merge admissions (sr and non-sr) with out-of-hospital states
all_states_full <- bind_rows(
  admissions_initial_states_merged,
  admissions_states_out_hospital,
  admissions_states_out_hospital_none
) %>%
  arrange(patient_uuid, age_state_start) %>%
  dplyr::select(-c(
    followup_duration,
    first_record_age_days
  ))

#' 5) Extend last state until the last medical record (if needed, as 4 patients
#' had the last medical record with a discharge day for a hospitalization)
all_states_full_fu <- all_states_full %>%
  group_by(patient_uuid) %>%
  arrange(age_state_end, age_state_start, .by_group = TRUE) %>% # sort within patient
  slice_tail(n = 1) %>% # keep last row per patient
  ungroup() %>%
  # extend only if observation ends after the last state
  filter(last_record_age_days > age_state_end) %>%
  group_by(patient_uuid) %>%
  mutate(
    out_hospital_start = age_state_end,
    out_hospital_end   = last_record_age_days
  ) %>%
  transmute(
    patient_uuid,
    state = "Out-of-hospital",
    age_state_start = out_hospital_start, # age in days
    age_state_end = out_hospital_end, # age in days
    duration_state = age_state_end - age_state_start
  )

# 6) Final merge of all states including observations until last medical record.
all_states_full_fu_joined <- bind_rows(all_states_full, all_states_full_fu) %>%
  arrange(patient_uuid, age_state_start) %>%
  dplyr::select(-last_record_age_days)
```

```{r}
#| label: msm-ag-model-covariates
#| echo: true
#| warning: false
#| results: false

# Add covariates to the one row per transition dataframe
# Preparing dataframe for MSM AG model with covariates
#' this process is similar to the one proposed by expand.covs from mstate package
#' for creating transition specific covariates
#'  https://cran.r-project.org/web/packages/mstate/vignettes/Tutorial.pdf

# 1) Add fixed time covariates 
# 1.1. Sex variable, as sex male = 1
msm_ag_model_sex <- all_states_full_fu_joined %>%
  left_join(
    # df defined for main analysis
    demographics_rec %>% distinct(patient_uuid, .keep_all = TRUE) %>%
      dplyr::select(
        patient_uuid,
        sex),
    by = "patient_uuid"
  ) %>%
  group_by(
    patient_uuid
  ) %>%
  mutate(
    sex_male = as.numeric(sex == "male")
  ) %>%
  dplyr::select(-sex)


# 1.2. add neonatal onset status, as neonatal onset yes = 1
msm_ag_model_neo_onset <- all_states_full_fu_joined %>%
  left_join(
    # df defined for main analysis
    neonatal_onset_rec %>% distinct(patient_uuid, .keep_all = TRUE) %>%
      dplyr::select(
        patient_uuid,
        neonatal_onset),
    by = "patient_uuid"
  ) %>%
  group_by(
    patient_uuid
  ) %>%
  mutate(
    neonatal_onset_yes = as.numeric(neonatal_onset == 1)
  ) %>%
  dplyr::select(-neonatal_onset, -state, -duration_state)

# 2) Add time-varying covariates
# 2.1 join quinidine treatment active on the transition day
msm_ag_model_quinidine <- all_states_full_fu_joined %>%
  #' Join derived df and quinidine episodes where the transition 
  #' start is within the quinidine use period
  left_join(
    # df defined for main analysis
    quinidine_use_episodes,
    join_by(
      patient_uuid,
      between(
        age_state_start,
        medication_age_days_firstDate,
        medication_age_days_lastDate
      )
    )
  ) %>%
  # Group by patient and admission periods
  group_by(
    patient_uuid,
    age_state_start,
    age_state_end
  ) %>%
  #' Summarise as binary numerical variable indicating quinidine treatment
  #'  at transition day
  summarise(
    quinidine_at_transition_day = 
      as.numeric(any(!is.na(medication_age_days_firstDate)) * 1),
      .groups = "drop"
  ) 

# 2.2. join number of distinct ASMs at transition day
msm_ag_model_asm <- all_states_full_fu_joined %>%
  #' Join  derived df and asm episodes where the transition 
  #' start is within the medication use period
  left_join(
    # df defined for main analysis
    medication_asm_df_episodes,
    join_by(
      patient_uuid,
      between(
        age_state_start,
        medication_age_days_firstDate,
        medication_age_days_lastDate
      )
    )
  ) %>%
  group_by(
    patient_uuid,
    age_state_start,
    age_state_end
  ) %>%
  summarise(
    #' count number of distinct ASM at transition day
    n_distinct_asm_at_transition_day = n_distinct(medication, na.rm = TRUE),
    .groups = "drop"
  )

# join ketogenic diet episodes to identify if under ketogenic diet at transition day
msm_ag_model_keto <- all_states_full_fu_joined %>%
  #' Join  derived df and ketogenic diet episodes where the transition 
  #' start is within the ketogenic diet effective intervention period
  left_join(
    # df defined for main analysis
    ketogenic_diet_episodes,
    # Check for overlap between *effective* procedure window and admission
    #' any part of the period starting from day 14 to the end of the
    #' ketogenic treatment overlaps with the admission
    join_by(
      patient_uuid,
      between(
        age_state_start,
        keto_effective_first_day,
        keto_effective_last_day)
    )
  ) %>%
  group_by(
    patient_uuid,
    age_state_start,
    age_state_end
  ) %>%
  #' Summarise a binary variable indicating was under ketogenic diet
  #'  for each hospital admission record (if there is a match at the day of admission)
  summarise(
    #' if the admission day is within the effective period, yes = 1, otherwise 0
    ketogenic_diet_at_transition_day = 
      as.numeric(any(!is.na(keto_effective_first_day)) * 1),
    .groups = "drop"
  )

# join VNS insertion ever before or on transition day
msm_ag_model_vns <- all_states_full_fu_joined %>% 
  #' Join derived df and vns insertion for each patient when where the transition 
  #' start occurred after the first VNS insertion.
  left_join(
    # df defined for main analysis
    first_vns_insertion_proc,
    join_by(
      patient_uuid,
      age_state_start >= first_event_date    
    )
  ) %>%
  group_by(
    patient_uuid,
    age_state_start,
    age_state_end
  ) %>%
  summarise(
    #' create VNS binary variable Presence "1" or absence "0" if at or before
    #'  the transition day, the patient had a VNS insertion
    vns_at_transition_ever = as.numeric(any(!is.na(first_event_date))),
    .groups = "drop"
  )

# match hospitalization records with first_tech_assistance dates
msm_ag_model_tech_assistance <- all_states_full_fu_joined %>%
  #' Join derived df and tech assistance first event for each patient where 
  #' the transition start occurred after the first tech assistance
  left_join(
    # df defined for main analysis
    first_tech_assistance_proc,
    join_by(
      patient_uuid,
      age_state_start >= first_event_date    
    )
  ) %>%
  group_by(
    patient_uuid,
    age_state_start,
    age_state_end
  ) %>%
  #' create tech assistance binary variable Presence "1" or absence "0"
  #' indicating if before or at transition day, the patient had at least
  #' 1 or more tech assistance procedures on or before the admission day.
  summarise(
    tech_assistance_at_transition_ever = 
      as.numeric(any(!is.na(first_event_date))),
    .groups = "drop")


# variables to join by
msm_join_var <- c(
  "patient_uuid",
  "age_state_start",
  "age_state_end"
)

# List of time-varying covariates to join
time_varying_msm_dfs <- list(
  msm_ag_model_neo_onset,
  msm_ag_model_quinidine,
  msm_ag_model_asm,
  msm_ag_model_keto,
  msm_ag_model_vns,
  msm_ag_model_tech_assistance
)

# joining hosptialization covariates for MSM AG model
msm_ag_model_cov <- purrr::reduce(
  time_varying_msm_dfs,
  left_join,
  .init = msm_ag_model_sex,
  by = msm_join_var
)
```

```{r}
#| label: msm-ag-model-cov-long-format
#| echo: true
#| warning: false

#' 1. Create the counting process long format dataframe based on:
# https://www.sciencedirect.com/science/article/pii/S0169260703001329 Paes (2004)
# http://link.springer.com/10.1007/978-1-4757-3294-8 Therneau and Grambsch (2000)
# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2692556/ Meira Machado (2009)
# https://onlinelibrary.wiley.com/doi/10.1002/sim.2712 Putter (2007)
msm_ag_model_cov_long <- msm_ag_model_cov %>%
arrange(patient_uuid, age_state_start) %>%
  group_by(patient_uuid) %>%
  mutate(
    next_state = lead(state),
    from_num = case_when(
      state == "Out-of-hospital"   ~ 1L,
      state == "SR admission"      ~ 2L,
      state == "Non-SR admission"  ~ 3L
    ),
    to_num = case_when(
      lead(state) == "Out-of-hospital"   ~ 1L,
      lead(state) == "SR admission"      ~ 2L,
      lead(state) == "Non-SR admission"  ~ 3L,
      TRUE ~ NA_integer_
    ),
    tstart = age_state_start,
    # check if last record per patient to avoid NA in tstop  +1
    tstop  = age_state_end
  ) %>%
  ungroup()

# 2. Specify allowed transitions and their numeric codes for inner join
allowed <- tibble::tribble(
  ~from, ~to, ~transition,
  1L,    2L, 1,   # 1->2
  1L,    3L, 2,   # 1->3
  2L,    1L, 3,   # 2->1
  3L,    1L, 4,   # 3->1
  2L,    3L, 5,   # 2->3 
  3L,    2L, 6    # 3->2 
)

# 3. Create final long format df for MSM AG model with covariates
#' it will add an extra row per allowed transition per time interval to have
#'  the time under risk for the other transitions
msm_ag_model_cov_long_df <- msm_ag_model_cov_long %>%
  inner_join(allowed, 
             by = c("from_num" = "from"), 
             relationship = "many-to-many"
             ) %>%
  mutate(
    # event indicator for the transition of interest
    # If the next state matches the 'to' state of the allowed transition, event = 1.
    #' If not, and there is a next state, event = 0 (no transition occurred,
    #'  right censored at end of follow-up).
    event = as.numeric(!is.na(to_num) 
                       & to_num == to),
    # label the transitions and specify levels
    transition = factor(transition, 
                        levels = 1:6,
                        labels = c("1-2","1-3","2-1","3-1","2-3","3-2"))
  ) %>%
  # order columns and arrange
  dplyr::select(
    patient_uuid, 
    state,
    transition, 
    tstart, 
    tstop,
    event,
    duration_state,
    sex_male,
    neonatal_onset_yes,
    quinidine_at_transition_day,
    n_distinct_asm_at_transition_day,
    ketogenic_diet_at_transition_day,
    vns_at_transition_ever,
    tech_assistance_at_transition_ever
  ) %>%
  arrange(patient_uuid, tstart, transition)
```

```{r}
#| label: msm-ag-modelling-data-export
#| eval: false
#| echo: false

# Upload to SAS the ptransit dataframe with TWO states (in and out of hospital)
dir.create("~/RWE-rare_disease/Hospitalizations/output/data")

# save df as rds file
saveRDS(msm_ag_model_cov_long_df, "data/msm_ag_model_cov_long_df.rds")

# Upload final df to GCS
gcs_upload("~/RWE-rare_disease/Hospitalizations/output/data/msm_ag_model_cov_long_df.rds",
  bucket = "YOUR_GCS_BUCKET",
  name = "MSM_Hospitalizations/msm_ag_model_cov_long_df.rds"
)
```

# Sensitivity analyses dataframes

### Merging hospitalization records when the admission day occurs on the same or next day as the prior discharge day

```{r}
#| label: no merged df 
#| echo: false
#| results: false

hospitalizations_no_merged <- hospitalization_main_df %>%
  dplyr::select(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    seizure_related_admission,
    planned_admission,
    birth_admission,
    outpatient_admission
  )
```

#### Hospitalization occurs on the SAME day as the prior discharge day

```{r}
#| label: merging hospital admission records same day
#| echo: false
#| results: false

#' gap_threshold = 0 → next hospitalization record starts on the *same day (+0)*
#'  as the previous discharge (for the same patient and with the same type of
#'  sr diagnosis)

# 1. Apply the episode-flagging function across all gap thresholds
#' Note: the function developed is in function.R
hospital_admission_flag_same_day <- flag_admission_episodes(
  hospitalizations_no_merged,
  gap_threshold = 0
) # same day

# 2. Merge admissions within the same episode
#' Internal note: 2 more records could be merged by removing the seizure-type
#' condition for merging
hospitalization_merged_same_day <- hospital_admission_flag_same_day %>%
  # Keep only unplanned, non-birth admissions and inpatient admissions
  filter(
    birth_admission == 0 &
    planned_admission == 0 &
    outpatient_admission == 0) %>%
  group_by(
    patient_uuid, # by patient
    episode_id, # by episode identified earlier
    seizure_related_admission # and by seizure-related flag
  ) %>%
  # merge admissions within patient, same episode and seizure diagnosis type
  summarise(
    # earliest start in episode
    admission_age_days_firstDate = min(admission_age_days_firstDate),
    # latest end in episode
    admission_age_days_lastDate = max(admission_age_days_lastDate),
    .groups = "drop"
  )

# 3. re-run the count of readmissions with new merged records
# function explained in detail in function.R and main derivation above.
readmissions_sensit_same_day <- hospitalization_merged_same_day %>%
  dplyr::select(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  find_readmissions(
    grouping = "patient_uuid",
    start_date = "admission_age_days_firstDate",
    end_date = "admission_age_days_lastDate"
  ) %>%
  arrange(patient_uuid, admission_age_days_firstDate)

# 4. re-derive count of 30-day UHRs per index admission with merged records
readmissions_sensit_same_day_counted <- readmissions_sensit_same_day %>%
  # Bring in deceased patient information
  left_join(deceased_patients, by = "patient_uuid") %>%
  # Bring last record age days to calculate last exposure time at risk
  left_join(
    rare_disease_followup_durations_days %>% dplyr::select(
      patient_uuid,
      last_record_age_days
    ),
    by = "patient_uuid"
  ) %>%
  #' Exclude admission record if a patient is deceased during the
  #' admission period, but retain prior ones for that patient
  filter(!(patient_uuid %in% deceased_patients &
    primary_diagnosis_age_days >= admission_age_days_firstDate &
    primary_diagnosis_age_days <= admission_age_days_lastDate)) %>%
  arrange(patient_uuid, admission_age_days_firstDate) %>%
  group_by(patient_uuid) %>%
  # identify 30-day readmissions for each index admission
  mutate(
    # flag index admissions based on readmission variable
    index_admission = if_else(readmission == "index", 1, 0),

    # create a sequential number for each index admission within patient
    index_number = cumsum(index_admission),

    # flag last index admission for each patient for last exptime derivation
    last_index = if_else(index_number == max(index_number), 1, 0),

    # flag 30-day readmissions
    is_uhr = str_detect(readmission, "30_days_uhr")
  ) %>%
  # Group by patient and index number
  group_by(patient_uuid, index_number) %>%
  mutate(
    # count readmissions for each index
    count_readmissions = sum(is_uhr, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  #  Keep only index admissions
  filter(index_admission == 1) %>%
  #' Define exposure time (exptime) for each index admission
  mutate(
    exptime = as.integer(
      case_when(
        last_index == 1 &
          (last_record_age_days - admission_age_days_lastDate < 30) ~
          last_record_age_days - admission_age_days_lastDate,
        TRUE ~ 30
      )
    )
  ) %>%
  #' Exclude exp time = 0 (i.e., last record == admission last day)
  filter(exptime > 0) %>%
  # order, rename and select variables of interest
  dplyr::select(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    count_readmissions,
    exptime
  )

# 5. Joined dataframe with merged hospitalizations and readmissions
hospitalizations_sensit_same_day_joined <- purrr::reduce(
  list(
    hospitalization_merged_same_day,
    readmissions_sensit_same_day,
    readmissions_sensit_same_day_counted
  ),
  left_join,
  by = c(
    "patient_uuid",
    "admission_age_days_firstDate",
    "admission_age_days_lastDate"
  )
)

#' 6. join previously derived covariates (from main df) that do not change with
#'  the merging and depends only on the patient id and admission start day
#'  (not period dependent)
hospitalizations_sensit_same_day_joined_cov <- 
  hospitalizations_sensit_same_day_joined %>%
  left_join(
    hospitalization_main_df %>% dplyr::select(
      patient_uuid,
      sex,
      other_epilepsy_mutations,
      recurrent_variant,
      neonatal_onset,
      age_admission_yr,
      admission_age_days_firstDate,
      tech_assistance_ever,
      vns_insertion_ever,
      n_distinct_asm_admission
    ),
    by = c(
      "patient_uuid",
      "admission_age_days_firstDate"
    )
  )

#' 7. re-derive covariates that may change with new periods in merged records
#' (i.e. length of stay and adverse effects during hospitalization)

#' 7.a. LOHS and length of stay categories
admissions_sensit_same_day_lohs <- hospitalizations_sensit_same_day_joined_cov %>%
  mutate(
    length_of_stay_days =
      (admission_age_days_lastDate - admission_age_days_firstDate) + 1,
    length_of_stay_days_cat = cut(
      length_of_stay_days,
      quantile(length_of_stay_days, c(0:3 / 3)),
      include.lowest = TRUE,
      labels = c("Bottom", "Middle", "Top")
    )
  ) %>%
  dplyr::select(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    length_of_stay_days,
    length_of_stay_days_cat
  )


#' 7.b. Respiratory support during admission
## -- derive first therapeutic procedures and clinical diagnoses
therapeutic_rec_same_day <- hospitalizations_sensit_same_day_joined_cov %>%
  # join therapeutic procedures with hospital admissions
  left_join(
    rare_disease_therapeutic_procedures,
    relationship = "many-to-many",
    by = "patient_uuid"
  ) %>%
  # compute overlap flags (either a date or an interval)
  mutate(
    procedure_overlap = if_else(
      !is.na(procedure_age_days_firstDate) &
        !is.na(procedure_age_days_lastDate) &
        !is.na(admission_age_days_firstDate) &
        !is.na(admission_age_days_lastDate) &
        (admission_age_days_firstDate <= procedure_age_days_lastDate) &
        (admission_age_days_lastDate >= procedure_age_days_firstDate),
      TRUE, FALSE
    ),
    # keep procedure name only if overlap
    procedure = if_else(procedure_overlap, procedure, NA_character_)
  ) %>%
  # group by admission
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  # summarise overlapping procedures (or NA if none)
  summarise(
    ther_procedure_name = str_c(
      unique(na.omit(procedure)),
      collapse = "; "
    ),
    .groups = "drop"
  ) %>%
  # rename to procedure for binding rows with clinical diagnoses
  rename(procedure = ther_procedure_name) %>%
  arrange(
    patient_uuid,
    admission_age_days_firstDate
  )


## -- idenfity clinical diagnoses during admissions
clinical_diagnosis_same_day <- hospitalizations_sensit_same_day_joined_cov %>%
  # Joining clinical diagnosis with hospital admission by patient UUID
  left_join(
    rare_disease_clinical_diagnosis,
    relationship = "many-to-many",
    by = "patient_uuid"
  ) %>%
  # Boolean to determine if the procedure has an explicit period that matches the admission period
  mutate(

    #' Boolean to determine if the clinical diagnosis (procedure) has an
    #' explicit period that matches the admission period in at least one day
    procedure_overlap = if_else(
      !is.na(clinical_diagnosis_age_days_firstDate) &
        !is.na(clinical_diagnosis_age_days_lastDate) &
        !is.na(admission_age_days_firstDate) &
        !is.na(admission_age_days_lastDate) &
        (admission_age_days_firstDate <= clinical_diagnosis_age_days_lastDate) &
        (admission_age_days_lastDate >= clinical_diagnosis_age_days_firstDate),
      TRUE, FALSE
    ),
    clinical_diagnosis = if_else(
      procedure_overlap, clinical_diagnosis, NA_character_
    )
  ) %>%
  #' Filter to retain procedures at hospital admissions:
  #' either a single-day procedure falls within admission period
  #' or the procedure period overlaps with the admission period
  # filter(procedure_overlap) %>%
  # Group by patient and admission periods
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  # Concatenate distinct values into a single cell as temporary variable
  summarise(
    clinical_diagnosis = str_c(
      unique(na.omit(clinical_diagnosis)),
      collapse = "; "
    ),
    .groups = "drop"
  ) %>%
  # rename to procedure for binding rows with therapeutic procedures
  rename(procedure = clinical_diagnosis) %>%
  arrange(patient_uuid, admission_age_days_firstDate)

## -- identify resp.support during admissions
admissions_sensit_same_day_resp_supp <- rbind(
  clinical_diagnosis_same_day,
  therapeutic_rec_same_day
) %>%
  # add a flag if procedure matches respiratory support terms
  mutate(
    respiratory_support_flag = case_when(
      str_detect(
        procedure,
        regex(str_c(RESP_SUPPORT_TERMS, collapse = "|"), ignore_case = TRUE)
      ) &
        !str_detect(
          procedure,
          regex("hypoventilation", ignore_case = TRUE)
        ) ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>%
  # group by admission
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  # summarise as binary factor: "1" if any respiratory support, else "0"
  summarise(
   respiratory_support_admission = factor(
      any(respiratory_support_flag, na.rm = TRUE) * 1,
       levels = c(0, 1)
    ),
  .groups = "drop"
  ) %>%
  arrange(patient_uuid, admission_age_days_firstDate)

#' 7.c. adverse events during admissions
admissions_sensit_same_day_aes <- hospitalizations_sensit_same_day_joined_cov %>%
  # Join adverse effects episodes
  left_join(
    rare_disease_adverse_effects,
    by = "patient_uuid",
    relationship = "many-to-many"
  ) %>%
  distinct(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    adverse_effect,
    medication_procedure,
    adverse_effect_age_days_firstDate,
    adverse_effect_age_days_lastDate
  ) %>%
  # Boolean overlap: adverse event overlaps with admission
  mutate(
    adverse_effect_overlap = if_else(
      !is.na(adverse_effect_age_days_firstDate) &
        !is.na(adverse_effect_age_days_lastDate) &
        !is.na(admission_age_days_firstDate) &
        !is.na(admission_age_days_lastDate) &
        (adverse_effect_age_days_firstDate <= admission_age_days_lastDate) &
        (adverse_effect_age_days_lastDate >= admission_age_days_firstDate),
      TRUE, FALSE
    )
  ) %>%
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  # Add binary variable (0/1) for adverse effect presence during admission
  # Note: for those admissions without AE, it was assumed no event (i.e., 0)
  summarise(
    adverse_effect_admission = factor(
      # if there is a match between the adverse effect and the admission
      # period, then it is an 1, otherwise 0
      any(adverse_effect_overlap, na.rm = TRUE) * 1),
    .groups = "drop"
  )  

#' 7.d- ketogenic diet during admission

#' 7.d. ketogenic diet during admission
# a. Derivation of ketogenic diet during admission period for 30-day and lohs analyses
adm_keto_overlap_same_day <- hospitalizations_sensit_same_day_joined_cov %>%
  left_join(
    ketogenic_diet_episodes,
# Check for overlap between *effective* procedure window and admission
  #' any part of the period starting from day 14 to the end of the
  #' ketogenic treatment overlaps with the admission
    join_by(
      patient_uuid,
      overlaps(
        admission_age_days_firstDate,
        admission_age_days_lastDate,
        keto_effective_first_day,
        keto_effective_last_day
      )
    )
  ) %>%
    # Summarise as binary presence indicator per admission
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
    #' Summarise a binary variable indicating was under ketogenic diet
  #'  for each hospital admission record (at least one day of match)
  summarise(
    #' if there is a match between the effective period of the
    #' ketogenic diet and the admission period, then it is an 1
    ketogenic_diet_admission = factor(any(!is.na(keto_effective_first_day)) * 1),
    .groups = "drop"
  )

# b. Derivation of ketogenic diet at admission day for sr analysis
adm_keto_on_admday_same_day <- hospitalizations_sensit_same_day_joined_cov %>%
  left_join(
    ketogenic_diet_episodes,
    # Check for overlap between *effective* procedure window and admission
  #' any part of the period starting from day 14 to the end of the
  #' ketogenic treatment overlaps with the admission
    join_by(
      patient_uuid,
      between(admission_age_days_firstDate,
              keto_effective_first_day,
              keto_effective_last_day)
    )
  ) %>%
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  #' Summarise a binary variable indicating was under ketogenic diet
  #'  for each hospital admission record (if there is a match at the day of admission)
  summarise(
    #' if the admission day is within the effective period, yes = 1, otherwise 0
    ketogenic_diet_at_admission_day = factor(any(!is.na(keto_effective_first_day)) * 1),
    .groups = "drop"
  )

admissions_sensit_same_day_keto <- adm_keto_overlap_same_day %>%
  left_join(
    adm_keto_on_admday_same_day,
    by = c(
      "patient_uuid",
      "admission_age_days_firstDate",
      "admission_age_days_lastDate"
    )
  )

#' 7.e. Quinidine during admission
#' a) Identify admissions with at least one day matching during admission episodr 
#' or at the day of admission match with quinidine use. 
quinidine_during_admission_rec_any_same_day <- hospitalizations_sensit_same_day_joined_cov %>%
  # Join hospital admissions and quinidine episodes on patient_uuid and overlap of periods
  left_join(
    quinidine_use_episodes,
    join_by(
      patient_uuid,
      overlaps(
        admission_age_days_firstDate,
        admission_age_days_lastDate,
        medication_age_days_firstDate,
        medication_age_days_lastDate
      )
    )
  ) %>%
  # Group by patient and admission periods
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  #' Summarise a binary variable indicating quinidine use for each hospital 
  #' admission record (at least one day of match)
  summarise(
    # if there is a match between the quinidine use and the admission
    # period, then it is an 1, otherwise 0
    quinidine_admission = factor(any(!is.na(medication_age_days_firstDate)) * 1),
    .groups = "drop"
  )

#2.a identify admissions with quinidine already active on the admission day
quinidine_during_admission_rec_on_day_same_day <- hospitalizations_sensit_same_day_joined_cov %>%
  # Join hospital admissions and quinidine episodes where admission start is within the quinidine use period
  left_join(
    quinidine_use_episodes,
    join_by(
      patient_uuid,
      between(
        admission_age_days_firstDate,
        medication_age_days_firstDate,
        medication_age_days_lastDate
      )
    )
  ) %>%
  # Group by patient and admission periods
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  #' Summarise a binary variable indicating quinidine use at the admission day
  summarise(
    quinidine_at_admission_day = factor(any(!is.na(medication_age_days_firstDate)) * 1),
    .groups = "drop"
  )

#2.b merge both variables into the final df (keep original name)
admissions_sensit_same_day_quinid <- quinidine_during_admission_rec_any_same_day %>%
  left_join(
    quinidine_during_admission_rec_on_day_same_day,
    by = c(
      "patient_uuid",
      "admission_age_days_firstDate",
      "admission_age_days_lastDate"
    )
  )

# 8- List of period-varying covariates to join
sensit_same_day_time_dfs <- list(
  admissions_sensit_same_day_lohs, # lohs and lohs categories
  admissions_sensit_same_day_resp_supp, # respiratory support variable
  admissions_sensit_same_day_aes, # adverse effects admission variable
  admissions_sensit_same_day_keto, # ketogenic diet admission variable
  admissions_sensit_same_day_quinid # quinidine use admission variable
)

# 9-Joining hospitalization-varying covariates for sensitivity analyses (same day)
admissions_sensit_same_day_full <- purrr::reduce(
  sensit_same_day_time_dfs,
  left_join,
  .init = hospitalizations_sensit_same_day_joined_cov, # start from prior joined df
  by = admission_join_var # based on start, end and patient uuid
)

# 10- Order variables in same way as in main df
#' Note. this dataframe can be used for all secondary endpoints, and for its use
#' in 30-day UHRss, the index in readmissions should be filtered.
admissions_sensit_same_day_full <- admissions_sensit_same_day_full %>%
  mutate(
    patient_uuid = as.factor(patient_uuid)
  ) %>%
  # select and order variables of interest in joined df
  dplyr::select(
    # patient identifier
    patient_uuid,
    # admissions related features
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    # demographics
    sex,
    age_admission_yr,
    # follow up and disease onset
    neonatal_onset,
    # genetic mutations
    other_epilepsy_mutations,
    recurrent_variant,
    # clinical history
    seizure_related_admission,
    length_of_stay_days,
    length_of_stay_days_cat,
    respiratory_support_admission,
    tech_assistance_ever,
    readmission,
    count_readmissions,
    exptime,
    # treatment history
    vns_insertion_ever,
    n_distinct_asm_admission,
    adverse_effect_admission,
    ketogenic_diet_admission,
    ketogenic_diet_at_admission_day,
    quinidine_admission,
    quinidine_at_admission_day
  ) %>%
  # arrange by patient uuid and admission day
  arrange(patient_uuid, admission_age_days_firstDate)
```

#### Hospitalization occurs on the NEXT day as the prior discharge day

```{r}
#| label: admissions-merged-episodes-nextday
#| echo: false
#| results: false
#' gap_threshold = 1 → next hospitalization record starts on the *next day (+1)*
#'  after the previous discharge (for the same patient and with the same type
#'   of sr diagnosis)

# 1. Apply the episode-flagging function across all gap thresholds
#' Note: the function developed is in function.R
hospital_admission_flag_next_day <- flag_admission_episodes(
  hospitalizations_no_merged,
  gap_threshold = 1
) # next day

# 2. Merge admissions within the same episode
#' Internal note: 4 more records could be merged if removing the seizure-type
#' condition for merging
hospitalization_merged_next_day <- hospital_admission_flag_next_day %>%
  # Keep only unplanned, non-birth admissions
  filter(
    birth_admission == 0 &
    planned_admission == 0 &
    outpatient_admission == 0) %>%
  group_by(
    patient_uuid, # by patient
    episode_id, # by episode identified earlier
    seizure_related_admission # and by seizure-related flag
  ) %>%
  # merge admissions within patient, same episode and seizure diagnosis type
  summarise(
    # earliest start in episode
    admission_age_days_firstDate = min(admission_age_days_firstDate),
    # latest end in episode
    admission_age_days_lastDate = max(admission_age_days_lastDate),
    .groups = "drop"
  )

# 3. Re-run the count of readmissions with new merged records
readmissions_sensit_next_day <- hospitalization_merged_next_day %>%
  dplyr::select(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  find_readmissions(
    grouping = "patient_uuid",
    start_date = "admission_age_days_firstDate",
    end_date = "admission_age_days_lastDate"
  ) %>%
  arrange(patient_uuid, admission_age_days_firstDate)

# 4. re-derive count of 30-day UHRs per index admission with merged records
readmissions_sensit_next_day_counted <- readmissions_sensit_next_day %>%
  # Bring in deceased patient information
  left_join(deceased_patients, by = "patient_uuid") %>%
  # Bring last record age days to calculate last exposure time at risk
  left_join(
    rare_disease_followup_durations_days %>% dplyr::select(
      patient_uuid,
      last_record_age_days
    ),
    by = "patient_uuid"
  ) %>%
  #' Exclude admission record if a patient is deceased during the
  #' admission period, but retain prior ones for that patient
  filter(!(patient_uuid %in% deceased_patients &
    primary_diagnosis_age_days >= admission_age_days_firstDate &
    primary_diagnosis_age_days <= admission_age_days_lastDate)) %>%
  arrange(patient_uuid, admission_age_days_firstDate) %>%
  group_by(patient_uuid) %>%
  # identify 30-day readmissions for each index admission
  mutate(
    # flag index admissions based on readmission variable
    index_admission = if_else(readmission == "index", 1, 0),

    # create a sequential number for each index admission within patient
    index_number = cumsum(index_admission),

    # flag last index admission for each patient for last exptime derivation
    last_index = if_else(index_number == max(index_number), 1, 0),

    # flag 30-day readmissions
    is_uhr = str_detect(readmission, "30_days_uhr")
  ) %>%
  # Group by patient and index number
  group_by(patient_uuid, index_number) %>%
  mutate(
    # count readmissions for each index
    count_readmissions = sum(is_uhr, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  #  Keep only index admissions
  filter(index_admission == 1) %>%
  #' Define exposure time (exptime) for each index admission:
  #' Use 30 days as default  unless censored earlier for last index admission
  mutate(
    exptime = as.integer(
      case_when(
        last_index == 1 &
          (last_record_age_days - admission_age_days_lastDate < 30) ~
          last_record_age_days - admission_age_days_lastDate,
        TRUE ~ 30
      )
    )
  ) %>%
  #' Exclude exp time = 0 (i.e., last record == admission last day),
  #' as there is no time to follow up after last index admission
  filter(exptime > 0) %>%
  # order, rename and select variables of interest
  dplyr::select(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    count_readmissions,
    exptime
  )

# 5. Joined dataframe with merged hospitalizations and readmissions
hospitalizations_sensit_next_day_joined <- purrr::reduce(
  list(
    hospitalization_merged_next_day,
    readmissions_sensit_next_day,
    readmissions_sensit_next_day_counted
  ),
  left_join,
  by = c(
    "patient_uuid",
    "admission_age_days_firstDate",
    "admission_age_days_lastDate"
  )
)

#' 6. join previously derived covariates (from main df) that do not change with
#'  the merging and depends only on the patient id and admission start day
#'  (not period dependent)
hospitalizations_sensit_next_day_joined_cov <- hospitalizations_sensit_next_day_joined %>%
  left_join(
    hospitalization_main_df %>% dplyr::select(
      patient_uuid,
      sex,
      other_epilepsy_mutations,
      recurrent_variant,
      neonatal_onset,
      age_admission_yr,
      admission_age_days_firstDate,
      tech_assistance_ever,
      vns_insertion_ever,
      n_distinct_asm_admission
    ),
    by = c(
      "patient_uuid",
      "admission_age_days_firstDate"
    )
  )

#' 7. re-derive covariates that may change with new periods in merged records
#' (i.e. length of stay and adverse effect during hospitalization)
#' 6.q. LOHS and length of stay categories
admissions_sensit_next_day_lohs <- hospitalizations_sensit_next_day_joined %>%
  mutate(
    # Derivation of continuous length of hospital stays in days
    length_of_stay_days =
      (admission_age_days_lastDate - admission_age_days_firstDate) + 1,

    #' Derivation of length of hospital stays in 3 categories by tertiles
    length_of_stay_days_cat = cut(
      length_of_stay_days,
      # cut in tertiles
      quantile(length_of_stay_days, c(0:3 / 3)),
      include.lowest = T,
      labels = c("Bottom", "Middle", "Top")
    )
  ) %>%
  # select variables of interest
  dplyr::select(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    length_of_stay_days,
    length_of_stay_days_cat
  )

#' 7.b.- respiratory support during admission
#--derive first therapeutic procedures and clinical diagnoses
therapeutic_rec_next_day <- hospitalizations_sensit_next_day_joined %>%
  # join therapeutic procedures with hospital admissions
  left_join(
    rare_disease_therapeutic_procedures,
    relationship = "many-to-many",
    by = "patient_uuid"
  ) %>%
  # compute overlap flags (either a date or an interval)
  mutate(
    procedure_overlap = if_else(
      !is.na(procedure_age_days_firstDate) &
        !is.na(procedure_age_days_lastDate) &
        !is.na(admission_age_days_firstDate) &
        !is.na(admission_age_days_lastDate) &
        (admission_age_days_firstDate <= procedure_age_days_lastDate) &
        (admission_age_days_lastDate >= procedure_age_days_firstDate),
      TRUE, FALSE
    ),
    # keep procedure name only if overlap
    procedure = if_else(procedure_overlap, procedure, NA_character_)
  ) %>%
  # group by admission
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  # summarise overlapping procedures (or NA if none)
  summarise(
    ther_procedure_name = str_c(
      unique(na.omit(procedure)),
      collapse = "; "
    ),
    .groups = "drop"
  ) %>%
  # rename to procedure for binding rows with clinical diagnoses
  rename(procedure = ther_procedure_name) %>%
  arrange(
    patient_uuid,
    admission_age_days_firstDate
  )


## -- idenfity clinical diagnoses during admissions
clinical_diagnosis_next_day <- hospitalizations_sensit_next_day_joined %>%
  # Joining clinical diagnosis with hospital admission by patient UUID
  left_join(
    rare_disease_clinical_diagnosis,
    relationship = "many-to-many",
    by = "patient_uuid"
  ) %>%
  # Boolean to determine if the procedure has an explicit period that matches the admission period
  mutate(

    #' Boolean to determine if the clinical diagnosis (procedure) has an
    #' explicit period that matches the admission period in at least one day
    procedure_overlap = if_else(
      !is.na(clinical_diagnosis_age_days_firstDate) &
        !is.na(clinical_diagnosis_age_days_lastDate) &
        !is.na(admission_age_days_firstDate) &
        !is.na(admission_age_days_lastDate) &
        (admission_age_days_firstDate <= clinical_diagnosis_age_days_lastDate) &
        (admission_age_days_lastDate >= clinical_diagnosis_age_days_firstDate),
      TRUE, FALSE
    ),
    clinical_diagnosis = if_else(
      procedure_overlap, clinical_diagnosis, NA_character_
    )
  ) %>%
  #' Filter to retain procedures at hospital admissions:
  #' either a single-day procedure falls within admission period
  #' or the procedure period overlaps with the admission period
  # filter(procedure_overlap) %>%
  # Group by patient and admission periods
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  # Concatenate distinct values into a single cell as temporary variable
  summarise(
    clinical_diagnosis = str_c(
      unique(na.omit(clinical_diagnosis)),
      collapse = "; "
    ),
    .groups = "drop"
  ) %>%
  # rename to procedure for binding rows with therapeutic procedures
  rename(procedure = clinical_diagnosis) %>%
  arrange(patient_uuid, admission_age_days_firstDate)

## -- identify resp.support during admissions
admissions_sensit_next_day_resp_supp <- rbind(
  clinical_diagnosis_next_day,
  therapeutic_rec_next_day
) %>%
  # add a flag if procedure matches respiratory support terms
  mutate(
    respiratory_support_flag = case_when(
      str_detect(
        procedure,
        regex(str_c(RESP_SUPPORT_TERMS, collapse = "|"), ignore_case = TRUE)
      ) &
        !str_detect(
          procedure,
          regex("hypoventilation", ignore_case = TRUE)
        ) ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>%
  # group by admission
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  # summaries as binary factor: "1" if any respiratory support, else "0"
  summarise(
   respiratory_support_admission = factor(
      any(respiratory_support_flag, na.rm = TRUE) * 1,
       levels = c(0, 1)
    ),
  .groups = "drop"
  ) %>%
  arrange(patient_uuid, admission_age_days_firstDate)

#' 7.c. adverse effect during admissions
admissions_sensit_next_day_aes <- hospitalizations_sensit_next_day_joined %>%
  # Join adverse effect episodes
  left_join(
    rare_disease_adverse_effects,
    by = "patient_uuid",
    relationship = "many-to-many"
  ) %>%
  distinct(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    adverse_effect,
    medication_procedure,
    adverse_effect_age_days_firstDate,
    adverse_effect_age_days_lastDate
  ) %>%
  # Boolean overlap: adverse effect overlaps with admission
  mutate(
    adverse_effect_overlap = if_else(
      !is.na(adverse_effect_age_days_firstDate) &
        !is.na(adverse_effect_age_days_lastDate) &
        !is.na(admission_age_days_firstDate) &
        !is.na(admission_age_days_lastDate) &
        (adverse_effect_age_days_firstDate <= admission_age_days_lastDate) &
        (adverse_effect_age_days_lastDate >= admission_age_days_firstDate),
      TRUE, FALSE
    )
  ) %>%
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  # Add binary variable (0/1) for adverse effect presence during admission
  # Note: for those admissions without AE, it was assumed no event (i.e., 0)
  summarise(
    adverse_effect_admission = factor(
      # if there is a match between the adverse effect and the admission
      # period, then it is an 1, otherwise 0
      any(adverse_effect_overlap, na.rm = TRUE) * 1),
    .groups = "drop"
  )  

#' 7.d. ketogenic diet during admission
# a. Derivation of ketogenic diet during admission period for 30-day and lohs analyses
adm_keto_overlap_next_day <- hospitalizations_sensit_next_day_joined %>%
  left_join(
    ketogenic_diet_episodes,
# Check for overlap between *effective* procedure window and admission
  #' any part of the period starting from day 14 to the end of the
  #' ketogenic treatment overlaps with the admission
    join_by(
      patient_uuid,
      overlaps(
        admission_age_days_firstDate,
        admission_age_days_lastDate,
        keto_effective_first_day,
        keto_effective_last_day
      )
    )
  ) %>%
    # Summarise as binary presence indicator per admission
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
    #' Summarise a binary variable indicating was under ketogenic diet
  #'  for each hospital admission record (at least one day of match)
  summarise(
    #' if there is a match between the effective period of the
    #' ketogenic diet and the admission period, then it is an 1
    ketogenic_diet_admission = factor(any(!is.na(keto_effective_first_day)) * 1),
    .groups = "drop"
  )

# b. Derivation of ketogenic diet at admission day for sr analysis
adm_keto_on_admday_next_day <- hospitalizations_sensit_next_day_joined %>%
  left_join(
    ketogenic_diet_episodes,
    # Check for overlap between *effective* procedure window and admission
  #' any part of the period starting from day 14 to the end of the
  #' ketogenic treatment overlaps with the admission
    join_by(
      patient_uuid,
      between(admission_age_days_firstDate,
              keto_effective_first_day,
              keto_effective_last_day)
    )
  ) %>%
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  #' Summarise a binary variable indicating was under ketogenic diet
  #'  for each hospital admission record (if there is a match at the day of admission)
  summarise(
    #' if the admission day is within the effective period, yes = 1, otherwise 0
    ketogenic_diet_at_admission_day = factor(any(!is.na(keto_effective_first_day)) * 1),
    .groups = "drop"
  )

admissions_sensit_next_day_keto <- adm_keto_overlap_next_day %>%
  left_join(
    adm_keto_on_admday_next_day,
    by = c(
      "patient_uuid",
      "admission_age_days_firstDate",
      "admission_age_days_lastDate"
    )
  )

#' 7.e. Quinidine during admission

#2. identify admissions with matching use of quinidine in at least one day
#   (using inequality join with overlaps)

#' a) Identify admissions with at least one day matching during admission episodr 
#' or at the day of admission match with quinidine use. 
quinidine_during_admission_rec_any_next_day <- hospitalizations_sensit_next_day_joined %>%
  # Join hospital admissions and quinidine episodes on patient_uuid and overlap of periods
  left_join(
    quinidine_use_episodes,
    join_by(
      patient_uuid,
      overlaps(
        admission_age_days_firstDate,
        admission_age_days_lastDate,
        medication_age_days_firstDate,
        medication_age_days_lastDate
      )
    )
  ) %>%
  # Group by patient and admission periods
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  #' Summarise a binary variable indicating quinidine use for each hospital 
  #' admission record (at least one day of match)
  summarise(
    # if there is a match between the quinidine use and the admission
    # period, then it is an 1, otherwise 0
    quinidine_admission = factor(any(!is.na(medication_age_days_firstDate)) * 1),
    .groups = "drop"
  )

#2.a identify admissions with quinidine already active on the admission day
quinidine_during_admission_rec_on_day_next_day <- hospitalizations_sensit_next_day_joined %>%
  # Join hospital admissions and quinidine episodes where admission start is within the quinidine use period
  left_join(
    quinidine_use_episodes,
    join_by(
      patient_uuid,
      between(
        admission_age_days_firstDate,
        medication_age_days_firstDate,
        medication_age_days_lastDate
      )
    )
  ) %>%
  # Group by patient and admission periods
  group_by(
    patient_uuid,
    admission_age_days_firstDate,
    admission_age_days_lastDate
  ) %>%
  #' Summarise a binary variable indicating quinidine use at the admission day
  summarise(
    quinidine_at_admission_day = factor(any(!is.na(medication_age_days_firstDate)) * 1),
    .groups = "drop"
  )

#2.b merge both variables into the final df (keep original name)
admissions_sensit_next_day_quinid <- quinidine_during_admission_rec_any_next_day %>%
  left_join(
    quinidine_during_admission_rec_on_day_next_day,
    by = c(
      "patient_uuid",
      "admission_age_days_firstDate",
      "admission_age_days_lastDate"
    )
  )

# 8- List of time-varying covariates to join
sensit_next_day_time_dfs <- list(
  admissions_sensit_next_day_lohs, # lohs and lohs categories
  admissions_sensit_next_day_resp_supp, # respiratory support variable
  admissions_sensit_next_day_aes, # adverse effect admission variable
  admissions_sensit_next_day_keto, # ketogenic diet admission variable
  admissions_sensit_next_day_quinid # quinidine use admission variable
)

# 9- joining hosptialization-varying covariates for sensitivity analyses (next day)
admissions_sensit_next_day_full <- purrr::reduce(
  sensit_next_day_time_dfs,
  left_join,
  .init = hospitalizations_sensit_next_day_joined_cov, # start from previos joined
  by = admission_join_var # based on start, end and patient uuid
)

#10- order variables in same way as in main df
#' Note: this dataframe can be used for all secondary endpoints, and for its use
#' in 30-day UHRss, the index in readmissions should be filtered.
admissions_sensit_next_day_full <- admissions_sensit_next_day_full %>%
  # select and order variables of interest in joined df
  mutate(
    patient_uuid = as.factor(patient_uuid)
  ) %>%
  # select and order variables of interest in joined df
  dplyr::select(
    # patient identifier
    patient_uuid,
    # admissions related features
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    # demographics
    sex,
    age_admission_yr,
    # follow up and disease onset
    neonatal_onset,
    # genetic mutations
    other_epilepsy_mutations,
    recurrent_variant,
    # clinical history
    seizure_related_admission,
    length_of_stay_days,
    length_of_stay_days_cat,
    respiratory_support_admission,
    tech_assistance_ever,
    readmission,
    count_readmissions,
    exptime,
    # treatment history
    vns_insertion_ever,
    n_distinct_asm_admission,
    adverse_effect_admission,
    ketogenic_diet_admission,
    ketogenic_diet_at_admission_day,
    quinidine_admission,
    quinidine_at_admission_day
  ) %>%
  # arrange by patient uuid and admission day
  arrange(patient_uuid, admission_age_days_firstDate)
```

```{r}
#| label: save-sensitivity-df-file-admissions-merging
#| results: false
#| warning: false
#| message: false

# create temp dir to save file
dir.create("~/RWE-rare_disease/Hospitalizations/output/data")

# save df as rds file
saveRDS(admissions_sensit_same_day_full, "data/admissions_sensit_same_day_full.rds")
saveRDS(admissions_sensit_next_day_full, "data/admissions_sensit_next_day_full.rds")

# Upload final df to GCS
gcs_upload("~/RWE-rare_disease/Hospitalizations/output/data/admissions_sensit_same_day_full.rds",
  bucket = "YOUR_GCS_BUCKET",
  name = "MSM_Hospitalizations/admissions_sensit_same_day_full.rds"
)

gcs_upload("~/RWE-rare_disease/Hospitalizations/output/data/admissions_sensit_next_day_full.rds",
  bucket = "YOUR_GCS_BUCKET",
  name = "MSM_Hospitalizations/admissions_sensit_next_day_full.rds"
)
```

### Including planned admissions

```{r}
#| label: include-planned-admissions-df-lohs-sensitivity-analyses
#| echo: false

#' Definition (SAP):
#' The main analysis will be repeated but extended to include hospital 
#' admissions records categorized as "planned” according to the ‘admission 
#' diagnosis’ variable in the hospital admission table. 
#' Note: this df will be used for both LOHS endpoint and SR admissions and 
#' we start from the main df with all covariates

lohs_sr_df_planned <- hospitalization_main_df %>%
  # do not remove planned admissions, only birth and outpatient admissions
  filter(
    birth_admission == 0, # Exclude birth admissions
    outpatient_admission == 0 # Exclude outpatient admissions
  ) %>%
  dplyr::select(
    # patient identifier
    patient_uuid,
    # admissions related features
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    # demographics
    sex,
    age_admission_yr,
    # follow up and disease onset
    neonatal_onset,
    # genetic mutations
    other_epilepsy_mutations,
    recurrent_variant,
    # clinical history
    seizure_related_admission,
    length_of_stay_days,
    length_of_stay_days_cat,
    respiratory_support_admission,
    tech_assistance_ever,
    # treatment history
    vns_insertion_ever,
    n_distinct_asm_admission,
    adverse_effect_admission,
    ketogenic_diet_admission,
    ketogenic_diet_at_admission_day,
    quinidine_admission,
    quinidine_at_admission_day
    )
```

```{r}
#| label: save-sensitivity-df-file-lohs-and-sr-admissions-planned
#| results: false
#| warning: false
#| message: false

# create temp dir to save file
dir.create("~/RWE-rare_disease/Hospitalizations/output/data")

# save df as rds file
saveRDS(lohs_sr_df_planned, "data/lohs_sr_df_planned.rds")

# Upload final df to GCS
gcs_upload("~/RWE-rare_disease/Hospitalizations/output/data/lohs_sr_df_planned.rds",
  bucket = "YOUR_GCS_BUCKET",
  name = "MSM_Hospitalizations/lohs_sr_df_planned.rds"
)
```

### Excluding outliers (only to LOHS)

```{r}
#| label: outlier-exclusion-df-sensitivity-analyses
#| echo: false

#' Definition (SAP) updated after checking an approximately log-normal distribution of LOHS
#' https://www.sciencedirect.com/science/article/pii/S0031699725074575?ref=cra_js_challenge&fr=RR-1
#' Outliers will be identified based on the distribution of the log-transformed 
#' length of hospital stay (LOHS). 
#' Specifically, any value of log(LOHS) below the first quartile minus 1.5 times
#'  the interquartile range (IQR), or above the third quartile plus 1.5 × IQR,
#'   will be considered an outlier and excluded from the analysis.
#'    
#'    For this sensitivity analysis, we will start from the main df with all
#'     covariates

lohs_df_outliers <- hospitalization_main_df %>%
  filter(
    birth_admission == 0, # Exclude birth admissions
    planned_admission == 0, # Exclude planned admissions
    outpatient_admission == 0 # Exclude outpatient admissions
  ) %>%
  dplyr::select(
    # patient identifier
    patient_uuid,
    # admissions related features
    admission_age_days_firstDate,
    admission_age_days_lastDate,
    # demographics
    sex,
    age_admission_yr,
    # follow up and disease onset
    neonatal_onset,
    # genetic mutations
    other_epilepsy_mutations,
    recurrent_variant,
    # clinical history
    seizure_related_admission,
    length_of_stay_days,
    respiratory_support_admission,
    tech_assistance_ever,
    # treatment history
    vns_insertion_ever,
    n_distinct_asm_admission,
    adverse_effect_admission,
    ketogenic_diet_admission,
    quinidine_admission
    ) %>%
  # add log(LOHS)
  mutate(log_lohs = log(length_of_stay_days)) 

# Compute IQR bounds on the LOG scale
Q1_log  <- quantile(lohs_df_outliers$log_lohs, 0.25, na.rm = TRUE)
Q3_log  <- quantile(lohs_df_outliers$log_lohs, 0.75, na.rm = TRUE)
IQR_log <- Q3_log - Q1_log

lower_log <- max(0, Q1_log - 1.5 * IQR_log)
upper_log <- Q3_log + 1.5 * IQR_log

# Filter rows by keeping ONLY observations within the LOG bounds
log_lohs_df_outliers <- lohs_df_outliers %>%
  filter(log_lohs >= lower_log, log_lohs <= upper_log) %>%
  dplyr::select(-log_lohs)
```

```{r}
#| label: save-sensitivity-df-file-lohs-outliers
#| results: false
#| warning: false
#| message: false

# create temp dir to save file
dir.create("~/RWE-rare_disease/Hospitalizations/output/data")

# save df as rds file
saveRDS(log_lohs_df_outliers, "data/log_lohs_df_outliers.rds")

# Upload final df to GCS
gcs_upload("~/RWE-rare_disease/Hospitalizations/output/data/log_lohs_df_outliers.rds",
  bucket = "YOUR_GCS_BUCKET",
  name = "MSM_Hospitalizations/log_lohs_df_outliers.rds"
)
```

```{r}
#| label: remove-temp-files-end
#| results: false
#| warning: false
#| message: false

# remove temp files
list.files(path = ".", pattern = NULL, all.files = FALSE, full.names = FALSE)

# delete from local files
unlink("~/RWE-rare_disease/Hospitalizations/output/data", recursive = TRUE)
```
